---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "12/29/2024"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/repeat-heat/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/repeat-heat/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("data.table)
#install.packages("RColorBrewer")
require(tidyverse)
require(data.table)
require(lme4)
require(RColorBrewer)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank10a <- read.csv("../../../data/repeat-heat/HOBO/t10a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp10a = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..t10a.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank10b <- read.csv("../../../data/repeat-heat/HOBO/t10b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp10b = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..t10b.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank11a <- read.csv("../../../data/repeat-heat/HOBO/t11a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp11a = Temp...C..LGR.S.N..21435404..SEN.S.N..21435404..LBL..t11a.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank11b <- read.csv("../../../data/repeat-heat/HOBO/t11b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp11b = Temp...C..LGR.S.N..21435408..SEN.S.N..21435408.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank12a <- read.csv("../../../data/repeat-heat/HOBO/t12a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp12a = Temp...C..LGR.S.N..21435413..SEN.S.N..21435413.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank12b <- read.csv("../../../data/repeat-heat/HOBO/t12b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp12b = Temp...C..LGR.S.N..21435401..SEN.S.N..21435401.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank13a <- read.csv("../../../data/repeat-heat/HOBO/t13a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp13a = Temp...C..LGR.S.N..20548754..SEN.S.N..20548754.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank13b <- read.csv("../../../data/repeat-heat/HOBO/t13b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp13b = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..t13b.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment

tank10a <- tank10a %>% 
  mutate(., tank = 10,
         treatment = "control")
tank10b <- tank10b %>% 
  mutate(., tank = 10,
         treatment = "control")

tank11a <- tank11a %>% 
  mutate(., tank = 11,
         treatment = "treatment")
tank11b <- tank11b %>% 
  mutate(., tank = 11,
         treatment = "treatment")

tank12a <- tank12a %>% 
  mutate(., tank = 12,
         treatment = "control")
tank12b <- tank12b %>% 
  mutate(., tank = 12,
         treatment = "control")

tank13a <- tank13a %>% 
  mutate(., tank = 13,
         treatment = "treatment")
tank13b <- tank13b %>% 
  mutate(., tank = 13,
         treatment = "treatment")
```

```{r}
#Format all temperature data files as data.tables

tank10a <- as.data.table(tank10a)
tank10b <- as.data.table(tank10b)

tank11a <- as.data.table(tank11a)
tank11b <- as.data.table(tank11b)

tank12a <- as.data.table(tank12a)
tank12b <- as.data.table(tank12b)

tank13a <- as.data.table(tank13a)
tank13b <- as.data.table(tank13b)
```

13b: 402
11b: 404
12a: 405
11a: 406
12b: 406

```{r}
temp <- tank13a[tank10b, on = "dateTime", roll = T, rollends = F][tank10a, on = "dateTime", roll = T, rollends = F][tank13b, on = "dateTime", roll = T, rollends = F][tank11b, on = "dateTime", roll = T, rollends = F][tank12a, on = "dateTime", roll = T, rollends = F][tank12b, on = "dateTime", roll = T, rollends = F][tank11a, on = "dateTime", roll = T, rollends = F] #Roll data by closest dateTime. List begins with the shortest file and ends with the longest file
temp
```

```{r}
tempClean <- temp %>%
  dplyr::select(dateTime, 
                temp10a, temp10b,
                temp11a, temp11b,
                temp12a, temp12b,
                temp13a, temp13b) #Select necessary columns from rolled data
head(tempClean) #Confirm changes
```

# Calculate treatment conditions

## For each tank

I need to calculate the average temperature for the four periods of the experiment for each treatment. The following lines correspond with each part of the experiment:

Pre-exposure: 1:22
Ramp up: 23:30
Pulse 1: 31:108
Ramp down: 109:112
Rest: 112:291
Ramp up: 292:300
Pulse 2: 301:406

```{r}
treatmentTempStats <- data.frame("tank" = c(10:13),
                                 "meanTempPre" = rep(0, times = 4),
                                 "sdTempPre" = rep(0, times = 4),
                                 "meanTempPulse1" = rep(0, times = 4),
                                 "sdTempPulse1" = rep(0, times = 4),
                                 "meanTempRest" = rep(0, times = 4),
                                 "sdTempRest" = rep(0, times = 4),
                                 "meanTempPulse2" = rep(0, times = 4),
                                 "sdTempPulse2" = rep(0, times = 4)) #Create an empty dataframe
treatmentTempStats #Confirm dataframe creation
```

```{r}
#Pre-exposure: 1:22

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),2] <-
    mean(rbind(tempClean[[i]][1:22], 
               tempClean[[i+1]][1:22]), na.rm = TRUE)
  treatmentTempStats[(i/2),3] <-
    sd(rbind(tempClean[[i]][1:22], 
             tempClean[[i+1]][1:22]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
#Pulse 1: 31:108

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),4] <-
    mean(rbind(tempClean[[i]][31:108], 
               tempClean[[i+1]][31:108]), na.rm = TRUE)
  treatmentTempStats[(i/2),5] <-
    sd(rbind(tempClean[[i]][31:108], 
             tempClean[[i+1]][31:108]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
#Rest: 112:291

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),6] <-
    mean(rbind(tempClean[[i]][112:291], 
               tempClean[[i+1]][112:291]), na.rm = TRUE)
  treatmentTempStats[(i/2),7] <-
    sd(rbind(tempClean[[i]][112:291], 
             tempClean[[i+1]][112:291]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
#Pulse 2: 301:406

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),8] <-
    mean(rbind(tempClean[[i]][301:406], 
               tempClean[[i+1]][301:406]), na.rm = TRUE)
  treatmentTempStats[(i/2),9] <-
    sd(rbind(tempClean[[i]][301:406], 
             tempClean[[i+1]][301:406]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

## For each treatment

```{r}
treatmentTempStatsOverall <- data.frame("treatment" = c("control", "treatment"),
                                        "meanTempPre" = rep(0, times = 2),
                                        "sdTempPre" = rep(0, times = 2),
                                        "meanTempPulse1" = rep(0, times = 2),
                                        "sdTempPulse1" = rep(0, times = 2),
                                        "meanTempRest" = rep(0, times = 2),
                                        "sdTempRest" = rep(0, times = 2),
                                        "meanTempPulse2" = rep(0, times = 2),
                                        "sdTempPulse2" = rep(0, times = 2)) #Create an empty dataframe
treatmentTempStatsOverall #Confirm dataframe creation
```

```{r}
#Pre-exposure: 1:22

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),2] <-
    mean(rbind(tempClean[[i]][1:22], 
               tempClean[[i+1]][1:22],
               tempClean[[i+4]][1:22], 
               tempClean[[i+5]][1:22]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),3] <-
    sd(rbind(tempClean[[i]][1:22], 
             tempClean[[i+1]][1:22],
             tempClean[[i+4]][1:22], 
             tempClean[[i+5]][1:22]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
#Pulse 1: 31:108

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),4] <-
    mean(rbind(tempClean[[i]][31:108], 
               tempClean[[i+1]][31:108],
               tempClean[[i+4]][31:108], 
               tempClean[[i+5]][31:108]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),5] <-
    sd(rbind(tempClean[[i]][31:108], 
             tempClean[[i+1]][31:108],
             tempClean[[i+4]][31:108], 
             tempClean[[i+5]][31:108]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
#Rest: 112:291

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),6] <-
    mean(rbind(tempClean[[i]][112:291], 
               tempClean[[i+1]][112:291],
               tempClean[[i+4]][112:291], 
               tempClean[[i+5]][112:291]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),7] <-
    sd(rbind(tempClean[[i]][112:291], 
             tempClean[[i+1]][112:291],
             tempClean[[i+4]][112:291], 
             tempClean[[i+5]][112:291]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
#Pulse 2: 301:406

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),8] <-
    mean(rbind(tempClean[[i]][301:406], 
               tempClean[[i+1]][301:406],
               tempClean[[i+4]][301:406], 
               tempClean[[i+5]][301:406]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),9] <-
    sd(rbind(tempClean[[i]][301:406], 
             tempClean[[i+1]][301:406],
             tempClean[[i+4]][301:406], 
             tempClean[[i+5]][301:406]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "overall-temp-mean-sd.csv", quote = FALSE)
```

# Statistical test

I will conduct a one-way ANOVA to confirm there are significant differences between treatments during the first pulse, and to confirm that there are no significant differences between treatments during the pre-ramp, rest period, and second pulse.

## Format data

I need to make an individual data frame for each section of the experiment (pre-condition, pulse 1, rest, and pulse 2).

```{r}
tempPreExpLong <- tempClean[4:22,] %>%
  pivot_longer(., cols = temp10a:temp13b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp10a" | logger == "temp10b" ~ "tank10",
                             logger == "temp11a" | logger == "temp11b" ~ "tank11",
                             logger == "temp12a" | logger == "temp12b" ~ "tank12",
                             logger == "temp13a" | logger == "temp13b" ~ "tank13")) %>%
  mutate(., condition = case_when(tank == "tank10" | tank == "tank12" ~ "control",
                                  tank == "tank11" | tank == "tank13" ~ "treatment")) %>% 
  dplyr::select(-logger) %>% 
  drop_na() #Take pre-exposure data and pivot longer. Create tank and condition columns and remove logger column. Drop rows with NAs
head(tempPreExpLong)
```

```{r}
tempPulse1Long <- tempClean[31:108,] %>%
  pivot_longer(., cols = temp10a:temp13b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp10a" | logger == "temp10b" ~ "tank10",
                             logger == "temp11a" | logger == "temp11b" ~ "tank11",
                             logger == "temp12a" | logger == "temp12b" ~ "tank12",
                             logger == "temp13a" | logger == "temp13b" ~ "tank13")) %>%
  mutate(., condition = case_when(tank == "tank10" | tank == "tank12" ~ "control",
                                  tank == "tank11" | tank == "tank13" ~ "treatment")) %>% 
  dplyr::select(-logger) %>% 
  drop_na() #Take pulse 1 data and pivot longer. Create tank and condition columns and remove logger column. Drop rows with NAs
head(tempPulse1Long)
```

```{r}
tempRestLong <- tempClean[112:291,] %>%
  pivot_longer(., cols = temp10a:temp13b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp10a" | logger == "temp10b" ~ "tank10",
                             logger == "temp11a" | logger == "temp11b" ~ "tank11",
                             logger == "temp12a" | logger == "temp12b" ~ "tank12",
                             logger == "temp13a" | logger == "temp13b" ~ "tank13")) %>%
  mutate(., condition = case_when(tank == "tank10" | tank == "tank12" ~ "control",
                                  tank == "tank11" | tank == "tank13" ~ "treatment")) %>% 
  dplyr::select(-logger) %>% 
  drop_na() #Take rest data and pivot longer. Create tank and condition columns and remove logger column. Drop rows with NAs
head(tempRestLong)
```

```{r}
tempPulse2Long <- tempClean[301:406,] %>%
  pivot_longer(., cols = temp10a:temp13b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp10a" | logger == "temp10b" ~ "tank10",
                             logger == "temp11a" | logger == "temp11b" ~ "tank11",
                             logger == "temp12a" | logger == "temp12b" ~ "tank12",
                             logger == "temp13a" | logger == "temp13b" ~ "tank13")) %>%
  mutate(., condition = case_when(tank == "tank10" | tank == "tank12" ~ "control",
                                  tank == "tank11" | tank == "tank13" ~ "treatment")) %>% 
  dplyr::select(-logger) %>% 
  drop_na() #Take pulse 2 data and pivot longer. Create tank and condition columns and remove logger column. Drop rows with NAs
head(tempPulse2Long)
```

## Perform ANOVAs

```{r}
tempPreExpANOVA <- aov(temperature ~ condition, data = tempPreExpLong)
summary(tempPreExpANOVA)[[1]][4][[1]][1] #F = 8.519706
summary(tempPreExpANOVA)[[1]][5][[1]][1] #p = 0.00411743
```

```{r}
tempPulse1ANOVA <- aov(temperature ~ condition, data = tempPulse1Long)
summary(tempPulse1ANOVA)[[1]][4][[1]][1] #F = 41860.97
summary(tempPulse1ANOVA)[[1]][5][[1]][1] #p = 0
```

```{r}
tempRestANOVA <- aov(temperature ~ condition, data = tempRestLong)
summary(tempRestANOVA)[[1]][4][[1]][1] #F = 0.10543
summary(tempRestANOVA)[[1]][5][[1]][1] #p = 0.7454668
```

```{r}
tempPulse2ANOVA <- aov(temperature ~ condition, data = tempPulse2Long)
summary(tempPulse2ANOVA)[[1]][4][[1]][1] #F = 442.4941
summary(tempPulse2ANOVA)[[1]][5][[1]][1] #p = 2.410091e-79
```

As expected, the conditions at Pulse 1 were different between treatments. While the rest conditions were similar between treatments, the pre-exposure and pulse 2 conditions were not. Looking at the mean ± SD, the pre-exposure conditions do slightly overlap. The differences in pulse 2 are likely due to lower temperature in Tank 13. However, those SD bars do overlap a touch at the edges.

# Plot full temperature conditions

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Reds")[7],
                brewer.pal(9, "Greys")[7]) #Create color scheme
```

- add pulse 2 ribbons
-treatment
-control
- add pulse 1 ribbons
-control

```{r}
treatmentTempStatsOverall
```


```{r}
tempClean[4:406,] %>%
  dplyr::select(dateTime, 
                temp10a, temp10b,
                temp11a, temp11b,
                temp12a, temp12b, 
                temp13a, temp13b) %>%
  rowwise() %>%
  mutate(., controlTemp = mean(c(temp10a, temp10b,
                                 temp12a, temp12b), na.rm = TRUE),
         treatmentTemp = mean(c(temp11a, temp11b,
                                temp13a, temp13b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = treatmentTemp)) +
  geom_line(color = plotColors[1]) + #Average treatment temperature
  geom_segment(aes(x = tempClean$dateTime[31], xend = tempClean$dateTime[108], 
                   y = treatmentTempStatsOverall$meanTempPulse1[2], yend = treatmentTempStatsOverall$meanTempPulse1[2]), 
               color = plotColors[1], linetype = 3) + #Pulse 1 treatment segment
  geom_ribbon(data = tempClean[31:108], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse1[2], 
                  ymin = treatmentTempStatsOverall$meanTempPulse1[2] - treatmentTempStatsOverall$sdTempPulse1[2], 
                  ymax = treatmentTempStatsOverall$meanTempPulse1[2] + treatmentTempStatsOverall$sdTempPulse1[2]), 
              fill = plotColors[1], alpha = 0.15) + #Pulse 1 treatment ribbon
  geom_segment(aes(x = tempClean$dateTime[301], xend = tempClean$dateTime[406], 
                   y = treatmentTempStatsOverall$meanTempPulse2[2], yend = treatmentTempStatsOverall$meanTempPulse2[2]), 
               color = plotColors[1], linetype = 3) + #Pulse 2 treatment segment
  geom_ribbon(data = tempClean[301:406], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse2[2], 
                  ymin = treatmentTempStatsOverall$meanTempPulse2[2] - treatmentTempStatsOverall$sdTempPulse2[2], 
                  ymax = treatmentTempStatsOverall$meanTempPulse2[2] + treatmentTempStatsOverall$sdTempPulse2[2]), 
              fill = plotColors[1], alpha = 0.15) + #Pulse 2 treatment ribbon
  geom_line(aes(x = dateTime, y = controlTemp), color = plotColors[2]) + #Control treatment temperature
  geom_segment(aes(x = tempClean$dateTime[31], xend = tempClean$dateTime[108], 
                   y = treatmentTempStatsOverall$meanTempPulse1[1], yend = treatmentTempStatsOverall$meanTempPulse1[1]), 
               color = plotColors[2], linetype = 3) + #Pulse 1 control segment
  geom_ribbon(data = tempClean[31:108], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse1[1], 
                  ymin = treatmentTempStatsOverall$meanTempPulse1[1] - treatmentTempStatsOverall$sdTempPulse1[1], 
                  ymax = treatmentTempStatsOverall$meanTempPulse1[1] + treatmentTempStatsOverall$sdTempPulse1[1]), 
              fill = plotColors[2], alpha = 0.15) + #Pulse 1 control ribbon
  geom_segment(aes(x = tempClean$dateTime[301], xend = tempClean$dateTime[406], 
                   y = treatmentTempStatsOverall$meanTempPulse2[1], yend = treatmentTempStatsOverall$meanTempPulse2[1]), 
               color = plotColors[2], linetype = 3) + #Pulse 2 control segment
  geom_ribbon(data = tempClean[301:406], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse2[1], 
                  ymin = treatmentTempStatsOverall$meanTempPulse2[1] - treatmentTempStatsOverall$sdTempPulse2[1], 
                  ymax = treatmentTempStatsOverall$meanTempPulse2[1] + treatmentTempStatsOverall$sdTempPulse2[1]), 
              fill = plotColors[2], alpha = 0.15) + #Pulse 2 treatment ribbon
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     limits = c(14.9,33),
                     breaks = c(seq(15,30,5), seq(31,33,1))) +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

