---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "12/29/2024"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/repeat-heat/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/repeat-heat/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("data.table)
#install.packages("RColorBrewer")
#install.packages("cowplot")
require(tidyverse)
require(data.table)
require(RColorBrewer)
require(cowplot)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank10a <- read.csv("../../../data/repeat-heat/HOBO/t10a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp10a = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..t10a.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank10b <- read.csv("../../../data/repeat-heat/HOBO/t10b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp10b = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..t10b.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank11a <- read.csv("../../../data/repeat-heat/HOBO/t11a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp11a = Temp...C..LGR.S.N..21435404..SEN.S.N..21435404..LBL..t11a.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank11b <- read.csv("../../../data/repeat-heat/HOBO/t11b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp11b = Temp...C..LGR.S.N..21435408..SEN.S.N..21435408.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank12a <- read.csv("../../../data/repeat-heat/HOBO/t12a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp12a = Temp...C..LGR.S.N..21435413..SEN.S.N..21435413.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank12b <- read.csv("../../../data/repeat-heat/HOBO/t12b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp12b = Temp...C..LGR.S.N..21435401..SEN.S.N..21435401.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank13a <- read.csv("../../../data/repeat-heat/HOBO/t13a.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp13a = Temp...C..LGR.S.N..20548754..SEN.S.N..20548754.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank13b <- read.csv("../../../data/repeat-heat/HOBO/t13b.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp13b = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..t13b.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment

tank10a <- tank10a %>% 
  mutate(., tank = 10,
         treatment = "control")
tank10b <- tank10b %>% 
  mutate(., tank = 10,
         treatment = "control")

tank11a <- tank11a %>% 
  mutate(., tank = 11,
         treatment = "treatment")
tank11b <- tank11b %>% 
  mutate(., tank = 11,
         treatment = "treatment")

tank12a <- tank12a %>% 
  mutate(., tank = 12,
         treatment = "control")
tank12b <- tank12b %>% 
  mutate(., tank = 12,
         treatment = "control")

tank13a <- tank13a %>% 
  mutate(., tank = 13,
         treatment = "treatment")
tank13b <- tank13b %>% 
  mutate(., tank = 13,
         treatment = "treatment")
```

```{r}
#Format all temperature data files as data.tables

tank10a <- as.data.table(tank10a)
tank10b <- as.data.table(tank10b)

tank11a <- as.data.table(tank11a)
tank11b <- as.data.table(tank11b)

tank12a <- as.data.table(tank12a)
tank12b <- as.data.table(tank12b)

tank13a <- as.data.table(tank13a)
tank13b <- as.data.table(tank13b)
```

```{r}
temp <- tank13a[tank10b, on = "dateTime", roll = T, rollends = F][tank10a, on = "dateTime", roll = T, rollends = F][tank13b, on = "dateTime", roll = T, rollends = F][tank11b, on = "dateTime", roll = T, rollends = F][tank12a, on = "dateTime", roll = T, rollends = F][tank12b, on = "dateTime", roll = T, rollends = F][tank11a, on = "dateTime", roll = T, rollends = F] #Roll data by closest dateTime. List begins with the shortest file and ends with the longest file
temp
```

```{r}
tempClean <- temp %>%
  dplyr::select(dateTime, 
                temp10a, temp10b,
                temp11a, temp11b,
                temp12a, temp12b,
                temp13a, temp13b) #Select necessary columns from rolled data
head(tempClean) #Confirm changes
```

# Calculate treatment conditions

## For each tank

I need to calculate the average temperature for the four periods of the experiment for each treatment. The following lines correspond with each part of the experiment:

Pre-exposure: 1:22
Ramp up: 23:30
Pulse 1: 31:108
Ramp down: 109:112
Rest: 112:291
Ramp up: 292:300
Pulse 2: 301:395

```{r}
treatmentTempStats <- data.frame("tank" = c(10:13),
                                 "meanTempPre" = rep(0, times = 4),
                                 "sdTempPre" = rep(0, times = 4),
                                 "meanTempPulse1" = rep(0, times = 4),
                                 "sdTempPulse1" = rep(0, times = 4),
                                 "meanTempRest" = rep(0, times = 4),
                                 "sdTempRest" = rep(0, times = 4),
                                 "meanTempPulse2" = rep(0, times = 4),
                                 "sdTempPulse2" = rep(0, times = 4)) #Create an empty dataframe
treatmentTempStats #Confirm dataframe creation
```

```{r}
#Pre-exposure: 4:22

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),2] <-
    mean(rbind(tempClean[[i]][4:22], 
               tempClean[[i+1]][4:22]), na.rm = TRUE)
  treatmentTempStats[(i/2),3] <-
    sd(rbind(tempClean[[i]][4:22], 
             tempClean[[i+1]][4:22]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
#Pulse 1: 31:108

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),4] <-
    mean(rbind(tempClean[[i]][31:108], 
               tempClean[[i+1]][31:108]), na.rm = TRUE)
  treatmentTempStats[(i/2),5] <-
    sd(rbind(tempClean[[i]][31:108], 
             tempClean[[i+1]][31:108]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
#Rest: 112:291

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),6] <-
    mean(rbind(tempClean[[i]][112:291], 
               tempClean[[i+1]][112:291]), na.rm = TRUE)
  treatmentTempStats[(i/2),7] <-
    sd(rbind(tempClean[[i]][112:291], 
             tempClean[[i+1]][112:291]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
#Pulse 2: 301:395

for (i in seq(2, 8, 2)) {
  treatmentTempStats[(i/2),8] <-
    mean(rbind(tempClean[[i]][301:395], 
               tempClean[[i+1]][301:395]), na.rm = TRUE)
  treatmentTempStats[(i/2),9] <-
    sd(rbind(tempClean[[i]][301:395], 
             tempClean[[i+1]][301:395]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

## For each treatment

```{r}
treatmentTempStatsOverall <- data.frame("treatment" = c("control", "treatment"),
                                        "meanTempPre" = rep(0, times = 2),
                                        "sdTempPre" = rep(0, times = 2),
                                        "meanTempPulse1" = rep(0, times = 2),
                                        "sdTempPulse1" = rep(0, times = 2),
                                        "meanTempRest" = rep(0, times = 2),
                                        "sdTempRest" = rep(0, times = 2),
                                        "meanTempPulse2" = rep(0, times = 2),
                                        "sdTempPulse2" = rep(0, times = 2)) #Create an empty dataframe
treatmentTempStatsOverall #Confirm dataframe creation
```

```{r}
#Pre-exposure: 4:22

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),2] <-
    mean(rbind(tempClean[[i]][4:22], 
               tempClean[[i+1]][4:22],
               tempClean[[i+4]][4:22], 
               tempClean[[i+5]][4:22]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),3] <-
    sd(rbind(tempClean[[i]][4:22], 
             tempClean[[i+1]][4:22],
             tempClean[[i+4]][4:22], 
             tempClean[[i+5]][4:22]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
#Pulse 1: 31:108

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),4] <-
    mean(rbind(tempClean[[i]][31:108], 
               tempClean[[i+1]][31:108],
               tempClean[[i+4]][31:108], 
               tempClean[[i+5]][31:108]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),5] <-
    sd(rbind(tempClean[[i]][31:108], 
             tempClean[[i+1]][31:108],
             tempClean[[i+4]][31:108], 
             tempClean[[i+5]][31:108]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
#Rest: 112:291

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),6] <-
    mean(rbind(tempClean[[i]][112:291], 
               tempClean[[i+1]][112:291],
               tempClean[[i+4]][112:291], 
               tempClean[[i+5]][112:291]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),7] <-
    sd(rbind(tempClean[[i]][112:291], 
             tempClean[[i+1]][112:291],
             tempClean[[i+4]][112:291], 
             tempClean[[i+5]][112:291]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
#Pulse 2: 301:395

for (i in seq(2, 4, 2)) {
  treatmentTempStatsOverall[(i/2),8] <-
    mean(rbind(tempClean[[i]][301:395], 
               tempClean[[i+1]][301:395],
               tempClean[[i+4]][301:395], 
               tempClean[[i+5]][301:395]), na.rm = TRUE)
  treatmentTempStatsOverall[(i/2),9] <-
    sd(rbind(tempClean[[i]][301:395], 
             tempClean[[i+1]][301:395],
             tempClean[[i+4]][301:395], 
             tempClean[[i+5]][301:395]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStatsOverall
head(treatmentTempStatsOverall) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "overall-temp-mean-sd.csv", quote = FALSE)
```

# Statistical test

I will use a Kruskal-Wallis to confirm there are significant differences between treatments during the first pulse, and to confirm that there are no significant differences between treatments during the second pulse. I am using a Kruskal-Wallis test because there is missing data from some of the loggers due to failure, and ANOVAs cannot handle a bunch of missing values very well.

## Format data

I need to make an individual data frame for each pulse.

```{r}
tempPulse1Long <- tempClean[31:108,] %>%
  pivot_longer(., cols = temp10a:temp13b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp10a" | logger == "temp10b" ~ "tank10",
                             logger == "temp11a" | logger == "temp11b" ~ "tank11",
                             logger == "temp12a" | logger == "temp12b" ~ "tank12",
                             logger == "temp13a" | logger == "temp13b" ~ "tank13")) %>%
  mutate(., condition = case_when(tank == "tank10" | tank == "tank12" ~ "control",
                                  tank == "tank11" | tank == "tank13" ~ "treatment")) %>% 
  dplyr::select(-logger) %>% 
  drop_na() #Take pulse 1 data and pivot longer. Create tank and condition columns and remove logger column. Drop rows with NAs
head(tempPulse1Long)
```

```{r}
tempPulse2Long <- tempClean[301:406,] %>%
  pivot_longer(., cols = temp10a:temp13b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp10a" | logger == "temp10b" ~ "tank10",
                             logger == "temp11a" | logger == "temp11b" ~ "tank11",
                             logger == "temp12a" | logger == "temp12b" ~ "tank12",
                             logger == "temp13a" | logger == "temp13b" ~ "tank13")) %>%
  mutate(., condition = case_when(tank == "tank10" | tank == "tank12" ~ "control",
                                  tank == "tank11" | tank == "tank13" ~ "treatment")) %>% 
  dplyr::select(-logger) %>% 
  drop_na() #Take pulse 2 data and pivot longer. Create tank and condition columns and remove logger column. Drop rows with NAs
head(tempPulse2Long)
```

## Perform ANOVAs

```{r}
tempPulse1Stats <- kruskal.test(temperature ~ condition, data = tempPulse1Long)
tempPulse1Stats #Significant differences between treatments during the first pulse.
```

```{r}
tempPulse2Stats <- kruskal.test(temperature ~ condition, data = tempPulse2Long)
tempPulse2Stats #Significant differences between treatments during the second pulse
```

As expected, the conditions at Pulse 1 were different between treatments, but the test also identified significant differences during the second pulse. I need to investigate this further.

```{r}
tempPulse2TankStats <- kruskal.test(temperature ~ tank, data = tempPulse2Long) #Investigate differences by tank
tempPulse2TankStats #Significant differences between tanks during the second pulse
```

```{r}
tempPulse2TankPairwise <- pairwise.wilcox.test(tempPulse2Long$temperature, tempPulse2Long$tank,
                     p.adjust.method = "bonferroni") #Control tanks (10 and 12) are not significantly different from eachother. Tanks 11 and 13 (treatment) are significantly different from all other tanks. Tank 13 seems to have the largest difference between all tanks.
tempPulse2TankPairwise
```


```{r}
rbind(broom::tidy(tempPulse1Stats) %>%
        mutate(., pulse = 1),
      broom::tidy(tempPulse2Stats) %>%
        mutate(., pulse = 2),
      broom::tidy(tempPulse2TankStats) %>%
        mutate(., pulse = 2)) %>%
  write.csv("kruskal-wallis-test-output.csv", quote = FALSE, row.names = FALSE) #Save table with KW test output
```


```{r}
broom::tidy(tempPulse2TankPairwise) %>%
  write.csv("pairwise-wilcox-test-output-pulse2vtank.csv", quote = FALSE, row.names = FALSE) #Save pairwise test output
```

Tank 13 being different from all other tanks makes sense based on the tempreautre averages calculated above. Moving forward, I'll use tank as a random effect for the TTR analysis. I'll also need to plot all four tank temperatures during the second pulse.

# Plot full temperature conditions

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Reds")[7],
                brewer.pal(9, "Greys")[7]) #Create color scheme
```


add all 4 tank temperatures ± SE for second pulse as an inset

```{r}
tempClean[4:406,] %>%
  dplyr::select(dateTime, 
                temp10a, temp10b,
                temp11a, temp11b,
                temp12a, temp12b, 
                temp13a, temp13b) %>%
  rowwise() %>%
  mutate(., controlTemp = mean(c(temp10a, temp10b,
                                 temp12a, temp12b), na.rm = TRUE),
         treatmentTemp = mean(c(temp11a, temp11b,
                                temp13a, temp13b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = treatmentTemp)) +
  geom_line(color = plotColors[1]) + #Average treatment temperature
  geom_segment(aes(x = tempClean$dateTime[31], xend = tempClean$dateTime[108], 
                   y = treatmentTempStatsOverall$meanTempPulse1[2], yend = treatmentTempStatsOverall$meanTempPulse1[2]), 
               color = plotColors[1], linetype = 3) + #Pulse 1 treatment segment
  geom_ribbon(data = tempClean[31:108], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse1[2], 
                  ymin = treatmentTempStatsOverall$meanTempPulse1[2] - treatmentTempStatsOverall$sdTempPulse1[2], 
                  ymax = treatmentTempStatsOverall$meanTempPulse1[2] + treatmentTempStatsOverall$sdTempPulse1[2]), 
              fill = plotColors[1], alpha = 0.15) + #Pulse 1 treatment ribbon
  geom_segment(aes(x = tempClean$dateTime[301], xend = tempClean$dateTime[406], 
                   y = treatmentTempStatsOverall$meanTempPulse2[2], yend = treatmentTempStatsOverall$meanTempPulse2[2]), 
               color = plotColors[1], linetype = 3) + #Pulse 2 treatment segment
  geom_ribbon(data = tempClean[301:406], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse2[2], 
                  ymin = treatmentTempStatsOverall$meanTempPulse2[2] - treatmentTempStatsOverall$sdTempPulse2[2], 
                  ymax = treatmentTempStatsOverall$meanTempPulse2[2] + treatmentTempStatsOverall$sdTempPulse2[2]), 
              fill = plotColors[1], alpha = 0.15) + #Pulse 2 treatment ribbon
  geom_line(aes(x = dateTime, y = controlTemp), color = plotColors[2]) + #Control treatment temperature
  geom_segment(aes(x = tempClean$dateTime[31], xend = tempClean$dateTime[108], 
                   y = treatmentTempStatsOverall$meanTempPulse1[1], yend = treatmentTempStatsOverall$meanTempPulse1[1]), 
               color = plotColors[2], linetype = 3) + #Pulse 1 control segment
  geom_ribbon(data = tempClean[31:108], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse1[1], 
                  ymin = treatmentTempStatsOverall$meanTempPulse1[1] - treatmentTempStatsOverall$sdTempPulse1[1], 
                  ymax = treatmentTempStatsOverall$meanTempPulse1[1] + treatmentTempStatsOverall$sdTempPulse1[1]), 
              fill = plotColors[2], alpha = 0.15) + #Pulse 1 control ribbon
  geom_segment(aes(x = tempClean$dateTime[301], xend = tempClean$dateTime[406], 
                   y = treatmentTempStatsOverall$meanTempPulse2[1], yend = treatmentTempStatsOverall$meanTempPulse2[1]), 
               color = plotColors[2], linetype = 3) + #Pulse 2 control segment
  geom_ribbon(data = tempClean[301:406], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse2[1], 
                  ymin = treatmentTempStatsOverall$meanTempPulse2[1] - treatmentTempStatsOverall$sdTempPulse2[1], 
                  ymax = treatmentTempStatsOverall$meanTempPulse2[1] + treatmentTempStatsOverall$sdTempPulse2[1]), 
              fill = plotColors[2], alpha = 0.15) + #Pulse 2 treatment ribbon
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     limits = c(14.9,33),
                     breaks = c(seq(15,30,5), seq(31,33,1))) +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

```{r}
mainPlot <- tempClean[4:406,] %>%
  dplyr::select(dateTime, 
                temp10a, temp10b,
                temp11a, temp11b,
                temp12a, temp12b, 
                temp13a, temp13b) %>%
  rowwise() %>%
  mutate(., controlTemp = mean(c(temp10a, temp10b,
                                 temp12a, temp12b), na.rm = TRUE),
         treatmentTemp = mean(c(temp11a, temp11b,
                                temp13a, temp13b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = treatmentTemp)) +
  geom_line(color = plotColors[1]) + #Average treatment temperature
  geom_segment(aes(x = tempClean$dateTime[31], xend = tempClean$dateTime[108], 
                   y = treatmentTempStatsOverall$meanTempPulse1[2], yend = treatmentTempStatsOverall$meanTempPulse1[2]), 
               color = plotColors[1], linetype = 3) + #Pulse 1 treatment segment
  geom_ribbon(data = tempClean[31:108], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse1[2], 
                  ymin = treatmentTempStatsOverall$meanTempPulse1[2] - treatmentTempStatsOverall$sdTempPulse1[2], 
                  ymax = treatmentTempStatsOverall$meanTempPulse1[2] + treatmentTempStatsOverall$sdTempPulse1[2]), 
              fill = plotColors[1], alpha = 0.15) + #Pulse 1 treatment ribbon
  geom_segment(aes(x = tempClean$dateTime[301], xend = tempClean$dateTime[406], 
                   y = treatmentTempStatsOverall$meanTempPulse2[2], yend = treatmentTempStatsOverall$meanTempPulse2[2]), 
               color = plotColors[1], linetype = 3) + #Pulse 2 treatment segment
  geom_ribbon(data = tempClean[301:406], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse2[2], 
                  ymin = treatmentTempStatsOverall$meanTempPulse2[2] - treatmentTempStatsOverall$sdTempPulse2[2], 
                  ymax = treatmentTempStatsOverall$meanTempPulse2[2] + treatmentTempStatsOverall$sdTempPulse2[2]), 
              fill = plotColors[1], alpha = 0.15) + #Pulse 2 treatment ribbon
  geom_line(aes(x = dateTime, y = controlTemp), color = plotColors[2]) + #Control treatment temperature
  geom_segment(aes(x = tempClean$dateTime[31], xend = tempClean$dateTime[108], 
                   y = treatmentTempStatsOverall$meanTempPulse1[1], yend = treatmentTempStatsOverall$meanTempPulse1[1]), 
               color = plotColors[2], linetype = 3) + #Pulse 1 control segment
  geom_ribbon(data = tempClean[31:108], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse1[1], 
                  ymin = treatmentTempStatsOverall$meanTempPulse1[1] - treatmentTempStatsOverall$sdTempPulse1[1], 
                  ymax = treatmentTempStatsOverall$meanTempPulse1[1] + treatmentTempStatsOverall$sdTempPulse1[1]), 
              fill = plotColors[2], alpha = 0.15) + #Pulse 1 control ribbon
  geom_segment(aes(x = tempClean$dateTime[301], xend = tempClean$dateTime[406], 
                   y = treatmentTempStatsOverall$meanTempPulse2[1], yend = treatmentTempStatsOverall$meanTempPulse2[1]), 
               color = plotColors[2], linetype = 3) + #Pulse 2 control segment
  geom_ribbon(data = tempClean[301:406], 
              aes(x = dateTime, 
                  y = treatmentTempStatsOverall$meanTempPulse2[1], 
                  ymin = treatmentTempStatsOverall$meanTempPulse2[1] - treatmentTempStatsOverall$sdTempPulse2[1], 
                  ymax = treatmentTempStatsOverall$meanTempPulse2[1] + treatmentTempStatsOverall$sdTempPulse2[1]), 
              fill = plotColors[2], alpha = 0.15) + #Pulse 2 treatment ribbon
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     limits = c(14.9,33),
                     breaks = c(seq(15,30,5), seq(31,33,1))) +
  theme_classic(base_size = 15)
```

```{r}
tempClean[301:394,] %>%
  dplyr::select(dateTime, 
                temp10a, temp10b,
                temp11a, temp11b,
                temp12a, temp12b, 
                temp13a, temp13b) %>%
  rowwise() %>%
  mutate(., tank10avg = mean(c(temp10a, temp10b), na.rm = TRUE),
         tank11avg = mean(c(temp11a, temp11b), na.rm = TRUE),
         tank12avg = mean(c(temp12a, temp12b), na.rm = TRUE),
         tank13avg = mean(c(temp13a, temp13b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = tank10avg)) +
  geom_line(color = plotColors[2]) +
  geom_line(aes(x = dateTime, y = tank11avg), color = plotColors[1]) + 
  geom_line(aes(x = dateTime, y = tank12avg), color = plotColors[2]) +
  geom_line(aes(x = dateTime, y = tank13avg), color = plotColors[1]) #Average treatment temperature #Take temperature data for the second pulse and average data at each time point for each tank.
```








