---
title: "respo-analysis-kelso"
output: html_document
date: "2023-08-05"
---

This code was written by JC Kelso, and modified by YR Venkataraman. This script analyzes respirometry data from the repeat heat exposure experiment with WA crabs.

# Set up R Markdown document

```{bash}
#Make output directory
pwd
mkdir 04-respirometry
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/repeat-heat/04-respirometry/")) #Set root directory
```

```{r}
getwd()
```

# Install Packages

```{r packages, include=FALSE}
#if ("tidyverse" %in% rownames(installed.packages()) == 'FALSE') install.packages("tidyverse")
#if ("RColorBrewer" %in% rownames(installed.packages()) == 'FALSE') install.packages("RColorBrewer")
#if ("lubridate" %in% rownames(installed.packages()) == 'FALSE') install.packages("lubridate")
#if ("respirometry" %in% rownames(installed.packages()) == 'FALSE') install.packages("respirometry")
require(tidyverse)
require(RColorBrewer)
require(lubridate)
require(respirometry)
```

```{r}
sessionInfo()
```

# Import Data

## Genotype

```{r}
genotypeData <- read.csv("../../../data/repeat-heat/time-to-right-genotype-julia.csv", header = TRUE) %>%
  filter(., is.na(probe.number) == FALSE) %>%
  mutate(., treatment = case_when(tank == "10" | tank == "12" ~ "No_Pulse",
                                  tank == "11" | tank == "13" ~ "Pulse")) %>%
  mutate(., tank = case_when(tank == 10 ~ "tank10",
                             tank == 11 ~ "tank11", 
                             tank == 12 ~ "tank12",
                             tank == 13 ~ "tank13")) %>%
  mutate(., probe_num = as.character(probe.number)) %>%
  mutate(., presenceC = case_when(genotype == "CC" ~ "Y",
                                  genotype == "CT" ~ "Y",
                                  genotype == "TT" ~ "N")) %>%
  mutate(., presenceT = case_when(genotype == "CC" ~ "N",
                                  genotype == "CT" ~ "Y",
                                  genotype == "TT" ~ "Y")) %>%
  dplyr::select(date, treatment, tank, probe_num, crab.ID, genotype, presenceC, presenceT) #Import and format genotype data
head(genotypeData) #Confirm changes
```

## 2023-08-01

```{r}
files20230801 <- fs::dir_ls("../../../data/repeat-heat/respirometry/20230801/", glob = "*tank??_20230801.txt", recurse = TRUE) #Create a list of respirometry files from specified date. Use recurse = TRUE to prompt search through child directories.
files20230801 #Confirm the list has the correct files
```

```{r}
dat20230801 <- read_delim(files20230801, id = "tank", delim = "\t", skip = 20, col_names = FALSE) %>%
  select(., c(1:4, 7:9, 11)) #Import data using "tank" as an ID column for all files in the designated list. Specify tab-delimited files and skip the first 20 lines. There are no column names. Select data from all four probes. Use all date/time columns since some probes ran for longer than others.
colnames(dat20230801) <- c("path",
                           "date", "time", "dt_s", 
                           "perc_air_sat_2", "perc_air_sat_3", "perc_air_sat_4",
                           "temp_C", 
                           "pressure_mbar") #Rename columns
tail(dat20230801) #Confirm changes
```

```{r}
dat20230801 <- dat20230801 %>%
  mutate(., tank = str_sub(path, 49, 54)) %>%
  select(-path) %>%
  mutate(., treatment = case_when(tank == "tank10" ~ "No_Pulse",
                                  tank == "tank11" ~ "Pulse",
                                  tank == "tank12" ~ "No_Pulse",
                                  tank == "tank13" ~ "Pulse"))
tail(dat20230801)
```

## 2023-08-04

```{r}
files20230804 <- fs::dir_ls("../../../data/repeat-heat/respirometry/20230804/", glob = "*tank??_20230804.txt", recurse = TRUE)
files20230804
```

```{r}
dat20230804 <- read_delim(files20230804, id = "tank", delim = "\t", skip = 20, col_names = FALSE) %>%
  select(., c(1:4, 7:9, 11, 14))

colnames(dat20230804) <- c("path",
                           "date", "time", "dt_s", 
                           "perc_air_sat_2", "perc_air_sat_3", "perc_air_sat_4",
                           "temp_C", 
                           "pressure_mbar")
tail(dat20230804)
```

```{r}
dat20230804 <- dat20230804 %>%
  mutate(., tank = str_sub(path, 49, 54)) %>%
  select(-path) %>%
  mutate(., treatment = case_when(tank == "tank10" ~ "No_Pulse",
                                  tank == "tank11" ~ "Pulse",
                                  tank == "tank12" ~ "No_Pulse",
                                  tank == "tank13" ~ "Pulse")) %>%
  mutate(., perc_air_sat_2 = as.numeric(perc_air_sat_2)) %>%
  mutate(., perc_air_sat_4 = as.numeric(perc_air_sat_4))
tail(dat20230804)
```

# Format Data

## 2023-08-01

```{r}
dat20230801mod <- dat20230801 %>%
  mutate(., datetime = dmy_hms(paste(date, time))) %>%
  mutate(., salinity = case_when(tank == "tank10" ~ 33,
                                 tank == "tank11" ~ 33,
                                 tank == "tank12" ~ 30,
                                 tank == "tank13" ~ 30,)) %>%
  group_by(., treatment, tank) %>%
  arrange(., .by_group = TRUE) %>%
  ungroup(.) %>%
  pivot_longer(., cols = c(perc_air_sat_2, perc_air_sat_3, perc_air_sat_4), names_to = "og_perc", values_to = "perc_air_sat") %>%
  mutate(., og_perc = gsub("perc_air_sat_", replacement = "", x = og_perc)) %>%
  dplyr::rename(., probe_num = og_perc)
head(dat20230801mod)
```

## 2023-08-04

```{r}
dat20230804mod <- dat20230804 %>%
  mutate(., datetime = dmy_hms(paste(date, time))) %>%
  mutate(., salinity = case_when(tank == "tank10" ~ 33,
                                 tank == "tank11" ~ 33,
                                 tank == "tank12" ~ 32,
                                 tank == "tank13" ~ 32,)) %>%
  group_by(., treatment, tank) %>%
  arrange(., .by_group = TRUE) %>%
  ungroup(.) %>%
  pivot_longer(., cols = c(perc_air_sat_2, perc_air_sat_3, perc_air_sat_4), names_to = "og_perc", values_to = "perc_air_sat") %>%
  mutate(., og_perc = gsub("perc_air_sat_", replacement = "", x = og_perc)) %>%
  dplyr::rename(., probe_num = og_perc)
head(dat20230804mod)
```

# Clean data based on figures

Moving forward, I am going to focus on the final data point. Looking at day 4 data will still allow me to understand if priming had an impact on respiration rate.

## 2023-08-04

```{r}
dat20230804clean <- dat20230804mod %>%
  filter(., perc_air_sat <= 100) %>% 
  filter(., !(tank == "tank11" & probe_num == 3 & dt_s > 120 & dt_s < 360)) %>%
  filter(., !(tank == "tank11" & probe_num == 2 & dt_s < 180)) %>%
  filter(., !(tank == "tank12" & probe_num == 2 & dt_s < 120)) %>%
  filter(., !(tank == "tank10" & probe_num == 3 & dt_s < 100)) %>%
  mutate(., dt_s = ifelse(tank == "tank11" & probe_num == 3, dt_s - 120, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank11" & probe_num == 2, dt_s - 180, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank12" & probe_num == 2, dt_s - 120, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank10" & probe_num == 3, dt_s - 100, dt_s)) %>%
  group_by(tank, probe_num) %>%
  filter(., cumsum(perc_air_sat <= 75) == 0) %>%
  ungroup(.) %>%
  ggplot(., mapping = aes(x = (dt_s/60), y = perc_air_sat, color = treatment)) +
  geom_point(size = 0.5) +
  geom_smooth(method = lm, se = FALSE, lty = 1, color = "grey50")+
  facet_grid(cols = vars(treatment, tank), rows = vars(probe_num), scales = "free_x") +
  xlab("Time (min)") + ylab("Oxygen Air Saturation (%)") +
  scale_y_continuous(breaks = seq(70, 100, 10), 
                     limits = c(70, 100)) +
  scale_x_continuous(breaks = seq(0,10,5),
                     limits = c(0,10)) +
  scale_color_manual(values = c("#eb8d88", "#9699eb"), 
                     name = "Treatment",
                     breaks = c("Pulse","No_Pulse"),
                     labels = c("Pulse","No_Pulse")) +
  theme_classic(base_size = 15) + theme(strip.text = element_blank(),
                                        axis.line = element_line())
dat20230804clean
```

```{r}
##raw data
dat20230804mod %>%
  filter(., perc_air_sat <= 100) %>% 
  filter(., tank == "tank12" & probe_num == 4) %>%
  #group_by(tank, probe_num) %>%
  #filter(., cumsum(perc_air_sat <= 80) == 0) %>%
  ungroup(.) %>%
  ggplot(., mapping = aes(x = (dt_s), y = perc_air_sat, color = treatment)) +
  geom_point(size = 0.5) +
  geom_smooth(method = lm, se = FALSE, lty = 1, color = "grey50")+
  facet_grid(cols = vars(treatment, tank), rows = vars(probe_num), scales = "free_x") +
  xlab("Time (min)") + ylab("Oxygen Air Saturation (%)") +
  scale_y_continuous(breaks = seq(71, 100, 10), 
                     limits = c(71, 100)) +
  scale_x_continuous(breaks = seq(200,340,50),
                     limits = c(200,340)) +
  scale_color_manual(values = c("#eb8d88", "#9699eb"), 
                     name = "Treatment",
                     breaks = c("Pulse","No_Pulse"),
                     labels = c("Pulse","No_Pulse")) +
  theme_classic(base_size = 15) + theme(strip.text = element_blank(),
                                        axis.line = element_line())
```

```{r}
dat20230804mod <- dat20230804mod %>%
  mutate(., DO_umol_L = conv_o2(o2 = perc_air_sat, from = "percent_a.s.",
                                to = "umol_per_l",
                                temp = temp_C,
                                sal = salinity,
                                atm_pres = pressure_mbar)) #Create a new DO variable in units umol/L from %saturation using the function conv_o2() in the respirometry package
head(dat20230804mod)
```

```{r}
dat20230804cleanDOumol <- dat20230804mod %>%
  filter(., perc_air_sat <= 100) %>% 
  filter(., !(tank == "tank11" & probe_num == 3 & dt_s > 120 & dt_s < 360)) %>%
  filter(., !(tank == "tank11" & probe_num == 2 & dt_s < 180)) %>%
  filter(., !(tank == "tank12" & probe_num == 2 & dt_s < 120)) %>%
  filter(., !(tank == "tank10" & probe_num == 3 & dt_s < 100)) %>%
  mutate(., dt_s = ifelse(tank == "tank11" & probe_num == 3, dt_s - 120, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank11" & probe_num == 2, dt_s - 180, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank12" & probe_num == 2, dt_s - 120, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank10" & probe_num == 3, dt_s - 100, dt_s)) %>%
  group_by(tank, probe_num) %>%
  filter(., cumsum(perc_air_sat <= 75) == 0) %>%
  ungroup(.) %>%
  ggplot(., mapping = aes(x = (dt_s/60), y = DO_umol_L, color = treatment)) +
  geom_point(size = 0.5) +
  geom_smooth(method = lm, se = FALSE, lty = 1, color = "grey50") +
  facet_grid(cols = vars(treatment, tank), rows = vars(probe_num), scale = "free_x") +
  xlab("Time (min)") + ylab("Dissolved Oxygen (µmol)") +
  scale_color_manual(values = c("#eb8d88", "#9699eb"), 
                     name = "Treatment",
                     breaks = c("Pulse","No_Pulse"),
                     labels = c("Pulse","No_Pulse")) +
  theme_classic(base_size = 15) + theme(strip.text = element_blank(),
                                        axis.line = element_line())
dat20230804cleanDOumol
```

# Extract slopes

### 2023-08-04

```{r}
dat20230804data <- dat20230804mod %>%
  filter(., perc_air_sat <= 100) %>% 
  filter(., !(tank == "tank11" & probe_num == 3 & dt_s > 120 & dt_s < 360)) %>%
  filter(., !(tank == "tank11" & probe_num == 2 & dt_s < 180)) %>%
  filter(., !(tank == "tank12" & probe_num == 2 & dt_s < 120)) %>%
  filter(., !(tank == "tank10" & probe_num == 3 & dt_s < 100)) %>%
  mutate(., dt_s = ifelse(tank == "tank11" & probe_num == 3, dt_s - 120, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank11" & probe_num == 2, dt_s - 180, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank12" & probe_num == 2, dt_s - 120, dt_s)) %>%
  mutate(., dt_s = ifelse(tank == "tank10" & probe_num == 3, dt_s - 100, dt_s)) %>%
  group_by(tank, probe_num) %>%
  filter(., cumsum(perc_air_sat <= 75) == 0) %>%
  ungroup(.)

head(dat20230804data)  
```

```{r}
dat20230804_MO2_slopes <- dat20230804data %>%
  nest(data = -c(treatment, tank, probe_num)) %>%
  mutate(fit = map(data, ~lm(DO_umol_L ~ dt_s, data = .))) %>%
  mutate(tidied = map(fit, tidy)) %>%
  unnest(tidied)
dat20230804_MO2_slopes
```

```{r}
dat20230804_MO2_glance <- dat20230804data %>%
  nest(data = -c(treatment, tank, probe_num)) %>%
  mutate(fit = map(data, ~lm(DO_umol_L ~ dt_s, data = .))) %>%
  mutate(glanced = map(fit, glance)) %>%
  unnest(glanced)
dat20230804_MO2_glance
```

```{r}
dat20230804_MO2_slope_results <- dat20230804_MO2_slopes %>%
  filter(term == "dt_s") %>%
  mutate(slope_nmol_hr = estimate*1000*60) %>%
  dplyr::select(treatment, tank, probe_num, slope_nmol_hr, p.value) %>%
  mutate(r.squared = (dat20230804_MO2_glance$r.squared))
head(dat20230804_MO2_slope_results)
```

# Blank-correct slopes

Next step: extract data from the last 10-20 minutes of each chamber. This is the background respiration.

### 2023-08-04

```{r}
dat20230804mod %>%
  ggplot(., mapping = aes(x = (dt_s), y = perc_air_sat, color = treatment)) +
  geom_point(size = 0.5) +
  geom_line(y = 80, lty = 1, color = "grey50") +
  facet_grid(cols = vars(treatment, tank), rows = vars(probe_num), scales = "free_x") +
  xlab("Time (min)") + ylab("Oxygen Air Saturation (%)") +
  scale_x_continuous(limits = c(1200,1800)) + #Identify the last 10 minutes of each trial to use for the background.
  scale_color_manual(values = c("#eb8d88", "#9699eb"), 
                     name = "Treatment",
                     breaks = c("Pulse","No_Pulse"),
                     labels = c("Pulse","No_Pulse")) +
  theme_classic(base_size = 15) + theme(strip.text = element_blank(),
                                        axis.line = element_line()) #Plot raw air saturation values and add a line for 80% saturation. Modify x and y axis labels, y axis breaks. Facet such that each column is a treatment/tank, and each row is a probe. This provides two columns per treatment.
```

```{r}
dat20230804mod %>%
  filter(., dt_s >= 1200 & dt_s <= 1800) %>%
  mutate(., dt_s = dt_s - 1200) %>%
  ggplot(., mapping = aes(x = (dt_s/60), y = perc_air_sat, color = treatment)) +
  geom_point(size = 0.5) +
  geom_smooth(method = lm, se = FALSE, lty = 1, color = "grey50")+
  facet_grid(cols = vars(treatment, tank), rows = vars(probe_num), scale = "free_x") +
  xlab("Time (min)") + ylab("Oxygen Air Saturation (%)") +
  scale_color_manual(values = c("#eb8d88", "#9699eb"), 
                     name = "Treatment",
                     breaks = c("Pulse","No_Pulse"),
                     labels = c("Pulse","No_Pulse")) +
  theme_classic(base_size = 15) + theme(strip.text = element_blank(),
                                        axis.line = element_line()) #Identify and plot ~10 min of background respiration data that will be used to calculate the background respiration slope
```

```{r}
dat20230804blanks <- dat20230804mod %>%
  filter(., !(tank == "tank13" & probe_num == 4)) %>% #Remove empty chamber information
  filter(., !(tank == "tank11" & probe_num == 4)) %>% #Remove empty chamber information
  filter(., dt_s >= 1200 & dt_s <= 1800) %>%
  mutate(., dt_s = dt_s - 1200)
head(dat20230804blanks) #Confirm dataframe creation
```

```{r}
dat20230804_MO2_slopes_blanks <- dat20230804blanks %>%
  nest(data = -c(treatment, tank, probe_num)) %>%
  mutate(fit = map(data, ~lm(DO_umol_L ~ dt_s, data = .))) %>%
  mutate(tidied = map(fit, tidy)) %>%
  unnest(tidied) #Create a list object that stores the regression results for each chamber. Nest data by treatment, tank, and probe number so the regression slopes are calculated appropriately. Use map to extract linear model fit information for DO_umol_L by time. Extract slopes using broom::tidy and unnest.
dat20230804_MO2_slopes_blanks #Confirm dataframe creation
```

```{r}
dat20230804_MO2_glance_blanks <- dat20230804blanks %>%
  nest(data = -c(treatment, tank, probe_num)) %>%
  mutate(fit = map(data, ~lm(DO_umol_L ~ dt_s, data = .))) %>%
  mutate(glanced = map(fit, glance)) %>%
  unnest(glanced) #Similar to above, but calculate the adjusted R-squared values by nesting treatment and probe information
dat20230804_MO2_glance_blanks #Confirm dataframe creation
```

```{r}
dat20230804_MO2_slopes_blanks_results <- dat20230804_MO2_slopes_blanks %>%
  filter(term == "dt_s") %>%
  mutate(slope_nmol_hr = estimate*1000*60) %>%
  dplyr::select(treatment, tank, probe_num, slope_nmol_hr, p.value) %>%
  mutate(r.squared = (dat20230804_MO2_glance_blanks$r.squared)) #Get the regression information for dt_s and not the intercept. Creates new a variable for respiration rate as nmol_per_hour (slope_nmol_hr). Select treatment, tank, probe number, slope over the hour, and p-value. Add the r-squared information from glance.
dat20230804_MO2_slopes_blanks_results #Confirm dataframe creation
```

```{r}
dat20230804_MO2_all_slope_results <- dat20230804_MO2_slope_results %>%
  left_join(., dat20230804_MO2_slopes_blanks_results, by = c("treatment", "tank", "probe_num"), suffix = c("", "_blank")) %>%
  mutate(., slope_nmol_hr_corrected = slope_nmol_hr - dat20230804_MO2_slopes_blanks_results$slope_nmol_hr) %>%
  mutate(., blank_ratio = abs(slope_nmol_hr_blank / slope_nmol_hr)) #Join dataframes by treatment, tank, and probe_num. Create a new column for the the corrected slope (uncorrected slope - blank slope) and the blank ratio (blank / uncorrected slope)
dat20230804_MO2_all_slope_results
```

```{r}
write.csv(dat20230804_MO2_all_slope_results, "20230804-MO2-slope-results.csv", quote = FALSE, row.names = FALSE) #Save results with uncorrected, blank, and corrected slopes
```

# Compare blank-corrected oxygen consumption by treatment

Given the small sample size, I will not run a statistical test for genotype.

```{r}
hist(dat20230804_MO2_all_slope_results$slope_nmol_hr_corrected * -1)
hist(x = log(dat20230804_MO2_all_slope_results$slope_nmol_hr_corrected * -1)) #Log transformation helps normalize data
hist(x = sqrt(dat20230804_MO2_all_slope_results$slope_nmol_hr_corrected * -1)) #sqrt transformation does nothing
```

```{r}
lm20230804_treatment <- lm((slope_nmol_hr_corrected * -1) ~ as.factor(treatment),
                           data = dat20230804_MO2_all_slope_results) #Mixed effects model examining treatment imapct on oxygen consumption. Include tank as a random effect due to tank temperature differences
summary(lm20230804_treatment)
```

No significant impact of treatment on oxygen consumption (t = -1.679, p = 0.131728).

```{r}
hist(residuals(lm20230804_treatment)) #Residuals normally distributed

plot(fitted(lm20230804_treatment), residuals(lm20230804_treatment)) #Lack of dispersion across x-axis is likely related to discrete variables
abline(h = 0, lty = 2, col = "grey")
```

# Plot

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Reds")[7],
                brewer.pal(9, "Greys")[7]) #Create color scheme
```

```{r}
dat20230804_MO2_all_slope_results
```

```{r}
dat20230804_MO2_all_slope_results %>%
  left_join(x = ., y = genotypeData %>%
              filter(., date == "8/4/23") %>%
              dplyr::select(treatment, tank, probe_num, genotype, presenceC, presenceT)) %>%
  ggplot(., aes(x = treatment, y = -slope_nmol_hr_corrected, color = treatment)) + 
  geom_boxplot() + geom_point(aes(shape = genotype), size = 2) +
  xlab("") + ylab("Oxygen Consumption (nmol/hr)") +
  scale_x_discrete(labels = c("Control", "Treatment")) +
  scale_color_manual(guide = "none",
                     values = c(plotColors[2], plotColors[1]),
                     labels = c("Control", "Treatment")) +
  scale_shape_manual(name = "Genotype",
                     values = c(15, 17, 19),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15)
ggsave("figures/2023-08-04-oxygen-consumption-nmolhr-corrected.pdf", width = 11, height = 8.5) #Create a boxplot for oxygen consumption (µmol/hr). Use -slope_nmol_hr so the larger the slope, the faster the oxygen consumption over the hour. 
```
