---
title: "04-TTR-analysis"
author: "Yaamini Venkataraman"
date: '2024-07-12'
output: html_document
---

In this document I'll examine how time-to-right (TTR) varies by temperature and time.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../output/03-TTR-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, include=FALSE}
#install.packages("tidyverse")
#install.packages("plotrix")
#install.packages("RColorBrewer")
#install.packages("lme4")
#install.packages("ggpubr")
#install.packages("patchwork")
#install.packages("respirometry")
require(tidyverse)
require(plotrix)
require(RColorBrewer)
require(lme4)
require(ggpubr)
require(patchwork)
require(respirometry)
```

```{r}
sessionInfo()
```

# Import data

```{r}
rawTTR <- read.csv("../../data/MA/time-to-right-MA-clean.csv", header = TRUE, na.strings = c("N/A", "NA")) #Import raw data
head(rawTTR) #Confirm import. Trial data is in seconds.
```

```{r}
crabMetadata <- read.csv("../../data/MA/crab-metadata-MA.csv", header = TRUE) %>%
  dplyr::select(., c("number", "tank", "carapace.width", "carapace.length")) %>%
  dplyr::rename(crab.ID = "number") #Import metadata. Select ID, treatment, width and length columns. Rename "ID" as "crab.ID"
head(crabMetadata) #Confirm changes
```

```{r}
# genotypeData <- read.csv("../../data/genotype/2023-crab-genotype.csv", header = TRUE, na.strings = "") %>%
#   select(., crab.ID, treatment, final.genotype) #Import genotype data. Remove rows without final genotypes
# head(genotypeData) #Confirm genotype dataframe creation
```

# Format data

```{r}
modTTR <- rawTTR %>%
  dplyr::select(., -notes) %>%
  left_join(., crabMetadata, by = c("crab.ID", "tank", "carapace.width")) %>%
  filter(., is.na(trial.1) == FALSE) %>%
  mutate(., integument.cont = case_when(integument.color == "B/G" ~ 0.5,
                                        integument.color == "G" ~ 1.0,
                                        integument.color == "Y/G" ~ 1.5,
                                        integument.color == "Y" ~ 2.0,
                                        integument.color == "Y/O" ~ 2.5,
                                        integument.color == "O" ~ 3,
                                        integument.color == "R/O" ~ 3.5,
                                        integument.color == "R" ~ 4)) %>%
  mutate(., treatment = case_when(tank == "T01" ~ "0C",
                                  tank == "T02" ~ "0C",
                                  tank == "T03" ~ "0C",
                                  tank == "T05" ~ "5C",
                                  tank == "T06" ~ "5C",
                                  tank == "T07" ~ "5C")) %>%
  mutate(., missing.swimmer = case_when(missing.swimmer == "Y" ~ missing.swimmer,
                                        missing.swimmer == "N" ~ missing.swimmer,
                                        missing.swimmer == "Y " ~ "Y")) %>%
  mutate(., trial.1 = na_if(trial.1, 91.00)) %>%
  mutate(., trial.2 = na_if(trial.2, 91.00)) %>%
  mutate(., trial.3 = na_if(trial.3, 91.00)) %>%
  rowwise(.) %>% 
  mutate(., TTRavg = mean(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  # filter(., is.na(TTRavg) == FALSE) %>%
  mutate(., TTRSE = std.error(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  # filter(., is.na(TTRSE) == FALSE) %>%
  ungroup(.) #Remove sampling ID and notes columns. Use case_when to create a column of continuous integument color metrics. Change all 91 to NA, calculate average and SE using data from the three TTR trials using rowwise operations. Ungroup. Join with genotype data.
head(modTTR) #Confirm formatting
```

# Data exploration

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Blues")[9],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

## Treatment

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = treatment)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  scale_color_manual(values = c(plotColors[1], plotColors[2]), 
                     name = "Temperature (ºC)",
                     breaks = c("0C", "5C"),
                     labels = c("0", "5")) +
  theme_classic(base_size = 15) #Plot average TTR and add lm fits without SE bars. Modify x-axis to only show experimental timepoints where TTR was measured. Assign colors to each treatment. Increase base font size.
ggsave("figures/time-to-right-raw.pdf", height = 8.5, width = 11)
```

I'm now going to make plots for various other factors to see if any of these need to be taken into consideration!

## Sex

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = sex)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-sex.pdf", height = 8.5, width = 11)
```

## Tank

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = tank)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-tank.pdf", height = 8.5, width = 11)
```

## Integument color

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = integument.color)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  scale_color_manual(values = c("#66C2A5","#A6D96A","#D9EF8B","#FFFF99","#FDAE61","#F46D43","#D53E4F"),
                     name = "Integument",
                     breaks = c("B/G", "G", "Y/G", "Y", "Y/O", "O", "R/O"),
                     labels = c("B/G", "G", "Y/G", "Y", "Y/O", "O", "R/O")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-integument.pdf", height = 8.5, width = 11)
```

## Missing swimmer

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = as.factor(missing.swimmer))) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-missswim.pdf", height = 8.5, width = 11)
```

## Weight

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = weight)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-weight.pdf", height = 8.5, width = 11)
```

## Carapace width

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = carapace.width)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-carapaceWidth.pdf", height = 8.5, width = 11)
```

## Genotype

```{r}
modTTR %>%
  filter(., is.na(final.genotype) == FALSE) %>% 
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = final.genotype)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 65, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-genotype.pdf", height = 8.5, width = 11)
```


# Assess differences by treatment

I will use a linear mixed effects model with individual as a random effect to analyze the data, following [this tutorial](https://jontalle.web.engr.illinois.edu/MISC/lme4/bw_LME_tutorial.pdf).

```{r}
hist(modTTR$TTRavg)
hist(x = log(modTTR$TTRavg)) #Log transformation helps normalize data
```

```{r}
TTRmodel <- lmer(log(TTRavg) ~ treatment + timepoint + treatment:timepoint + 
                   as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                   (1|crab.ID), 
                 data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run model with treatment, timepoint, their interaction, sex, integument color, carapace width, weight, and missing swimmer information as fixed effects, and individual as a random effect. Change the likelihood estimator (REML = FALSE) to compare models with LRT later. Use data without rows with NAs
summary(TTRmodel) #Look at model output.
#Residuals explain more random effect variation than crab ID, crab ID explains very little variation
```
Now that I've run the full model, I want to test the importance of treatment, timepoint, and their interaction as predictors. I'll include all other factors in the null model to account for any variation they may contribute.

```{r}
TTRnullTreatment <- lmer(log(TTRavg) ~ timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run null model without treatment to identify significance of predictor
anova(TTRnullTreatment, TTRmodel) #Treatment is a significant predictor
```

```{r}
TTRnullTimepoint <- lmer(log(TTRavg) ~ treatment + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run null model without timepoint to identify significance of predictor
anova(TTRnullTreatment, TTRmodel) #Timepoint is a significant predictor
```

```{r}
TTRnullInteraction <- lmer(log(TTRavg) ~ treatment + timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run model with interaction to identify significance of predictor. The full model constructed does not include an interaction
anova(TTRnullInteraction, TTRmodel) #Timepoint is a significant predictor
```

# Threshold analysis

In our data, trials for crabs that didn't right within 90 seconds are noted as NAs. I can examine these data in two ways:

1. How many crabs didn't right within 90 seconds in the first, second, or third trial?
2. Did a crab not right within 90 seconds in any trial?

The first would be easily answered in a plot. For the second, I can use a binomial model similar to [Coyle et al. (2019)](https://doi.org/10.1242/jeb.203521).

## Define threshold

```{r}
modTTR <- modTTR %>%
  mutate(., TTRthresh = case_when(!(is.na(trial.1 & trial.2 & trial.3) == FALSE) ~ 0,
                                  is.na(trial.1 & trial.2 & trial.3) == FALSE ~ 1)) #Create a new column where 1 = righted within 90 seconds in all trials and 0 = failure to right within that threshold
head(modTTR) #Confirm modifications
```

## Binomial model

```{r}
TTRthreshModel <- glmer(I(TTRthresh == 1) ~ treatment + timepoint + treatment:timepoint +
                          as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                          (1|crab.ID), 
                        family = binomial(), 
                        data = modTTR) #Create a binomial model, where 1 = success of righting wihtin the defined threshold
summary(TTRthreshModel) #Only weight is significant
```

# Final plots

## Average TTR plot

```{r}
modTTR %>%
  dplyr::select(., c(timepoint, treatment, TTRavg)) %>%
  filter(., is.na(TTRavg) == FALSE) %>%
  distinct(.) %>%
  ggplot(mapping = aes(x = as.character(timepoint), y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_dodge(width = 0.75)) +
  ylab("Average Time-to-Right (s)") +
  scale_x_discrete(name = "Exposure Time (Hours)",
                   breaks = unique(modTTR$timepoint),
                   labels = c(0, 4, 22, 28, 46, 94)) +
  scale_color_manual(values = c(plotColors[1], plotColors[2]),
                     name = "Temperature (ºC)",
                     breaks = c("0C", "5C"),
                     labels = c("0", "5")) +
  scale_shape_manual(values = c(15, 19),
                     name = "Temperature (ºC)",
                     breaks = c("0C", "5C"),
                     labels = c("0", "5")) +
  theme_classic(base_size = 15) #Select unique timepoint, treatment, and TTRavg data, and remove rows where TTRavg = NA. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Assign colors and shapes to each treatment. Increase base font size.
ggsave("figures/time-to-right-avg-boxplot.pdf", height = 8.5, width = 11)
```

## Threshold analysis plot

# Individual-level analysis

## Crabs tracked throught the experiment

I have 27 individuals (three per tank) with TTR and respirometry data! I'm going to examine that subset to understand individual variation within each treatment, and how that may or may not change over time.

```{r}
modTTRSubsetR <- modTTR %>%
  filter(., respirometry == "Y") %>%
  unique(.) #Subset data for crabs used in respirometry and TTR
head(modTTRSubsetR)
nrow(modTTRSubsetR)
```

First things first: I want to understand if there are certain individuals that vary more over time! The simplest way to do this is to just plot the data.

```{r}
modTTRSubsetR %>%
  dplyr::select(., c(crab.ID, timepoint, tank, treatment, TTRavg)) %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment == "15C" ~ "15C",
                                  treatment == "25C" ~ "25C")) %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, group = crab.ID, color = treatment, shape = treatment)) +
  geom_point(size = 1, alpha = 0.75) + geom_line(linetype = 3) +
  facet_wrap(vars(treatment, tank), ncol = 3) +
  scale_x_continuous(name = "timepoint",
                     breaks = c(seq(0, 42, 14), 49),
                     limits = c(0, 49)) +
  scale_y_continuous(name = "Individual Time-to-Right (s)",
                     breaks = c(0, seq(1, 5, 2), seq(5, 15, 5))) +
  # scale_y_continuous(name = "Average Time-to-Right (s)") +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(strip.text.x = element_blank(),
                                        strip.background = element_blank()) #Plot TTR for each individual crab that was tracked over the entire experiment. Facet by tank and treatment so it's easier to visualize. Assign colors to each treatment.
ggsave("figures/time-to-right-avg-individuals.pdf", height = 8.5, width = 11)
```

Alright, this plot tells us what we already know: that there's more variation in the 5ºC than the other two.

I also want to look at difference in TTR over time by looking at the difference in TTR from the average...

```{r}
modTTRSubsetR %>%
  dplyr::select(., c(crab.ID, timepoint, tank, treatment, TTRavg, TTRavgFull)) %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment == "15C" ~ "15C",
                                  treatment == "25C" ~ "25C")) %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg - TTRavgFull, group = crab.ID, color = treatment, shape = treatment)) +
  geom_point(size = 1, alpha = 0.75) + geom_line(linetype = 3) +
  facet_wrap(vars(treatment, tank), ncol = 3) +
  scale_x_continuous(name = "timepoint",
                     breaks = c(seq(0, 42, 14), 49),
                     limits = c(0, 49)) +
  scale_y_continuous(name = "Difference in Time-to-Right (s)",
                     breaks = seq(-5, 10, 5)) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(strip.text.x = element_blank(),
                                        strip.background = element_blank()) #Plot difference in individual TTR from average TTR for each individual crab that was tracked over the entire experiment. Facet by tank and treatment so it's easier to visualize. Assign colors to each treatment.
ggsave("figures/time-to-right-avg-individuals-diff.pdf", height = 8.5, width = 11)
```

```{r}
modTTRSubsetR %>%
  dplyr::select(., c(crab.ID, timepoint, tank, treatment, TTRavg, TTRavgFull, final.genotype)) %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment == "15C" ~ "15C",
                                  treatment == "25C" ~ "25C")) %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg - TTRavgFull, group = crab.ID, color = treatment, shape = final.genotype)) +
  geom_point(size = 1, alpha = 0.75) + geom_line(linetype = 3) +
  facet_wrap(vars(treatment, tank), ncol = 3) +
  scale_x_continuous(name = "timepoint",
                     breaks = c(seq(0, 42, 14), 49),
                     limits = c(0, 49)) +
  scale_y_continuous(name = "Difference in Time-to-Right (s)",
                     breaks = seq(-5, 10, 5)) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(strip.text.x = element_blank(),
                                        strip.background = element_blank()) #Plot difference in individual TTR from average TTR for each individual crab that was tracked over the entire experiment. Facet by tank and treatment so it's easier to visualize. Assign colors to each treatment.
ggsave("figures/time-to-right-avg-individuals-diff-genotype.pdf", height = 8.5, width = 11)
```

...and percent change from the average.

```{r}
modTTRSubsetR %>%
  dplyr::select(., c(crab.ID, timepoint, tank, treatment, TTRavg, TTRavgFull)) %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment == "15C" ~ "15C",
                                  treatment == "25C" ~ "25C")) %>%
  ggplot(., mapping = aes(x = timepoint, y = ((TTRavg - TTRavgFull)/TTRavgFull)*100, group = crab.ID, color = treatment, shape = treatment)) +
  geom_point(size = 1, alpha = 0.75) + geom_line(linetype = 3) +
  facet_wrap(vars(treatment, tank), ncol = 3, scale = "free_y") +
  scale_x_continuous(name = "timepoint",
                     breaks = c(seq(0, 42, 14), 49),
                     limits = c(0, 49)) +
  scale_y_continuous(name = "Percent Change in Time-to-Right") +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]), 
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(strip.text.x = element_blank(),
                                        strip.background = element_blank()) #Plot difference in individual TTR from average TTR for each individual crab that was tracked over the entire experiment. Facet by tank and treatment so it's easier to visualize. Assign colors to each treatment.
ggsave("figures/time-to-right-avg-individuals-percChange.pdf", height = 8.5, width = 11)
```

## Genotype analysis

I'm going to examine the effect of genotype on TTR in a few ways. 

### Average TTR

I'm going to start by looking the influnce of genotype variables on overall TTR:

1. Genotype
2. Presence of "C" (warm-adapted allele) (the inverse )

```{r}
modTTRGenotype <- modTTR %>%
  mutate(., presenceC = case_when(final.genotype == "CC" ~ "Y",
                                  final.genotype == "CT" ~ "Y",
                                  final.genotype == "TT" ~ "N")) %>%
  mutate(., presenceT = case_when(final.genotype == "CC" ~ "N",
                                  final.genotype == "CT" ~ "Y",
                                  final.genotype == "TT" ~ "Y")) #Subset data with genotype information. Add two columns for presence of C or T allele
head(modTTRGenotype)
```

```{r}
TTRmodelSubset <- glm(log(TTRavg) ~ treatment + timepoint + treatment*timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + as.factor(final.genotype), data = modTTRGenotype) #ANOVA, with log(TTRavg) as the response variable, and treatment, timepoint, the interaction between treatment and timepoint, sex, integument color (continuous), carapace width, weight, whether or not a crab is missing legs, and genotype as explanatory variables. Cannot add presence of C or T allele into the model since they are colinear with genotype
summary(TTRmodelSubset) #Various factors significant
```

```{r}
step(TTRmodelSubset, direction = "backward")
```

```{r}
TTRmodelSubset2 <- glm(log(TTRavg) ~ treatment + timepoint + treatment*timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + as.factor(presenceC) + as.factor(presenceT), data = modTTRGenotype) #ANOVA, with log(TTRavg) as the response variable, and treatment, timepoint, the interaction between treatment and timepoint, sex, integument color (continuous), carapace width, weight, whether or not a crab is missing legs, and presence of C or T allele
summary(TTRmodelSubset2) #Various factors significant
```

```{r}
step(TTRmodelSubset2, direction = "backward")
```

No genotype variable has a significant influence on average TTR. Can we see if any of our outliers in our average TTR boxplots are associated with a particular genotype?

```{r}
facet_labels <- c("05C" = "5ºC",
                  "15C" = "15ºC",
                  "25C" = "25ºC") #Assign labels for facets
```

```{r}
modTTRGenotype %>%
  dplyr::select(., c(timepoint, treatment, TTRavg, final.genotype)) %>%
  distinct(.) %>%
  filter(., treatment != "15C") %>%
  mutate(., timepoint = case_when(timepoint == 1 ~ "01",
                                  timepoint == 2 ~ "02",
                                  timepoint == 7 ~ "07",
                                  timepoint != 1 | 2 | 7 ~ as.character(timepoint))) %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = timepoint, y = TTRavg, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, nrow = 3, scales = "free_y", labeller = labeller(treatment = facet_labels)) +
  ylab("Average Time-to-Right (s)") +
  scale_x_discrete(name = "timepoint",
                   breaks = c("0", "01", "02", "07", "12", "14", "19", "21", "26", "28", "42", "49"),
                   labels = c("0", "1", "2", "7", "12", "14", "19", "21", "26", "28", "42", "49")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) #Select unique timepoint, treatment, and TTRavg data. Replace numbers with characters to facilitate better ordering. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Add horizontal line for boxplot outlier threshold. Assign colors to each treatment. Increase base font size.
ggsave("figures/time-to-right-avg-boxplot-genotype.pdf", height = 8.5, width = 11)
```

### Difference in TTR

I'm going to do a similar analysis, but this time look at the effects of genotype on the difference in TTR between timepoints 2 and 7, timepoints 2 and 21, and timepoints 2 and 42. Although the 5ºC treatment was run for 49 timepoints, since we know time has a significant influence on TTR I'm going to use the same time point for all variables. Looking at differences with timepoint 7 can represent short-term differences, while timepoints 21 and 42 would be longer-term differences. timepoint 21 is also a nice parallel to the three-week experiment run in 2022. I also have timepoint 0 data for some of the crabs.

```{r}
modTTRSubsetDiff <- modTTRGenotype %>%
  select(., timepoint, crab.ID, TTRavg, treatment, final.genotype, presenceC, presenceT) %>%
  filter(., timepoint == 0 | timepoint == 2 | timepoint == 7 | timepoint == 21 | timepoint == 42) %>%
  pivot_wider(., names_from = timepoint, values_from = TTRavg, names_prefix = "timepoint") %>%
  mutate(., timepoint2v0 = timepoint2 - timepoint0) %>%
  mutate(., timepoint7v0 = timepoint7 - timepoint0) %>%
  mutate(., timepoint21v0 = timepoint21 - timepoint0) %>%
  mutate(., timepoint42v0 = timepoint42 - timepoint0) %>%
  mutate(., timepoint7v2 = timepoint7 - timepoint2) %>%
  mutate(., timepoint21v2 = timepoint21 - timepoint2) %>%
  mutate(., timepoint42v2 = timepoint42 - timepoint2) #For subset of the crabs, select timepoint, crab ID, average TTR, treatmet, and genotype information. Filter timepoints of interest and pivot wider, using timepoints as columns and values from TTRavg. Add "timepoint" as a prefix to all columns. Get the difference in TTR for comparisons of interest.
head(modTTRSubsetDiff)
```

#### Statistical comparison

```{r}
modTTRSubsetDiffStats <- modTTRSubsetDiff %>%
  filter(., is.na(timepoint0) == FALSE & is.na(timepoint2) == FALSE & is.na(timepoint7) == FALSE) %>%
  select(., -c(timepoint0, timepoint2, timepoint7, timepoint21, timepoint42, timepoint7v2, timepoint21v2, timepoint42v2)) %>%
  pivot_longer(., cols = c(timepoint2v0, timepoint7v0, timepoint21v0, timepoint42v0), names_to = "timepoint", values_to = "TTRdiff") %>%
  mutate(., timepoint = gsub(pattern = "timepoint", replacement = "", x = timepoint)) %>%
  mutate(., timepoint = gsub(pattern = "v0", replacement = "", x = timepoint)) %>%
  mutate(., timepoint = as.numeric(timepoint))
#Take difference in TTR information and subset crabs with data for timepoints 0, 2, and 7. Remove extraneous columns. Pivot data longer with values in TTRdiff and timepoint in a new column. Remove extra characters from timepoint values and convert into a numeric column
head(modTTRSubsetDiffStats) #Confirm dataframe creation
```

```{r}
TTRdiffGLM <- glm(log(TTRdiff + 1) ~ final.genotype + treatment*timepoint, data = modTTRSubsetDiffStats) #Run GLM to examine if difference in TTR from timepoint 0 is impacted by genotype, treatment, or timepoint
summary(TTRdiffGLM) #No effect of genotype
```

```{r}
step(TTRdiffGLM, direction = "backward") #Genotype not included in most parsimonious model
```

```{r}
TTRdiffGLM2 <- glm(log(TTRdiff + 1) ~ presenceC + presenceT + treatment*timepoint, data = modTTRSubsetDiffStats) #Run GLM to examine if difference in TTR from timepoint 0 is impacted by binary genotype, treatment, or timepoint
summary(TTRdiffGLM2) #No effect of genotype
```

```{r}
step(TTRdiffGLM2, direction = "backward") #Genotype not included in most parsimonious model
```

#### Plot

##### Facet by genotype

```{r}
modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = treatment, y = timepoint7v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = treatment), size = 2) +
  facet_wrap(. ~ final.genotype) + 
  labs(x = "Temperature (ºC)", y = "Difference in Individual TTR (s)", title = "timepoint 7 vs. timepoint 2") +
  scale_x_discrete(breaks = c("05C", "15C", "25C"),
                   labels = c("5", "15", "25")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
ggsave("figures/time-to-right-average-individuals-genotype-timepoint7v2.pdf", height = 8.5, width = 11)
```

```{r}
modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = treatment, y = timepoint21v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = treatment), size = 2) +
  facet_wrap(. ~ final.genotype) + 
  labs(x = "Temperature (ºC)", y = "Difference in Individual TTR (s)", title = "timepoint 21 vs. timepoint 2") +
  scale_x_discrete(breaks = c("05C", "15C", "25C"),
                   labels = c("5", "15", "25")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
ggsave("figures/time-to-right-average-individuals-genotype-timepoint21v2.pdf", height = 8.5, width = 11)
```

```{r}
modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = treatment, y = timepoint42v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = treatment), size = 2) +
  facet_wrap(. ~ final.genotype) + 
  labs(x = "Temperature (ºC)", y = "Difference in Individual TTR (s)", title = "timepoint 42 vs. timepoint 2") +
  scale_x_discrete(breaks = c("05C", "15C", "25C"),
                   labels = c("5", "15", "25")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
ggsave("figures/time-to-right-average-individuals-genotype-timepoint42v2.pdf", height = 8.5, width = 11)
```

```{r}
timepoint7vs2Plot <- modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = treatment, y = timepoint7v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = treatment), size = 2) +
  facet_wrap(. ~ final.genotype) + 
  labs(x = "", y = "", title = "A. timepoint 7 vs. timepoint 2") +
  scale_y_continuous(breaks = seq(-5, 15, 5),
                     limits = c(-5,15)) +
  scale_x_discrete(breaks = c("05C", "15C", "25C"),
                   labels = c("5", "15", "25")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(legend.position = "none",
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank())
timepoint7vs2Plot
```

```{r}
timepoint21vs2Plot <- modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = treatment, y = timepoint21v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = treatment), size = 2) +
  facet_wrap(. ~ final.genotype) + 
  labs(x = "", y = "Difference in Individual TTR (s)", title = "B. timepoint 21 vs. timepoint 2") +
  scale_y_continuous(breaks = seq(-1,5,1),
                     limits = c(-1,5)) +
  scale_x_discrete(breaks = c("05C", "15C", "25C"),
                   labels = c("5", "15", "25")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(legend.position = "none",
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank())
timepoint21vs2Plot
```

```{r}
timepoint42v2Plot <- modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = treatment, y = timepoint42v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = treatment), size = 2) +
  facet_wrap(. ~ final.genotype) + 
  labs(x = "Temperature (ºC)", y = "", title = "C. timepoint 42 vs. timepoint 2") +
  scale_y_continuous(breaks = seq(-1,6,1),
                     limits = c(-1,6)) +
  scale_x_discrete(breaks = c("05C", "15C", "25C"),
                   labels = c("5", "15", "25")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(15, 19, 17),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
timepoint42v2Plot
```

```{r}
timepoint7vs2Plot / timepoint21vs2Plot / timepoint42v2Plot
ggsave("figures/time-to-right-average-individuals-genotype-multipanel.pdf", height = 11, width = 8.5)
```

##### Facet by temperature

```{r}
modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = final.genotype, y = timepoint7v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) +
  labs(x = "Genotype", y = "Difference in Individual TTR (s)", title = "timepoint 7 vs. timepoint 2") +
  scale_x_discrete(breaks = c("CC", "CT", "TT")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
ggsave("figures/time-to-right-average-individuals-genotype-bytemp-timepoint7v2.pdf", height = 8.5, width = 11)
```

```{r}
modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = final.genotype, y = timepoint21v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) +
  labs(x = "Genotype", y = "Difference in Individual TTR (s)", title = "timepoint 21 vs. timepoint 2") +
  scale_x_discrete(breaks = c("CC", "CT", "TT")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
ggsave("figures/time-to-right-average-individuals-genotype-bytemp-timepoint21v2.pdf", height = 8.5, width = 11)
```

```{r}
modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = final.genotype, y = timepoint42v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) +
  labs(x = "Genotype", y = "Difference in Individual TTR (s)", title = "timepoint 42 vs. timepoint 2") +
  scale_x_discrete(breaks = c("CC", "CT", "TT")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
ggsave("figures/time-to-right-average-individuals-genotype-bytemp-timepoint42v2.pdf", height = 8.5, width = 11)
```

```{r}
timepoint7vs2byTempPlot <- modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = final.genotype, y = timepoint7v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) + 
  labs(x = "", y = "", title = "A. timepoint 7 vs. timepoint 2") +
  scale_x_discrete(breaks = c("CC", "CT", "TT")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(legend.position = "none",
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank())
timepoint7vs2byTempPlot
```

```{r}
timepoint21vs2byTempPlot <- modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = final.genotype, y = timepoint21v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) + 
  labs(x = "", y = "Difference in Individual TTR (s)", title = "B. timepoint 21 vs. timepoint 2") +
  scale_x_discrete(breaks = c("CC", "CT", "TT")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(legend.position = "none",
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank())
timepoint21vs2byTempPlot
```

```{r}
timepoint42v2byTempPlot <- modTTRSubsetDiff %>%
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment)) %>%
  ggplot(mapping = aes(x = final.genotype, y = timepoint42v2, color = treatment)) +
  geom_boxplot(outlier.shape = NA) + geom_point(aes(shape = final.genotype), size = 2) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) + 
  labs(x = "Genotype", y = "", title = "C. timepoint 42 vs. timepoint 2") +
  scale_x_discrete(breaks = c("CC", "CT", "TT")) +
  scale_color_manual(values = c(plotColors[3], plotColors[2], plotColors[1]),
                     name = "Temperature (ºC)",
                     breaks = c("05C", "15C", "25C"),
                     labels = c("5", "15", "25")) +
  scale_shape_manual(values = c(3, 4, 1),
                     name = "Genotype",
                     breaks = c("CC", "CT", "TT"),
                     labels = c("CC", "CT", "TT")) +
  theme_classic(base_size = 15) + theme(legend.position = "none")
timepoint42v2byTempPlot
```

```{r}
timepoint7vs2byTempPlot / timepoint21vs2byTempPlot / timepoint42v2byTempPlot
ggsave("figures/time-to-right-average-individuals-genotype-byTemp-multipanel.pdf", height = 11, width = 8.5)
```

## Q10 analysis

Q10 is a metric used to quantify temperature sensitivity, generally defined as the change in a physiological response over a 10 degree temperature interval. For most organisms, Q10 is around 2 (a two-fold change in a physiological or chemical response due to a 10 degree temperature change). I will use the full dataset to calculate Q10 values between timepoint 0 and 2, timepoint 0 and 7, timepoint 0 and 21, and timepoint 0 and 42. I can then investigate how genotype influences Q10.

```{r}
modTTRQ10 <- modTTR %>%
  select(., timepoint, crab.ID, TTRavg, treatment, final.genotype) %>%
  mutate(., presenceC = case_when(final.genotype == "CC" ~ "Y",
                                  final.genotype == "CT" ~ "Y",
                                  final.genotype == "TT" ~ "N")) %>%
  mutate(., presenceT = case_when(final.genotype == "CC" ~ "N",
                                  final.genotype == "CT" ~ "Y",
                                  final.genotype == "TT" ~ "Y")) %>%
  filter(., timepoint == 0 | timepoint == 2 | timepoint == 7 | timepoint == 21 | timepoint == 42) %>%
  filter(., treatment != "15C") %>%
  pivot_wider(., names_from = timepoint, values_from = TTRavg, names_prefix = "timepoint") %>%
  filter(., is.na(timepoint0) == FALSE) %>% 
  mutate(temperature = case_when(treatment == "15C" ~ 15,
                                 treatment == "25C" ~ 25,
                                 treatment == "5C" ~ 5)) %>%
  arrange(., temperature) %>%
  mutate(., timepoint2_Q10 = Q10(R1 = timepoint0, 
                                 R2 = timepoint2, 
                                 T1 = 15, 
                                 T2 = temperature)) %>%
  mutate(., timepoint7_Q10 = Q10(R1 = timepoint0, 
                                 R2 = timepoint7, 
                                 T1 = 15, 
                                 T2 = temperature)) %>%
  mutate(., timepoint21_Q10 = Q10(R1 = timepoint0, 
                                  R2 = timepoint21, 
                                  T1 = 15, 
                                  T2 = temperature)) %>%
  mutate(., timepoint42_Q10 = Q10(R1 = timepoint0, 
                                  R2 = timepoint42, 
                                  T1 = 15, 
                                  T2 = temperature)) %>% 
  select(., crab.ID, treatment, final.genotype, presenceC, presenceT, timepoint2_Q10, timepoint7_Q10, timepoint21_Q10, timepoint42_Q10) %>%
  pivot_longer(., cols = !c(crab.ID, treatment, final.genotype, presenceC, presenceT), names_to = "timepoint", values_to = "Q10") %>%
  filter(., is.na(Q10) == FALSE) %>%
  mutate(., timepoint = gsub(x = timepoint, pattern = "_Q10", replacement = "")) %>%
  mutate(., timepoint = gsub(x = timepoint, pattern = "timepoint", replacement = "")) %>%
  mutate(., timepoint = case_when(timepoint == "2" ~ "02",
                                  timepoint == "7" ~ "07",
                                  timepoint != "2" | timepoint != "7" ~ timepoint)) %>% 
  mutate(., treatment = case_when(treatment == "5C" ~ "05C",
                                  treatment != "5C" ~ treatment))
#Select columns of interest. Add columns for presence of C or T alleles. Filter timepoints of interest, remove 15C crabs, and pivot wider. Remove crabs without a timepoint 0 measurement. Add a column for temperature based on the treatment column, and arrange data by temperature. Use the Q10 function within the respirometry package to calculate Q10 values for timepoint 0 vs. timepoint 2, 7, 21, and 42. Remove any rows without a timepoint 2 Q10 value. Select columns of interest and pivot longer. Remove any rows without a Q10 value. Use a series of gsub to modify timepoint column and convert to a character format that would be organized numerically. Use a similar approach for treatment.
head(modTTRQ10)
length(modTTRQ10$crab.ID %>% unique(.)) #18 crabs
```

```{r}
modTTRQ10 %>%
  ggplot(., mapping = aes(x = timepoint, y = Q10, color = treatment, shape = treatment, group = crab.ID)) +
  geom_point(size = 2) + geom_line(lty = 3) +
  facet_wrap(. ~ treatment, scales = "free_y", labeller = labeller(treatment = facet_labels)) +
  scale_x_discrete(name = "timepoint",
                   breaks = c("02", "07", "21", "42"),
                   labels = c("2", "7", "21", "42")) +
  scale_color_manual(values = c(plotColors[3], plotColors[1]),
                     guide = "none") +
  scale_shape_manual(values = c(15, 17),
                     guide = "none") +
  theme_classic(base_size = 15) #Plot Q10 data with timepoint on the x-axis, Q10 on the y-axis, and color and shape defined by treatment. Group by crab.ID. Add points and lines connecting values for each crab. Facet by treatment, allow y-axis to vary by each facet, and include facet labels. Add x-axis label and specify x-axis tick labels. Add shape and color values for each treatment but do not include a legend.
ggsave("figures/time-to-right-average-individuals-Q10.pdf", height = 8.5, width = 11)
```

