---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "12/29/2024"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/MA/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/MA/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("RColorBrewer")
require(tidyverse)
require(RColorBrewer)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1a <- read.csv("../../../data/MA/HOBO/T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..21435405..SEN.S.N..21435405.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1b <- read.csv("../../../data/MA/HOBO/T1B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..20548756..SEN.S.N..20548756..LBL..T1B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2a <- read.csv("../../../data/MA/HOBO/T2A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..21435404..SEN.S.N..21435404..LBL..T2A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2b <- read.csv("../../../data/MA/HOBO/T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..T2B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank3a <- read.csv("../../../data/MA/HOBO/T3A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3a = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..T3A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank3b <- read.csv("../../../data/MA/HOBO/T3B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3b = Temp...C..LGR.S.N..21435416..SEN.S.N..21435416.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank5a <- read.csv("../../../data/MA/HOBO/T5A.csv", skip = 1,  header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5a = Temp...C..LGR.S.N..20548752..SEN.S.N..20548752.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank5b <- read.csv("../../../data/MA/HOBO/T5B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5b = Temp...C..LGR.S.N..21435412..SEN.S.N..21435412.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank6a <- read.csv("../../../data/MA/HOBO/T6A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6a = Temp...C..LGR.S.N..21435414..SEN.S.N..21435414.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank6b <- read.csv("../../../data/MA/HOBO/T6B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6b = Temp...C..LGR.S.N..21435417..SEN.S.N..21435417..LBL..T6B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank7a <- read.csv("../../../data/MA/HOBO/T7A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7a = Temp...C..LGR.S.N..20548776..SEN.S.N..20548776..LBL..T7A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank7b <- read.csv("../../../data/MA/HOBO/T7B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7b = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..T7B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment
#Retain only rows with data before the experiment ended

tank1a <- tank1a %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 10:422)
tank1b <- tank1b %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 10:422)

tank2a <- tank2a %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 10:422)
tank2b <- tank2b %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 10:422)

tank3a <- tank3a %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)
tank3b <- tank3b %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)

tank5a <- tank5a %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 10:422)
tank5b <- tank5b %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 10:422)

tank6a <- tank6a %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)
tank6b <- tank6b %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)

tank7a <- tank7a %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:422)
tank7b <- tank7b %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:422)
```

# Calcualte temperature conditions before ramp down

```{r}
hobos <- list(tank1a, tank2a, tank3a, tank5a, tank6a, tank7a,
              tank1b, tank2b, tank3b, tank5b, tank6b, tank7b) #Create a list of filenames to cycle through
hobos
```

The first ~40 rows are the end of the acclimation period.

```{r}
#Acclimation: Mean = 14.83809, SD = 0.5346238
acclimationMeanMA <- mean(rbind(hobos[[1]][,2][1:40],
                                hobos[[2]][,2][1:40],
                                hobos[[3]][,2][1:40],
                                hobos[[4]][,2][1:40],
                                hobos[[5]][,2][1:40],
                                hobos[[6]][,2][1:40],
                                hobos[[7]][,2][1:40],
                                hobos[[8]][,2][1:40],
                                hobos[[9]][,2][1:40]),
                          na.rm = TRUE)
acclimationSDMA <- sd(rbind(hobos[[1]][,2][1:40],
                            hobos[[2]][,2][1:40],
                            hobos[[3]][,2][1:40],
                            hobos[[4]][,2][1:40],
                            hobos[[5]][,2][1:40],
                            hobos[[6]][,2][1:40],
                            hobos[[7]][,2][1:40],
                            hobos[[8]][,2][1:40],
                            hobos[[9]][,2][1:40]),
                      na.rm = TRUE)
acclimationMeanMA
acclimationSDMA
```

# Calculate treatment conditions after ramp down



## Calculate treatment mean and SE

### For each tank

Row 631 is when HOBO loggers are recording data on the first full day of the experiment after the temperatures have reached the treatment setpoints, so I'll use that as my cutoff.

```{r}
treatmentTempStats <- data.frame("tank" = c(1:3, 5:7),
                                 "meanTemp" = rep(0, times = 6),
                                 "sdTemp" = rep(0, times = 6)) #Create an empty dataframe
treatmentTempStats
```

```{r}
for (i in 1:(length(hobos)/2)) {
  treatmentTempStats[i,2] <-
    mean(rbind(hobos[[i]][,2][88:length(hobos[[i]][,2])],
               hobos[[i+6]][,2][88:length(hobos[[i+6]][,2])]), na.rm = TRUE)
  treatmentTempStats[i,3] <-
    sd(rbind(hobos[[i]][,2][88:length(hobos[[i]][,2])],
             hobos[[i+6]][,2][88:length(hobos[[i+6]][,2])]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers,either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

### For each treatment

```{r}
#Colder: Mean = 1.66, SD = 0.3231426
colderMeanMA <- mean(rbind(hobos[[1]][,2][88:length(hobos[[1]][,2])],
                           hobos[[2]][,2][88:length(hobos[[2]][,2])],
                           hobos[[3]][,2][88:length(hobos[[3]][,2])],
                           hobos[[7]][,2][88:length(hobos[[7]][,2])],
                           hobos[[8]][,2][88:length(hobos[[8]][,2])],
                           hobos[[9]][,2][88:length(hobos[[9]][,2])]),
                     na.rm = TRUE)
colderSDMA <- sd(rbind(hobos[[1]][,2][88:length(hobos[[1]][,2])],
                       hobos[[2]][,2][88:length(hobos[[2]][,2])],
                       hobos[[3]][,2][88:length(hobos[[3]][,2])],
                       hobos[[7]][,2][88:length(hobos[[7]][,2])],
                       hobos[[8]][,2][88:length(hobos[[8]][,2])],
                       hobos[[9]][,2][88:length(hobos[[9]][,2])]),
                 na.rm = TRUE)
colderMeanMA
colderSDMA
```

```{r}
#Colder: Mean = 5.385082, SD = 0.3004675
coldMeanMA <- mean(rbind(hobos[[4]][,2][88:length(hobos[[4]][,2])],
                         hobos[[5]][,2][88:length(hobos[[5]][,2])],
                         hobos[[6]][,2][88:length(hobos[[6]][,2])],
                         hobos[[10]][,2][88:length(hobos[[10]][,2])],
                         hobos[[11]][,2][88:length(hobos[[11]][,2])],
                         hobos[[12]][,2][88:length(hobos[[12]][,2])]),
                   na.rm = TRUE)
coldSDMA <- sd(rbind(hobos[[4]][,2][88:length(hobos[[4]][,2])],
                     hobos[[5]][,2][88:length(hobos[[5]][,2])],
                     hobos[[6]][,2][88:length(hobos[[6]][,2])],
                     hobos[[10]][,2][88:length(hobos[[10]][,2])],
                     hobos[[11]][,2][88:length(hobos[[11]][,2])],
                     hobos[[12]][,2][88:length(hobos[[12]][,2])]),
               na.rm = TRUE)
coldMeanMA
coldSDMA
```

# Statistical test

I will conduct use a Kruskal-Wallis test to confirm there are significant differences between treatments.

```{r}
tempDataLong <- rbind(tank1a %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp1a"), 
                      tank1b %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp1b"),
                      tank2a %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp2a"), 
                      tank2b %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp2b"),
                      tank3a %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp3a"), 
                      tank3b %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp3b"),
                      tank5a %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp5a"), 
                      tank5b %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp5b"),
                      tank6a %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp6a"), 
                      tank6b %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp6b"),
                      tank7a %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp7a"), 
                      tank7b %>%
                        slice(-c(1:87)) %>%
                        dplyr::rename(temp = "temp7b")) %>% 
  drop_na() #Create long dataframe for statistical testing. Remove ramp down temperatures and rows with NAs.
tail(tempDataLong) #Confirm dataframe creation
```

```{r}
tempStats <- kruskal.test(temp ~ treatment, data = tempDataLong) #One-way ANOVA by treatment
tempStats #Significant differences between temperature treatments. Chi-sq: 2941.37, p-value = 0

broom::tidy(tempStats)
```

# Plot full temperature conditions (including ramp down)

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Blues")[9],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

```{r}
dataTempPlotMA <- cbind(tank1a, tank1b,
      tank2a, tank2b,
      tank3a, tank3b,
      tank5a, tank5b,
      tank6a, tank6b,
      tank7a, tank7b) %>%
  dplyr::select(., c(1, 4, 
                     2, 6, 
                     10, 14, 
                     18, 22,
                     26, 30,
                     34, 38,
                     42, 46)) %>%
  rowwise() %>%
  mutate(., colderTemp = mean(c(temp1a, temp1b,
                                temp2a, temp2b,
                                temp3a, temp3b)),
         coldTemp = mean(c(temp5a, temp5b,
                           temp6a, temp6b,
                           temp7a, temp7b))) #Create dataframe to use for plotting
head(dataTempPlotMA) #Confirm changes
dataTempPlotMA$dateTime[88] #Date-time for start of experimental period
dataTempPlotMA$dateTime[413] #Date-time for end of experimental period
```

```{r}
dataTempPlotMA %>%
  ggplot(., aes(x = dateTime, y = colderTemp)) +
  geom_line(color = plotColors[1]) +
  geom_ribbon(data = dataTempPlotMA[88:413,], aes(x = dateTime, y = coldMeanMA, ymin = coldMeanMA - coldSDMA, ymax = coldMeanMA + coldSDMA), fill = plotColors[2], alpha = 0.15) + #Start SE ribbon at the right place
  geom_segment(x = as.POSIXct(ymd_hms("2024-06-25 06:00:00")), xend = as.POSIXct(ymd_hms("2024-06-28 09:51:00")),
               y = coldMeanMA, yend = coldMeanMA, 
               colour = plotColors[2], linetype = 3) + #Start average line at the right place
  geom_line(aes(x = dateTime, y = coldTemp), color = plotColors[2]) +
  geom_ribbon(data = dataTempPlotMA[88:413,], aes(x = dateTime, y = colderMeanMA, ymin = colderMeanMA - colderSDMA, ymax = colderMeanMA + colderSDMA), fill = plotColors[1], alpha = 0.15) + #Start SE ribbon at the right place
  geom_segment(x = as.POSIXct(ymd_hms("2024-06-25 06:00:00")), xend = as.POSIXct(ymd_hms("2024-06-28 09:51:00")),
               y = colderMeanMA, yend = colderMeanMA,
               colour = plotColors[1], linetype = 3) + #Start average line at the right place
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-24 07:30:00")), linetype = 3, color = "black") + #Hour 0 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-24 18:00:00")), linetype = 3, color = "black") + #Hour 4 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-25 09:00:00")), linetype = 3, color = "black") + #Hour 22 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-25 15:00:00")), linetype = 3, color = "black") + #Hour 28 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-26 09:00:00")), linetype = 3, color = "black") + #Hour 46 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-28 09:00:00")), linetype = 3, color = "black") + #Hour 96 TTR
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     breaks = c(seq(0,5,1), seq(5,15,5))) +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

```{r}
MAconditions <- dataTempPlotMA %>%
  ggplot(., aes(x = dateTime, y = colderTemp)) +
  geom_line(color = plotColors[1]) +
  geom_ribbon(data = dataTempPlotMA[88:413,], aes(x = dateTime, y = coldMeanMA, ymin = coldMeanMA - coldSDMA, ymax = coldMeanMA + coldSDMA), fill = plotColors[2], alpha = 0.15) + #Start SE ribbon at the right place
  geom_segment(x = as.POSIXct(ymd_hms("2024-06-25 06:00:00")), xend = as.POSIXct(ymd_hms("2024-06-28 09:51:00")),
               y = coldMeanMA, yend = coldMeanMA, 
               colour = plotColors[2], linetype = 3) + #Start average line at the right place
  geom_line(aes(x = dateTime, y = coldTemp), color = plotColors[2]) +
  geom_ribbon(data = dataTempPlotMA[88:413,], aes(x = dateTime, y = colderMeanMA, ymin = colderMeanMA - colderSDMA, ymax = colderMeanMA + colderSDMA), fill = plotColors[1], alpha = 0.15) + #Start SE ribbon at the right place
  geom_segment(x = as.POSIXct(ymd_hms("2024-06-25 06:00:00")), xend = as.POSIXct(ymd_hms("2024-06-28 09:51:00")),
               y = colderMeanMA, yend = colderMeanMA,
               colour = plotColors[1], linetype = 3) + #Start average line at the right place
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-24 07:30:00")), linetype = 3, color = "black") + #Hour 0 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-24 18:00:00")), linetype = 3, color = "black") + #Hour 4 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-25 09:00:00")), linetype = 3, color = "black") + #Hour 22 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-25 15:00:00")), linetype = 3, color = "black") + #Hour 28 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-26 09:00:00")), linetype = 3, color = "black") + #Hour 46 TTR
  geom_vline(xintercept = as.POSIXct(ymd_hms("2024-06-28 09:00:00")), linetype = 3, color = "black") + #Hour 96 TTR
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     breaks = c(seq(0,5,1), seq(5,15,5))) +
  ggtitle("A. MA Temperature Conditions") +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.

MAconditions
```

```{r}
save.image("~/Documents/cold-green-crab/output/MA/01-temp-conditions/01-temp-conditions.RData") #Save R Data to facilitate making a multipanel figure
```

