---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "12/29/2024"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/MA/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/MA/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("RColorBrewer")
require(tidyverse)
require(RColorBrewer)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1a <- read.csv("../../../data/MA/HOBO/T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..21435405..SEN.S.N..21435405.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1b <- read.csv("../../../data/MA/HOBO/T1B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..20548756..SEN.S.N..20548756..LBL..T1B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2a <- read.csv("../../../data/MA/HOBO/T2A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..21435404..SEN.S.N..21435404..LBL..T2A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2b <- read.csv("../../../data/MA/HOBO/T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..T2B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank3a <- read.csv("../../../data/MA/HOBO/T3A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3a = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..T3A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank3b <- read.csv("../../../data/MA/HOBO/T3B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3b = Temp...C..LGR.S.N..21435416..SEN.S.N..21435416.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank5a <- read.csv("../../../data/MA/HOBO/T5A.csv", skip = 1,  header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5a = Temp...C..LGR.S.N..20548752..SEN.S.N..20548752.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank5b <- read.csv("../../../data/MA/HOBO/T5B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5b = Temp...C..LGR.S.N..21435412..SEN.S.N..21435412.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank6a <- read.csv("../../../data/MA/HOBO/T6A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6a = Temp...C..LGR.S.N..21435414..SEN.S.N..21435414.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank6b <- read.csv("../../../data/MA/HOBO/T6B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6b = Temp...C..LGR.S.N..21435417..SEN.S.N..21435417..LBL..T6B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank7a <- read.csv("../../../data/MA/HOBO/T7A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7a = Temp...C..LGR.S.N..20548776..SEN.S.N..20548776..LBL..T7A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank7b <- read.csv("../../../data/MA/HOBO/T7B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7b = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..T7B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment
#Retain only rows with data before the experiment ended

tank1a <- tank1a %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 10:422)
tank1b <- tank1b %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 10:422)

tank2a <- tank2a %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 10:422)
tank2b <- tank2b %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 10:422)

tank3a <- tank3a %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)
tank3b <- tank3b %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)

tank5a <- tank5a %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 10:422)
tank5b <- tank5b %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 10:422)

tank6a <- tank6a %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)
tank6b <- tank6b %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)

tank7a <- tank7a %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:422)
tank7b <- tank7b %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:422)
```

# Calculate treatment conditions after ramp down

## Calculate treatment mean and SE

```{r}
hobos <- list(tank1a, tank2a, tank3a, tank5a, tank6a, tank7a,
              tank1b, tank2b, tank3b, tank5b, tank6b, tank7b) #Create a list of filenames to cycle through
hobos
```

### For each tank

Row 631 is when HOBO loggers are recording data on the first full day of the experiment after the temperatures have reached the treatment setpoints, so I'll use that as my cutoff.

```{r}
treatmentTempStats <- data.frame("tank" = c(1:3, 5:7),
                                 "meanTemp" = rep(0, times = 6),
                                 "sdTemp" = rep(0, times = 6)) #Create an empty dataframe
treatmentTempStats
```

```{r}
for (i in 1:(length(hobos)/2)) {
  treatmentTempStats[i,2] <-
    mean(rbind(hobos[[i]][,2][88:length(hobos[[i]][,2])],
               hobos[[i+6]][,2][88:length(hobos[[i+6]][,2])]), na.rm = TRUE)
  treatmentTempStats[i,3] <-
    sd(rbind(hobos[[i]][,2][88:length(hobos[[i]][,2])],
             hobos[[i+6]][,2][88:length(hobos[[i+6]][,2])]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers,either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

### For each treatment

```{r}
#Colder: Mean = 1.66, SD = 0.3231426
colderMean <- mean(rbind(hobos[[1]][,2][88:length(hobos[[1]][,2])],
                         hobos[[2]][,2][88:length(hobos[[2]][,2])],
                         hobos[[3]][,2][88:length(hobos[[3]][,2])],
                         hobos[[7]][,2][88:length(hobos[[7]][,2])],
                         hobos[[8]][,2][88:length(hobos[[8]][,2])],
                         hobos[[9]][,2][88:length(hobos[[9]][,2])]),
                   na.rm = TRUE)
colderSD <- sd(rbind(hobos[[1]][,2][88:length(hobos[[1]][,2])],
                     hobos[[2]][,2][88:length(hobos[[2]][,2])],
                     hobos[[3]][,2][88:length(hobos[[3]][,2])],
                     hobos[[7]][,2][88:length(hobos[[7]][,2])],
                     hobos[[8]][,2][88:length(hobos[[8]][,2])],
                     hobos[[9]][,2][88:length(hobos[[9]][,2])]),
               na.rm = TRUE)
colderMean
colderSD
```

```{r}
#Colder: Mean = 5.385082, SD = 0.3004675
coldMean <- mean(rbind(hobos[[4]][,2][88:length(hobos[[4]][,2])],
                       hobos[[5]][,2][88:length(hobos[[5]][,2])],
                       hobos[[6]][,2][88:length(hobos[[6]][,2])],
                       hobos[[10]][,2][88:length(hobos[[10]][,2])],
                       hobos[[11]][,2][88:length(hobos[[11]][,2])],
                       hobos[[12]][,2][88:length(hobos[[12]][,2])]),
                 na.rm = TRUE)
coldSD <- sd(rbind(hobos[[4]][,2][88:length(hobos[[4]][,2])],
                   hobos[[5]][,2][88:length(hobos[[5]][,2])],
                   hobos[[6]][,2][88:length(hobos[[6]][,2])],
                   hobos[[10]][,2][88:length(hobos[[10]][,2])],
                   hobos[[11]][,2][88:length(hobos[[11]][,2])],
                   hobos[[12]][,2][88:length(hobos[[12]][,2])]),
             na.rm = TRUE)
coldMean
coldSD
```

# Statistical test

I will conduct a one-way ANOVA to confirm there are significant differences between treatments.

```{r}
rbind(tank1a, tank1b)
```
```{r}

```


```{r}
tempDataLong <- rbind(tank1a %>%
                        dplyr::rename(temp = "temp1a"), 
                      tank1b %>%
                        dplyr::rename(temp = "temp1b"),
                      tank2a %>%
                        dplyr::rename(temp = "temp2a"), 
                      tank2b %>%
                        dplyr::rename(temp = "temp2b"),
                      tank3a %>%
                        dplyr::rename(temp = "temp3a"), 
                      tank3b %>%
                        dplyr::rename(temp = "temp3b"),
                      tank5a %>%
                        dplyr::rename(temp = "temp5a"), 
                      tank5b %>%
                        dplyr::rename(temp = "temp5b"),
                      tank6a %>%
                        dplyr::rename(temp = "temp6a"), 
                      tank6b %>%
                        dplyr::rename(temp = "temp6b"),
                      tank7a %>%
                        dplyr::rename(temp = "temp7a"), 
                      tank7b %>%
                        dplyr::rename(temp = "temp7b")) %>% 
  drop_na() #Create long dataframe for ANOVA and remove rows with NAs
tail(tempDataLong) #Confirm dataframe creation
```

```{r}
tempANOVA <- aov(temp ~ treatment, data = tempDataLong) #One-way ANOVA by treatment
summary(tempANOVA)[[1]][4][[1]][1] # F = 1028.818
summary(tempANOVA)[[1]][5][[1]][1] # P-value = 2.797163e-205
```

# Plot full temperature conditions (including ramp down)

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Blues")[9],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

```{r}
cbind(tank1a, tank1b,
      tank2a, tank2b,
      tank3a, tank3b,
      tank5a, tank5b,
      tank6a, tank6b,
      tank7a, tank7b) %>%
  dplyr::select(., c(1, 4, 
                     2, 6, 
                     10, 14, 
                     18, 22,
                     26, 30,
                     34, 38,
                     42, 46)) %>%
  rowwise() %>%
  mutate(., colderTemp = mean(c(temp1a, temp1b,
                                temp2a, temp2b,
                                temp3a, temp3b)),
         coldTemp = mean(c(temp5a, temp5b,
                           temp6a, temp6b,
                           temp7a, temp7b))) %>%
  ggplot(., aes(x = dateTime, y = colderTemp)) +
  geom_line(color = plotColors[1]) +
  geom_ribbon(aes(x = dateTime, y = coldMean, ymin = coldMean - coldSD, ymax = coldMean + coldSD), fill = plotColors[2], alpha = 0.15) +
  geom_hline(yintercept = coldMean, colour = plotColors[2], linetype = 3) +
  geom_line(aes(x = dateTime, y = coldTemp), color = plotColors[2]) +
  geom_ribbon(aes(x = dateTime, y = colderMean, ymin = colderMean - colderSD, ymax = colderMean + colderSD), fill = plotColors[1], alpha = 0.15) +
  geom_hline(yintercept = colderMean, colour = plotColors[1], linetype = 3) +
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     breaks = c(seq(0,5,1), seq(5,15,5))) +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

