---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "8/18/2022"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/MA/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/MA/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("qpcR")
#install.packages("RColorBrewer")
#install.packages("patchwork")
require(qpcR)
require(tidyverse)
require(RColorBrewer)
require(patchwork)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1a <- read.csv("../../../data/MA/HOBO/T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435405..SEN.S.N..21435405.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1b <- read.csv("../../../data/MA/HOBO/T1B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..20548756..SEN.S.N..20548756..LBL..T1B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2a <- read.csv("../../../data/MA/HOBO/T2A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435404..SEN.S.N..21435404..LBL..T2A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2b <- read.csv("../../../data/MA/HOBO/T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..T2B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank3a <- read.csv("../../../data/MA/HOBO/T3A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..T3A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank3b <- read.csv("../../../data/MA/HOBO/T3B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435416..SEN.S.N..21435416.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank5a <- read.csv("../../../data/MA/HOBO/T5A.csv", skip = 1,  header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..20548752..SEN.S.N..20548752.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank5b <- read.csv("../../../data/MA/HOBO/T5B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435412..SEN.S.N..21435412.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank6a <- read.csv("../../../data/MA/HOBO/T6A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435414..SEN.S.N..21435414.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank6b <- read.csv("../../../data/MA/HOBO/T6B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..21435417..SEN.S.N..21435417..LBL..T6B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank7a <- read.csv("../../../data/MA/HOBO/T7A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..20548776..SEN.S.N..20548776..LBL..T7A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank7b <- read.csv("../../../data/MA/HOBO/T7B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..T7B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment

tank1a <- tank1a %>% 
  mutate(., tank = 1,
         treatment = "colder")
tank1b <- tank1b %>% 
  mutate(., tank = 1,
         treatment = "colder")

tank2a <- tank2a %>% 
  mutate(., tank = 2,
         treatment = "colder")
tank2b <- tank2b %>% 
  mutate(., tank = 2,
         treatment = "colder")

tank3a <- tank3a %>% 
  mutate(., tank = 3,
         treatment = "colder")
tank3b <- tank3b %>% 
  mutate(., tank = 3,
         treatment = "colder")

tank5a <- tank5a %>% 
  mutate(., tank = 5,
         treatment = "cold")
tank5b <- tank5b %>% 
  mutate(., tank = 5,
         treatment = "cold")

tank6a <- tank6a %>% 
  mutate(., tank = 6,
         treatment = "cold")
tank6b <- tank6b %>% 
  mutate(., tank = 6,
         treatment = "cold")

tank7a <- tank7a %>% 
  mutate(., tank = 7,
         treatment = "cold")
tank7b <- tank7b %>% 
  mutate(., tank = 7,
         treatment = "cold")
```

# Treatment conditions

## Calculate treatment mean and SE

```{r}
hobos <- list(tank1a, tank2a, tank3a, tank5a, tank6a, tank7a,
              tank1b, tank2b, tank3b, tank5b, tank6b, tank7b) #Create a list of filenames to cycle through
```

### For each tank

Row 631 is when HOBO loggers are recording data on the first full day of the experiment after the temperatures have reached the treatment setpoints, so I'll use that as my cutoff.

```{r}
treatmentTempStats <- data.frame("tank" = c(seq(1:3), seq(5:7)),
                                 "meanTemp" = rep(0, times = 6),
                                 "sdTemp" = rep(0, times = 6)) #Create an empty dataframe
treatmentTempStats
```

```{r}
for (i in 1:(length(hobos)/2)) {
  treatmentTempStats[i,2] <- mean(rbind(hobos[[i]][,2][631:length(hobos[[i]][,2])], hobos[[i+6]][,2][631:length(hobos[[i+6]][,2])]), na.rm = TRUE)
  treatmentTempStats[i,3] <- sd(rbind(hobos[[i]][,2][631:length(hobos[[i]][,2])], hobos[[i+6]][,2][631:length(hobos[[i+6]][,2])]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers,  either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

### For each treatment

```{r}
#Ambient: Mean = 13.79212, SD = 0.9136427
ambientMean <- mean(rbind(hobos[[1]][,2][631:length(hobos[[1]][,2])],
                          hobos[[7]][,2][631:length(hobos[[7]][,2])],
                          hobos[[4]][,2][631:length(hobos[[4]][,2])],
                          hobos[[10]][,2][631:length(hobos[[10]][,2])]),
                    na.rm = TRUE)
ambientSD <- sd(rbind(hobos[[1]][,2][631:length(hobos[[1]][,2])],
                      hobos[[7]][,2][631:length(hobos[[7]][,2])],
                      hobos[[4]][,2][631:length(hobos[[4]][,2])],
                      hobos[[10]][,2][631:length(hobos[[10]][,2])]),
                na.rm = TRUE)
ambientMean
ambientSD
```

```{r}
#Cold: Mean = 5.666688, SD = 0.3637059
coldMean <- mean(rbind(hobos[[2]][,2][631:length(hobos[[2]][,2])],
                       hobos[[8]][,2][631:length(hobos[[8]][,2])],
                       hobos[[5]][,2][631:length(hobos[[5]][,2])],
                       hobos[[11]][,2][631:length(hobos[[11]][,2])]),
                 na.rm = TRUE)
coldSD <- sd(rbind(hobos[[2]][,2][631:length(hobos[[2]][,2])],
                   hobos[[8]][,2][631:length(hobos[[8]][,2])],
                   hobos[[5]][,2][631:length(hobos[[5]][,2])],
                   hobos[[11]][,2][631:length(hobos[[11]][,2])]),
             na.rm = TRUE)
coldMean
coldSD
```

```{r}
#Warm: Mean = 30.81052, SD = 2.360069
warmMean <- mean(rbind(hobos[[3]][,2][631:length(hobos[[3]][,2])],
                       hobos[[9]][,2][631:length(hobos[[9]][,2])],
                       hobos[[6]][,2][631:length(hobos[[6]][,2])],
                       hobos[[12]][,2][631:length(hobos[[12]][,2])]),
                 na.rm = TRUE)
warmSD <- sd(rbind(hobos[[3]][,2][631:length(hobos[[3]][,2])],
                   hobos[[9]][,2][631:length(hobos[[9]][,2])],
                   hobos[[6]][,2][631:length(hobos[[6]][,2])],
                   hobos[[12]][,2][631:length(hobos[[12]][,2])]),
             na.rm = TRUE)
warmMean
warmSD
```

# Statistical test

I will conduct a one-way ANOVA to confirm there are significant differences between treatments.

```{r}
tempDataLong <- rbind(tank1a, tank1b,
                      tank2a, tank2b,
                      tank3a, tank3b,
                      tank4a, tank4b,
                      tank5a, tank5b,
                      tank6a, tank6b) %>% drop_na() #Create long dataframe for ANOVA and remove rows with NAs
tail(tempDataLong) #Confirm dataframe creation
```

```{r}
tempANOVA <- aov(temp ~ treatment, data = tempDataLong) #One-way ANOVA by treatment
summary(tempANOVA)[[1]][4][[1]][1] # F = 42002.94
summary(tempANOVA)[[1]][5][[1]][1] # P-value = 0
```

# Plot temperature

## Average temperature by tank and treatment

```{r}
ambientTemp <- qpcR:::cbind.na(tank1a, tank1b, tank4a, tank4b) #Use cbind.na to cbind dataframes with uneven numbers of rows, calculate mean temperature, and only retain treatment period data
ambientTemp$avgAmbientTemp <- rowMeans(ambientTemp[,c(2, 6, 10, 14)], na.rm = TRUE) #Calculate mean temperature
ambientTemp <- ambientTemp[,c(13,17)] %>% slice(-(1:630)) #Drop extra dateTime columns and only retain treatment period data
tail(ambientTemp) #Confirm formatting
```

```{r}
coldTemp <- qpcR:::cbind.na(tank2a, tank2b, tank5a, tank5b) #Use cbind.na to cbind dataframes with uneven numbers of rows
coldTemp$avgColdTemp <- rowMeans(coldTemp[,c(2, 6, 10, 14)], na.rm = TRUE) #Calculate mean temperature
coldTemp <- coldTemp[,c(13,17)] %>% slice(-(1:630)) #Drop extra dateTime columns and only retain treatment period data
tail(coldTemp) #Confirm formatting
```

```{r}
warmTemp <- qpcR:::cbind.na(tank3a, tank3b, tank6a, tank6b) #Use cbind.na to cbind dataframes with uneven numbers of rows
warmTemp$avgWarmTemp <- rowMeans(warmTemp[,c(2, 6, 10, 14)], na.rm = TRUE) #Calculate mean temperature
warmTemp <- warmTemp[,c(13,17)] %>% slice(-(1:630)) #Drop extra columns and only retain treatment period data
tail(warmTemp) #Confirm formatting
```

```{r}
allTemps <- qpcR:::cbind.na(ambientTemp,
                            coldTemp,
                            warmTemp) #Combine columns for figure into same dataframe
allTemps <- allTemps[,c(5,2,4,6)] #Drop extra dateTime and reorder columns
head(allTemps)
```

## Create plot

```{r}
plotColors <- c(brewer.pal(9, "Reds")[7],
                brewer.pal(9, "Greys")[7],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

```{r}
#Create individual plots for each temperature

warmPlot <- ggplot(data = allTemps) + 
  geom_ribbon(aes(x = dateTime, y = avgWarmTemp, ymin = warmMean - warmSD, ymax = warmMean + warmSD), fill = plotColors[1], alpha = 0.15) +
  geom_line(aes(x = dateTime, y = avgWarmTemp), color = plotColors[1]) +
  geom_hline(yintercept = 30.8, colour = plotColors[1], linetype = 3) +
  labs(x = "", y = "") + scale_y_continuous(breaks = seq(23.5, 33.5, 2)) +
  theme_classic(base_size = 15) + theme(axis.line.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank(),
                                        plot.margin = margin(2,0,1,0))

ambientPlot <- ggplot(data = allTemps) + 
  geom_ribbon(aes(x = dateTime, y = avgAmbientTemp, ymin = ambientMean - ambientSD, ymax = ambientMean + ambientSD), fill = plotColors[2], alpha = 0.15) +
  geom_line(aes(x = dateTime, y = avgAmbientTemp), color = plotColors[2]) + 
  geom_hline(yintercept = 13.8, colour = plotColors[2], linetype = 3) +
  labs(x = "", y = "Temperature (ºC)") + scale_y_continuous(limits = c(12, 16), breaks = seq(12, 16, 1)) +
  theme_classic(base_size = 15) + theme(axis.line.x = element_blank(),
                                        axis.ticks.x = element_blank(),
                                        axis.text.x = element_blank(),
                                        plot.margin = margin(2,0,1,0))

#Cold: Mean = 5.666688, SD = 0.3637059
coldPlot <- ggplot(data = allTemps) + 
  geom_ribbon(aes(x = dateTime, y = avgColdTemp, ymin = coldMean - coldSD, ymax = coldMean + coldSD), fill = plotColors[3], alpha = 0.15) +
  geom_line(aes(x = dateTime, y = avgColdTemp), color = plotColors[3]) + 
  geom_hline(yintercept = 5.7, colour = plotColors[3], linetype = 3) +
  labs(x = "", y = "") + 
  scale_x_datetime(date_breaks = "1 day", date_labels = "%B %d") +
  scale_y_continuous(limits = c(5, 7.5), breaks = seq(5, 7.5, 0.5)) + 
  theme_classic(base_size = 15) + theme(axis.text.x = element_text(angle = 40, vjust = 1.0, hjust = 1.0),
                                        plot.margin = margin(2,0,1,0))

warmPlot / ambientPlot / coldPlot #Assemble individual plots using patchwork
ggsave("exp-temp.pdf", width = 11, height = 8.5)
```
