---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "12/29/2024"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/WA/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/WA/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("qpcR")
#install.packages("RColorBrewer")
require(tidyverse)
require(qpcR)
require(RColorBrewer)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
read.csv("../../../data/WA/HOBO/T8B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3)
```


```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1a <- read.csv("../../../data/WA/HOBO/T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..21435405..SEN.S.N..21435405.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1b <- read.csv("../../../data/WA/HOBO/T1B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..20239627..SEN.S.N..20239627.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2a <- read.csv("../../../data/WA/HOBO/T2A-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..21435415..SEN.S.N..21435415.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2b <- read.csv("../../../data/WA/HOBO/T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..T2B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank3a <- read.csv("../../../data/WA/HOBO/T3A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3a = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..T3A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank3b <- read.csv("../../../data/WA/HOBO/T3B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3b = Temp...C..LGR.S.N..21435416..SEN.S.N..21435416.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank4a <- read.csv("../../../data/WA/HOBO/T4A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp4a = Temp...C..LGR.S.N..21435413..SEN.S.N..21435413.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank4b <- read.csv("../../../data/WA/HOBO/T4B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp4b = Temp...C..LGR.S.N..21435199..SEN.S.N..21435199.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank5a <- read.csv("../../../data/WA/HOBO/T5A-2.csv", skip = 1,  header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5a = Temp...C..LGR.S.N..21435407..SEN.S.N..21435407.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank5b <- read.csv("../../../data/WA/HOBO/T5B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5b = Temp...C..LGR.S.N..20548770..SEN.S.N..20548770.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank6a <- read.csv("../../../data/WA/HOBO/T6A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6a = Temp...C..LGR.S.N..21435414..SEN.S.N..21435414.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank6b <- read.csv("../../../data/WA/HOBO/T6B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6b = Temp...C..LGR.S.N..21435417..SEN.S.N..21435417..LBL..T6B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank7a <- read.csv("../../../data/WA/HOBO/T7A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7a = Temp...C..LGR.S.N..20548776..SEN.S.N..20548776..LBL..T7A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank7b <- read.csv("../../../data/WA/HOBO/T7B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7b = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..T7B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank8a <- read.csv("../../../data/WA/HOBO/T8A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp8a = Temp...C..LGR.S.N..21435410..SEN.S.N..21435410.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank8b <- read.csv("../../../data/WA/HOBO/T8B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp8b = Temp...C..LGR.S.N..21435200..SEN.S.N..21435200.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment
#Retain only rows with data before the experiment ended

tank1a <- tank1a %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 10:422)
tank1b <- tank1b %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 1:186)

tank2a <- tank2a %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 1:186)
tank2b <- tank2b %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 10:421)

tank3a <- tank3a %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:420)
tank3b <- tank3b %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)

tank4a <- tank4a %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)
tank4b <- tank4b %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:423)

tank5a <- tank5a %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 1:258)
tank5b <- tank5b %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 1:70)

tank6a <- tank6a %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)
tank6b <- tank6b %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)

tank7a <- tank7a %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:421)
tank7b <- tank7b %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:421)

tank8a <- tank8a %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:421)
tank8b <- tank8b %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 1:260)
```


# CODE IN PROGRESS


# Calculate treatment conditions after ramp down

## Calculate treatment mean and SE

```{r}
hobos <- list(tank1a, tank2a, tank3a, tank5a, tank6a, tank7a,
              tank1b, tank2b, tank3b, tank5b, tank6b, tank7b) #Create a list of filenames to cycle through
hobos
```

### For each tank

```{r}
treatmentTempStats <- data.frame("tank" = c(1:3, 5:7),
                                 "meanTemp" = rep(0, times = 6),
                                 "sdTemp" = rep(0, times = 6)) #Create an empty dataframe
treatmentTempStats
```

```{r}
for (i in 1:(length(hobos)/2)) {
  treatmentTempStats[i,2] <-
    mean(rbind(hobos[[i]][,2][88:length(hobos[[i]][,2])],
               hobos[[i+6]][,2][88:length(hobos[[i+6]][,2])]), na.rm = TRUE)
  treatmentTempStats[i,3] <-
    sd(rbind(hobos[[i]][,2][88:length(hobos[[i]][,2])],
             hobos[[i+6]][,2][88:length(hobos[[i+6]][,2])]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers,either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

### For each treatment

```{r}
#Colder: Mean = 1.66, SD = 0.3231426
colderMean <- mean(rbind(hobos[[1]][,2][88:length(hobos[[1]][,2])],
                         hobos[[2]][,2][88:length(hobos[[2]][,2])],
                         hobos[[3]][,2][88:length(hobos[[3]][,2])],
                         hobos[[7]][,2][88:length(hobos[[7]][,2])],
                         hobos[[8]][,2][88:length(hobos[[8]][,2])],
                         hobos[[9]][,2][88:length(hobos[[9]][,2])]),
                   na.rm = TRUE)
colderSD <- sd(rbind(hobos[[1]][,2][88:length(hobos[[1]][,2])],
                     hobos[[2]][,2][88:length(hobos[[2]][,2])],
                     hobos[[3]][,2][88:length(hobos[[3]][,2])],
                     hobos[[7]][,2][88:length(hobos[[7]][,2])],
                     hobos[[8]][,2][88:length(hobos[[8]][,2])],
                     hobos[[9]][,2][88:length(hobos[[9]][,2])]),
               na.rm = TRUE)
colderMean
colderSD
```

```{r}
#Colder: Mean = 5.385082, SD = 0.3004675
coldMean <- mean(rbind(hobos[[4]][,2][88:length(hobos[[4]][,2])],
                       hobos[[5]][,2][88:length(hobos[[5]][,2])],
                       hobos[[6]][,2][88:length(hobos[[6]][,2])],
                       hobos[[10]][,2][88:length(hobos[[10]][,2])],
                       hobos[[11]][,2][88:length(hobos[[11]][,2])],
                       hobos[[12]][,2][88:length(hobos[[12]][,2])]),
                 na.rm = TRUE)
coldSD <- sd(rbind(hobos[[4]][,2][88:length(hobos[[4]][,2])],
                   hobos[[5]][,2][88:length(hobos[[5]][,2])],
                   hobos[[6]][,2][88:length(hobos[[6]][,2])],
                   hobos[[10]][,2][88:length(hobos[[10]][,2])],
                   hobos[[11]][,2][88:length(hobos[[11]][,2])],
                   hobos[[12]][,2][88:length(hobos[[12]][,2])]),
             na.rm = TRUE)
coldMean
coldSD
```

# Statistical test

I will conduct a one-way ANOVA to confirm there are significant differences between treatments.

```{r}
rbind(tank1a, tank1b)
```
```{r}

```


```{r}
tempDataLong <- rbind(tank1a %>%
                        dplyr::rename(temp = "temp1a"), 
                      tank1b %>%
                        dplyr::rename(temp = "temp1b"),
                      tank2a %>%
                        dplyr::rename(temp = "temp2a"), 
                      tank2b %>%
                        dplyr::rename(temp = "temp2b"),
                      tank3a %>%
                        dplyr::rename(temp = "temp3a"), 
                      tank3b %>%
                        dplyr::rename(temp = "temp3b"),
                      tank5a %>%
                        dplyr::rename(temp = "temp5a"), 
                      tank5b %>%
                        dplyr::rename(temp = "temp5b"),
                      tank6a %>%
                        dplyr::rename(temp = "temp6a"), 
                      tank6b %>%
                        dplyr::rename(temp = "temp6b"),
                      tank7a %>%
                        dplyr::rename(temp = "temp7a"), 
                      tank7b %>%
                        dplyr::rename(temp = "temp7b")) %>% 
  drop_na() #Create long dataframe for ANOVA and remove rows with NAs
tail(tempDataLong) #Confirm dataframe creation
```

```{r}
tempANOVA <- aov(temp ~ treatment, data = tempDataLong) #One-way ANOVA by treatment
summary(tempANOVA)[[1]][4][[1]][1] # F = 1028.818
summary(tempANOVA)[[1]][5][[1]][1] # P-value = 2.797163e-205
```

# Plot full temperature conditions (including ramp down)

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Blues")[9],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

```{r}
cbind(tank1a, tank1b,
      tank2a, tank2b,
      tank3a, tank3b,
      tank5a, tank5b,
      tank6a, tank6b,
      tank7a, tank7b) %>%
  dplyr::select(., c(1, 4, 
                     2, 6, 
                     10, 14, 
                     18, 22,
                     26, 30,
                     34, 38,
                     42, 46)) %>%
  rowwise() %>%
  mutate(., colderTemp = mean(c(temp1a, temp1b,
                                temp2a, temp2b,
                                temp3a, temp3b)),
         coldTemp = mean(c(temp5a, temp5b,
                           temp6a, temp6b,
                           temp7a, temp7b))) %>%
  ggplot(., aes(x = dateTime, y = colderTemp)) +
  geom_line(color = plotColors[1]) +
  geom_ribbon(aes(x = dateTime, y = coldMean, ymin = coldMean - coldSD, ymax = coldMean + coldSD), fill = plotColors[2], alpha = 0.15) +
  geom_hline(yintercept = coldMean, colour = plotColors[2], linetype = 3) +
  geom_line(aes(x = dateTime, y = coldTemp), color = plotColors[2]) +
  geom_ribbon(aes(x = dateTime, y = colderMean, ymin = colderMean - colderSD, ymax = colderMean + colderSD), fill = plotColors[1], alpha = 0.15) +
  geom_hline(yintercept = colderMean, colour = plotColors[1], linetype = 3) +
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     breaks = c(seq(0,5,1), seq(5,15,5))) +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

