---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "12/29/2024"
output: html_document
---

# Set up R Markdown document

```{bash}
mkdir ../../output/WA/01-temp-conditions #Make output directory
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/WA/01-temp-conditions")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("RColorBrewer")
#install.packages("data.table")
#install.packages("patchwork")
require(tidyverse)
require(RColorBrewer)
require(data.table)
require(patchwork)
```

```{r}
sessionInfo()
```

# Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1a <- read.csv("../../../data/WA/HOBO/T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..21435405..SEN.S.N..21435405.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1b <- read.csv("../../../data/WA/HOBO/T1B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..20239627..SEN.S.N..20239627.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2a <- read.csv("../../../data/WA/HOBO/T2A-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..21435415..SEN.S.N..21435415.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2b <- read.csv("../../../data/WA/HOBO/T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20548773..SEN.S.N..20548773..LBL..T2B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank3a <- read.csv("../../../data/WA/HOBO/T3A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3a = Temp...C..LGR.S.N..21435198..SEN.S.N..21435198..LBL..T3A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank3b <- read.csv("../../../data/WA/HOBO/T3B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp3b = Temp...C..LGR.S.N..21435416..SEN.S.N..21435416.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank4a <- read.csv("../../../data/WA/HOBO/T4A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp4a = Temp...C..LGR.S.N..21435413..SEN.S.N..21435413.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank4b <- read.csv("../../../data/WA/HOBO/T4B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp4b = Temp...C..LGR.S.N..21435199..SEN.S.N..21435199.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank5a <- read.csv("../../../data/WA/HOBO/T5A-2.csv", skip = 1,  header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5a = Temp...C..LGR.S.N..21435407..SEN.S.N..21435407.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank5b <- read.csv("../../../data/WA/HOBO/T5B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp5b = Temp...C..LGR.S.N..20548770..SEN.S.N..20548770.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank6a <- read.csv("../../../data/WA/HOBO/T6A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6a = Temp...C..LGR.S.N..21435414..SEN.S.N..21435414.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank6b <- read.csv("../../../data/WA/HOBO/T6B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp6b = Temp...C..LGR.S.N..21435417..SEN.S.N..21435417..LBL..T6B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank7a <- read.csv("../../../data/WA/HOBO/T7A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7a = Temp...C..LGR.S.N..20548776..SEN.S.N..20548776..LBL..T7A.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank7b <- read.csv("../../../data/WA/HOBO/T7B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp7b = Temp...C..LGR.S.N..20548774..SEN.S.N..20548774..LBL..T7B.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank8a <- read.csv("../../../data/WA/HOBO/T8A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp8a = Temp...C..LGR.S.N..21435410..SEN.S.N..21435410.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank8b <- read.csv("../../../data/WA/HOBO/T8B-2.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp8b = Temp...C..LGR.S.N..21435200..SEN.S.N..21435200.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment
#Retain only rows with data before the experiment ended

tank1a <- tank1a %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 10:422)
tank1b <- tank1b %>% 
  mutate(., tank = 1,
         treatment = "colder") %>%
  slice(., 1:186)

tank2a <- tank2a %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 1:186)
tank2b <- tank2b %>% 
  mutate(., tank = 2,
         treatment = "colder") %>%
  slice(., 10:421)

tank3a <- tank3a %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:420)
tank3b <- tank3b %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)

tank4a <- tank4a %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:422)
tank4b <- tank4b %>% 
  mutate(., tank = 3,
         treatment = "colder") %>%
  slice(., 10:423)

tank5a <- tank5a %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 1:258)
tank5b <- tank5b %>% 
  mutate(., tank = 5,
         treatment = "cold") %>%
  slice(., 1:70)

tank6a <- tank6a %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)
tank6b <- tank6b %>% 
  mutate(., tank = 6,
         treatment = "cold") %>%
  slice(., 10:422)

tank7a <- tank7a %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:421)
tank7b <- tank7b %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:421)

tank8a <- tank8a %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 10:421)
tank8b <- tank8b %>% 
  mutate(., tank = 7,
         treatment = "cold") %>%
  slice(., 1:260)
```

# Calculate treatment conditions after ramp down

## Calculate treatment mean and SE

```{r}
hobos <- list(tank1a, tank2a, tank3a, tank4a, tank5a, tank6a, tank7a, tank8a,
              tank1b, tank2b, tank3b, tank4b, tank5b, tank6b, tank7b, tank8b) #Create a list of filenames to cycle through
hobos
```

### For each tank

```{r}
treatmentTempStats <- data.frame("tank" = c(1:8),
                                 "meanTemp" = rep(0, times = 8),
                                 "sdTemp" = rep(0, times = 8)) #Create an empty dataframe
treatmentTempStats
```

```{r}
for (i in 1:(length(hobos)/2)) {
  treatmentTempStats[i,2] <-
    mean(rbind(hobos[[i]][,2][115:length(hobos[[i]][,2])],
               hobos[[i+8]][,2][115:length(hobos[[i+8]][,2])]), na.rm = TRUE)
  treatmentTempStats[i,3] <-
    sd(rbind(hobos[[i]][,2][115:length(hobos[[i]][,2])],
             hobos[[i+8]][,2][115:length(hobos[[i+8]][,2])]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStats) #Confirm calculations
```

I made these calculations starting at the 115th row (after 7 a.m. on 2024-07-23). However, some of the HOBO loggers were put in after that point since the original logger batteries died! I'll need to replace values for certain tanks. 

Loggers that started after 7 a.m. on 2024-07-23:

- tank 1b
- tank 2a
- tank 5a
- tank 5b
- tank 8b

```{r}
treatmentTempStats$meanTemp[1] <- mean(rbind(hobos[[1]][,2][115:length(hobos[[1]][,2])],
                                             hobos[[1+8]][,2][1:length(hobos[[1+8]][,2])]), na.rm = TRUE) #Tank 1 revised mean
treatmentTempStats$sdTemp[1] <- sd(rbind(hobos[[1]][,2][115:length(hobos[[1]][,2])],
                                         hobos[[1+8]][,2][1:length(hobos[[1+8]][,2])]), na.rm = TRUE) #Tank 1 revised SD

treatmentTempStats$meanTemp[2] <- mean(rbind(hobos[[2]][,2][1:length(hobos[[2]][,2])],
                                             hobos[[2+8]][,2][115:length(hobos[[2+8]][,2])]), na.rm = TRUE) #Tank 2 revised mean
treatmentTempStats$sdTemp[2] <- sd(rbind(hobos[[2]][,2][1:length(hobos[[2]][,2])],
                                         hobos[[2+8]][,2][115:length(hobos[[2+8]][,2])]), na.rm = TRUE) #Tank 2 revised SD

treatmentTempStats$meanTemp[5] <- mean(rbind(hobos[[5]][,2][1:length(hobos[[5]][,2])],
                                             hobos[[5+8]][,2][1:length(hobos[[5+8]][,2])]), na.rm = TRUE) #Tank 5 revised mean
treatmentTempStats$sdTemp[5] <- sd(rbind(hobos[[5]][,2][1:length(hobos[[5]][,2])],
                                         hobos[[5+8]][,2][1:length(hobos[[5+8]][,2])]), na.rm = TRUE) #Tank 5 revised SD

treatmentTempStats$meanTemp[8] <- mean(rbind(hobos[[8]][,2][115:length(hobos[[8]][,2])],
                                             hobos[[8+8]][,2][1:length(hobos[[8+8]][,2])]), na.rm = TRUE) #Tank 8 revised mean
treatmentTempStats$sdTemp[8] <- sd(rbind(hobos[[8]][,2][115:length(hobos[[8]][,2])],
                                         hobos[[8+8]][,2][1:length(hobos[[8+8]][,2])]), na.rm = TRUE) #Tank 8 revised SD

treatmentTempStats #Confirm changes
```

```{r}
write.csv(treatmentTempStats, "temp-mean-sd.csv", quote = FALSE)
```

### For each treatment

```{r}
#Colder: Mean = 2.137014, SD = 0.3486562
colderMean <- mean(rbind(hobos[[1]][,2][115:length(hobos[[1]][,2])],
                         hobos[[2]][,2][1:length(hobos[[2]][,2])],
                         hobos[[3]][,2][115:length(hobos[[3]][,2])],
                         hobos[[4]][,2][115:length(hobos[[4]][,2])],
                         hobos[[9]][,2][1:length(hobos[[9]][,2])],
                         hobos[[10]][,2][115:length(hobos[[10]][,2])],
                         hobos[[11]][,2][115:length(hobos[[11]][,2])],
                         hobos[[12]][,2][115:length(hobos[[11]][,2])]),
                   na.rm = TRUE)
colderSD <- sd(rbind(hobos[[1]][,2][115:length(hobos[[1]][,2])],
                     hobos[[2]][,2][1:length(hobos[[2]][,2])],
                     hobos[[3]][,2][115:length(hobos[[3]][,2])],
                     hobos[[4]][,2][115:length(hobos[[4]][,2])],
                     hobos[[9]][,2][1:length(hobos[[9]][,2])],
                     hobos[[10]][,2][115:length(hobos[[10]][,2])],
                     hobos[[11]][,2][115:length(hobos[[11]][,2])],
                     hobos[[12]][,2][115:length(hobos[[11]][,2])]),
               na.rm = TRUE)
colderMean
colderSD
```

```{r}
#Cold: Mean = 5.199698, SD = 0.2157631
coldMean <- mean(rbind(hobos[[5]][,2][1:length(hobos[[5]][,2])],
                       hobos[[6]][,2][115:length(hobos[[6]][,2])],
                       hobos[[7]][,2][115:length(hobos[[7]][,2])],
                       hobos[[8]][,2][115:length(hobos[[8]][,2])],
                       hobos[[13]][,2][1:length(hobos[[13]][,2])],
                       hobos[[14]][,2][115:length(hobos[[14]][,2])],
                       hobos[[15]][,2][115:length(hobos[[15]][,2])],
                       hobos[[16]][,2][1:length(hobos[[16]][,2])]),
                 na.rm = TRUE)
coldSD <- sd(rbind(hobos[[5]][,2][1:length(hobos[[5]][,2])],
                   hobos[[6]][,2][115:length(hobos[[6]][,2])],
                   hobos[[7]][,2][115:length(hobos[[7]][,2])],
                   hobos[[8]][,2][115:length(hobos[[8]][,2])],
                   hobos[[13]][,2][1:length(hobos[[13]][,2])],
                   hobos[[14]][,2][115:length(hobos[[14]][,2])],
                   hobos[[15]][,2][115:length(hobos[[15]][,2])],
                   hobos[[16]][,2][1:length(hobos[[16]][,2])]),
             na.rm = TRUE)
coldMean
coldSD
```

# Statistical test

I will conduct a one-way ANOVA to confirm there are significant differences between treatments after the ramp down.

```{r}
tempDataLong <- rbind(tank1a %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp1a"), 
                      tank1b %>%
                        dplyr::rename(temp = "temp1b"),
                      tank2a %>%
                        dplyr::rename(temp = "temp2a"), 
                      tank2b %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp2b"),
                      tank3a %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp3a"),
                      tank3b %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp3b"),
                      tank4a %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp4a"), 
                      tank4b %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp4b"),
                      tank5a %>%
                        dplyr::rename(temp = "temp5a"), 
                      tank5b %>%
                        dplyr::rename(temp = "temp5b"),
                      tank6a %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp6a"), 
                      tank6b %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp6b"),
                      tank7a %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp7a"), 
                      tank7b %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp7b"),
                      tank8a %>%
                        slice(., -(1:114)) %>%
                        dplyr::rename(temp = "temp8a"), 
                      tank8b %>%
                        dplyr::rename(temp = "temp8b")) %>% 
  drop_na() #Create long dataframe for ANOVA and remove rows with NAs
tail(tempDataLong) #Confirm dataframe creation
```

```{r}
tempANOVA <- aov(temp ~ treatment, data = tempDataLong) #One-way ANOVA by treatment
summary(tempANOVA)[[1]][4][[1]][1] # F = 113927.9
summary(tempANOVA)[[1]][5][[1]][1] # P-value = 0
```

# Plot full temperature conditions (including ramp down)

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Blues")[9],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

```{r}
#Format all temperature data files as data.tables

tank1a <- as.data.table(tank1a)
tank1b <- as.data.table(tank1b)

tank2a <- as.data.table(tank2a)
tank2b <- as.data.table(tank2b)

tank3a <- as.data.table(tank3a)
tank3b <- as.data.table(tank3b)

tank4a <- as.data.table(tank4a)
tank4b <- as.data.table(tank4b)

tank5a <- as.data.table(tank5a)
tank5b <- as.data.table(tank5b)

tank6a <- as.data.table(tank6a)
tank6b <- as.data.table(tank6b)

tank7a <- as.data.table(tank7a)
tank7b <- as.data.table(tank7b)

tank8a <- as.data.table(tank8a)
tank8b <- as.data.table(tank8b)
```

```{r}
temp <- tank3a[tank2b, on = "dateTime", roll = T, rollends = F][tank3b, on = "dateTime", roll = T, rollends = F][tank4a, on = "dateTime", roll = T, rollends = F][tank4b, on = "dateTime", roll = T, rollends = F][tank6a, on = "dateTime", roll = T, rollends = F][tank6b, on = "dateTime", roll = T, rollends = F][tank7a, on = "dateTime", roll = T, rollends = F][tank7b, on = "dateTime", roll = T, rollends = F][tank8a, on = "dateTime", roll = T, rollends = F][tank1a, on = "dateTime", roll = T, rollends = F] #Use data.table rolling join to match HOBO logger data together by nearest dateTime. Use only the loggers wiht at least 411 observations
temp #Confirm changes
```

```{r}
temp2 <- tank5b[tank1b, on = "dateTime", roll = T, rollends = F][tank2a, on = "dateTime", roll = T, rollends = F][tank5a, on = "dateTime", roll = T, rollends = F][tank8b, on = "dateTime", roll = T, rollends = F][temp, on = "dateTime", roll = T, rollends = F] #Roll HOBO loggers with limited data. First file named is the smallest one, final file rolled is temp (largest file with the most data)
temp2 #Confirm changes
```

```{r}
temp2 %>%
  dplyr::select(dateTime, 
                temp1a, temp1b,
                temp2a, temp2b,
                temp3a, temp3b, 
                temp4a, temp4b,
                temp5a, temp5b,
                temp6a, temp6b,
                temp7a, temp7b,
                temp8a, temp8b) %>%
  rowwise() %>%
  mutate(., colderTemp = mean(c(temp1a, temp1b,
                                temp2a, temp2b,
                                temp3a, temp3b,
                                temp4a, temp4b), na.rm = TRUE),
         coldTemp = mean(c(temp5a, temp5b,
                           temp6a, temp6b,
                           temp7a, temp7b,
                           temp8a, temp8b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = colderTemp)) +
  geom_line(color = plotColors[1]) +
  geom_ribbon(aes(x = dateTime, y = coldMean, ymin = coldMean - coldSD, ymax = coldMean + coldSD), fill = plotColors[2], alpha = 0.15) +
  geom_hline(yintercept = coldMean, colour = plotColors[2], linetype = 3) +
  geom_line(aes(x = dateTime, y = coldTemp), color = plotColors[2]) +
  geom_ribbon(aes(x = dateTime, y = colderMean, ymin = colderMean - colderSD, ymax = colderMean + colderSD), fill = plotColors[1], alpha = 0.15) +
  geom_hline(yintercept = colderMean, colour = plotColors[1], linetype = 3) +
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     breaks = c(seq(0,5,1), seq(5,15,5))) +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

```{r}
WAconditions <- temp2 %>%
  dplyr::select(dateTime, 
                temp1a, temp1b,
                temp2a, temp2b,
                temp3a, temp3b, 
                temp4a, temp4b,
                temp5a, temp5b,
                temp6a, temp6b,
                temp7a, temp7b,
                temp8a, temp8b) %>%
  rowwise() %>%
  mutate(., colderTemp = mean(c(temp1a, temp1b,
                                temp2a, temp2b,
                                temp3a, temp3b,
                                temp4a, temp4b), na.rm = TRUE),
         coldTemp = mean(c(temp5a, temp5b,
                           temp6a, temp6b,
                           temp7a, temp7b,
                           temp8a, temp8b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = colderTemp)) +
  geom_line(color = plotColors[1]) +
  geom_ribbon(aes(x = dateTime, y = coldMean, ymin = coldMean - coldSD, ymax = coldMean + coldSD), fill = plotColors[2], alpha = 0.15) +
  geom_hline(yintercept = coldMean, colour = plotColors[2], linetype = 3) +
  geom_line(aes(x = dateTime, y = coldTemp), color = plotColors[2]) +
  geom_ribbon(aes(x = dateTime, y = colderMean, ymin = colderMean - colderSD, ymax = colderMean + colderSD), fill = plotColors[1], alpha = 0.15) +
  geom_hline(yintercept = colderMean, colour = plotColors[1], linetype = 3) +
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)",
                     breaks = c(seq(0,5,1), seq(5,15,5))) +
  ggtitle("B. WA Temperature Conditions") +
  theme_classic(base_size = 15)
#Combine data from all dataframes and select columns of interest. Use rowwise operations to calculate average temperature for each treatment at each timepoint. Plot colder temepratures for each timepoint. Add a ribbon with the overall average colder temperature and SD. Add a line for the overall average colder temperature. Repeat with the cold temperature. Modify x-axis label. Modify y-axis.
```


# Make multipanel plot

```{r}
# Load objects from MA code
attach("../../MA/01-temp-conditions/01-temp-conditions.RData")
MAconditions <- MAconditions
coldMeanMA <- coldMeanMA
coldSDMA <- coldSDMA
colderMeanMA <- colderMeanMA
colderSDMA <- colderSDMA
detach("file:../../MA/01-temp-conditions/01-temp-conditions.RData")
```

```{r}
MAconditions
```

```{r}
MAconditions / WAconditions #Create multipanel plot with patchwork
ggsave("figures/multipanel-MA-WA-temperature-plot.pdf", height = 8.5, width = 11)
```

