---
title: "03-TTR-analysis-WA"
author: "Yaamini Venkataraman"
date: '2024-08-20'
output: html_document
---

In this document I'll examine how time-to-right (TTR) varies by temperature and time for WA crabs.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/WA/03-TTR-analysis/")) #Set root directory
```

```{r}
getwd()
```

# Install packages

```{r packages, include=FALSE}
#install.packages("tidyverse")
#install.packages("plotrix")
#install.packages("RColorBrewer")
#install.packages("lme4")
#install.packages("ggpubr")
#install.packages("patchwork")
#install.packages("respirometry")
require(tidyverse)
require(plotrix)
require(RColorBrewer)
require(lme4)
require(ggpubr)
require(patchwork)
require(respirometry)
```

```{r}
sessionInfo()
```

# Import data

```{r}
rawTTR <- read.csv("../../../data/WA/time-to-right-WA-clean.csv", header = TRUE, na.strings = c("N/A", "NA")) #Import raw data
head(rawTTR) #Confirm import. Trial data is in seconds.
```

```{r}
crabMetadata <- read.csv("../../../data/WA/crab-metadata-WA.csv", header = TRUE) %>%
  dplyr::select(., c("number", "tank", "carapace.width", "carapace.length")) %>%
  dplyr::rename(crab.ID = "number") #Import metadata. Select ID, treatment, width and length columns. Rename "ID" as "crab.ID"
head(crabMetadata) #Confirm changes
```

```{r}
genotypeData <- read.csv("../../../data/WA/2024-crab-genotype-WA.csv", header = TRUE, na.strings = "") %>%
  select(., -c(timepoint:gel.date, notes)) %>%
  rename(., crab.ID = "number") %>%
  filter(., is.na(genotype) == FALSE) %>%
  mutate(., presenceC = case_when(genotype == "CC" ~ "Y",
                                  genotype == "CT" ~ "Y",
                                  genotype == "TT" ~ "N")) %>%
  mutate(., presenceT = case_when(genotype == "CC" ~ "N",
                                  genotype == "CT" ~ "Y",
                                  genotype == "TT" ~ "Y")) #Import genotype data. Remove unnecessary columns. Create a crab.ID column to match with other data and ensure it is numeric. Add columns for precense of C and T
head(genotypeData) #Confirm genotype dataframe creation
```

# Format data

```{r}
modTTR <- rawTTR %>%
  dplyr::select(., -notes) %>%
  left_join(., crabMetadata, by = c("crab.ID", "tank", "carapace.width")) %>%
  filter(., is.na(trial.1) == FALSE) %>%
  mutate(., integument.cont = case_when(integument.color == "BG" ~ 0.5,
                                        integument.color == "G" ~ 1.0,
                                        integument.color == "YG" ~ 1.5,
                                        integument.color == "Y" ~ 2.0,
                                        integument.color == "YO" ~ 2.5,
                                        integument.color == "O" ~ 3,
                                        integument.color == "RO" ~ 3.5,
                                        integument.color == "R" ~ 4)) %>%
  mutate(., treatment = case_when(tank == "T01" ~ "1.5C",
                                  tank == "T02" ~ "1.5C",
                                  tank == "T03" ~ "1.5C",
                                  tank == "T04" ~ "1.5C",
                                  tank == "T05" ~ "5C",
                                  tank == "T06" ~ "5C",
                                  tank == "T07" ~ "5C",
                                  tank == "T08" ~ "5C")) %>%
  mutate(., missing.swimmer = case_when(missing.swimmer == "Y" ~ missing.swimmer,
                                        missing.swimmer == "N" ~ missing.swimmer,
                                        missing.swimmer == "Y " ~ "Y")) %>%
  mutate(., trial.1 = na_if(trial.1, 91.00)) %>%
  mutate(., trial.2 = na_if(trial.2, 91.00)) %>%
  mutate(., trial.3 = na_if(trial.3, 91.00)) %>%
  rowwise(.) %>% 
  mutate(., TTRavg = mean(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  mutate(., TTRSE = std.error(c(trial.1, trial.2, trial.3), na.rm = TRUE)) %>%
  left_join(x = ., y = (genotypeData %>% dplyr::select(crab.ID, genotype:presenceT)), by = "crab.ID") %>%
  ungroup(.) #Remove sampling ID and notes columns. Use case_when to create a column of continuous integument color metrics. Change all 91 to NA, calculate average and SE using data from the three TTR trials using rowwise operations. Ungroup. Join with genotype data.
head(modTTR) #Confirm formatting
```

# Data exploration

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Blues")[9],
                brewer.pal(9, "Blues")[7]) #Create color scheme
```

## Treatment

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = treatment)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "Timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  scale_color_manual(values = c(plotColors[1], plotColors[2]), 
                     name = "Temperature (ºC)",
                     breaks = c("1.5C", "5C"),
                     labels = c("1.5", "5")) +
  theme_classic(base_size = 15) #Plot average TTR and add lm fits without SE bars. Modify x-axis to only show experimental timepoints where TTR was measured. Assign colors to each treatment. Increase base font size.
ggsave("figures/time-to-right-raw.pdf", height = 8.5, width = 11)
```

I'm now going to make plots for various other factors to see if any of these need to be taken into consideration!

## Sex

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = sex)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-sex.pdf", height = 8.5, width = 11)
```

## Tank

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = tank)) +
  geom_jitter(alpha = 0.75) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-tank.pdf", height = 8.5, width = 11)
```

## Integument color

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = integument.color)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  scale_color_manual(values = c("#66C2A5","#A6D96A","#D9EF8B","#FFFF99","#FDAE61","#F46D43","#D53E4F"),
                     name = "Integument",
                     breaks = c("BG", "G", "YG", "Y", "YO", "O", "RO"),
                     labels = c("BG", "G", "YG", "Y", "YO", "O", "RO")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-integument.pdf", height = 8.5, width = 11)
```

## Missing swimmer

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = as.factor(missing.swimmer))) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-missswim.pdf", height = 8.5, width = 11)
```

## Weight

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = weight)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-weight.pdf", height = 8.5, width = 11)
```

## Carapace width

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = carapace.width)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 100, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-carapaceWidth.pdf", height = 8.5, width = 11)
```

## Genotype

```{r}
modTTR %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRavg, color = genotype)) +
  geom_jitter(alpha = 0.8) +
  scale_x_continuous(name = "timepoint",
                     breaks = unique(modTTR$timepoint)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = seq(0, 95, 10)) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-genotype.pdf", height = 8.5, width = 11)
```

```{r}
genotypeData %>%
  filter(., is.na(genotype) == FALSE) %>%
  group_by(., presenceT) %>%
  summarize(., count = n()) %>%
  ggplot(mapping = aes(x = "", y = count, fill = presenceT)) +
  geom_col(color = "black") +
  coord_polar(theta = "y") +
  scale_fill_manual(name = "Genotype",
                    values = c("orange", "blue")) +
  theme_void(base_size = 15) #Group data by genotype and summarize. Plot counts as a pie chart.
ggsave("../../../../Desktop/mockdata.jpg", height = 8.5, width = 11)
```


```{r}
genotypeData %>%
  filter(., is.na(genotype) == FALSE) %>%
  group_by(., genotype) %>%
  summarize(., count = n()) %>%
  ggplot(mapping = aes(x = "", y = count, fill = genotype)) +
  geom_col(color = "black") +
  coord_polar(theta = "y") +
  scale_fill_manual(name = "Genotype",
                    values = c("grey90", "grey45", "grey0")) +
  theme_void(base_size = 15) #Group data by genotype and summarize. Plot counts as a pie chart.
ggsave("figures/genotype-pie-chart.jpg", height = 8.5, width = 11)
```

### Hardy-Weinberg equilibrium

Since I have the genotype data, I want to see if my population is in Hardy-Weinbergy equilibrium. I will need to do this for the full MA population, as well as for the individual New Bedford and Eel Pond populations. I used [this calculator](https://www.cog-genomics.org/software/stats) to obtain p-values.

```{r}
genSummarized <- genotypeData %>%
  filter(., is.na(genotype) == FALSE) %>%
  group_by(., genotype) %>%
  summarize(., count = n())
genSummarized
```

Full population p-value: 0.00169236548011

# Assess differences in TTR based on treatment and time

I will use a linear mixed effects model with individual as a random effect to analyze the data, following [this tutorial](https://jontalle.web.engr.illinois.edu/MISC/lme4/bw_LME_tutorial.pdf).

```{r}
hist(modTTR$TTRavg)
hist(x = log(modTTR$TTRavg)) #Log transformation helps normalize data
```

```{r}
TTRmodel <- lmer(log(TTRavg) ~ treatment + timepoint + treatment:timepoint + 
                   as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                   (1|crab.ID), 
                 data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run model with treatment, timepoint, their interaction, sex, integument color, carapace width, weight, and missing swimmer information as fixed effects, and individual as a random effect. Change the likelihood estimator (REML = FALSE) to compare models with LRT later. Use data without rows with NAs
summary(TTRmodel) #Look at model output.
#Residuals explain more random effect variation than crab ID, crab ID explains very little variation
```

Now that I've run the full model, I want to test the importance of treatment, timepoint, and their interaction as predictors. I'll include all other factors in the null model to account for any variation they may contribute.

```{r}
TTRnullTreatment <- lmer(log(TTRavg) ~ timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run null model without treatment to identify significance of predictor
anova(TTRnullTreatment, TTRmodel) #Treatment is a significant predictor
```

```{r}
TTRnullTimepoint <- lmer(log(TTRavg) ~ treatment + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run null model without timepoint to identify significance of predictor
anova(TTRnullTreatment, TTRmodel) #Timepoint is a significant predictor
```

```{r}
TTRnullInteraction <- lmer(log(TTRavg) ~ treatment + timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE)), REML = FALSE) #Run model without interaction to identify significance of predictor. The full model constructed does not include an interaction
anova(TTRnullInteraction, TTRmodel) #Interaction is a significant predictor
```

```{r}
TTRModelComparisons<- rbind(broom::tidy(anova(TTRnullTreatment, TTRmodel)),
                            broom::tidy(anova(TTRnullTimepoint, TTRmodel)),
                            broom::tidy(anova(TTRnullInteraction, TTRmodel))) #Create table for model comparison output
TTRModelComparisons
```

```{r}
write.csv(TTRModelComparisons, "ttr-model-comparison-stat-output.csv", quote = FALSE, row.names = FALSE) #Save model output
```

# Effect of genotype

## Impact on average TTR

Now that I know treatment, time, and their interaction significantly impact TTR, I want to add in genotype information.

```{r}
TTRmodelReduced <- lmer(log(TTRavg) ~ treatment + timepoint + treatment:timepoint + 
                          as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                          (1|crab.ID), 
                        data = (modTTR %>% filter(., is.na(TTRavg) == FALSE) %>% filter(., is.na(genotype) == FALSE)), REML = FALSE) #Run model with treatment, timepoint, their interaction, sex, integument color, carapace width, weight, and missing swimmer information as fixed effects, and individual as a random effect. Change the likelihood estimator (REML = FALSE) to compare models with LRT later. Use data without rows with NAs
summary(TTRmodelReduced) #Look at model output.
#Residuals explain more random effect variation than crab ID, crab ID explains very little variation
```

```{r}
TTRfullGenotype <- lmer(log(TTRavg) ~ treatment + timepoint + treatment:timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + as.factor(genotype) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE) %>% filter(., is.na(genotype) == FALSE)), REML = FALSE) #Run model with genotype to identify significance of predictor
anova(TTRmodelReduced, TTRfullGenotype) #No significant impact of genotype
```

```{r}
TTRpresenceC <- lmer(log(TTRavg) ~ treatment + timepoint + treatment:timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + as.factor(presenceC) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE) %>% filter(., is.na(genotype) == FALSE)), REML = FALSE) #Run model with genotype to identify significance of predictor
anova(TTRmodelReduced, TTRpresenceC) #No significant impact of C allele
```

```{r}
TTRpresenceT <- lmer(log(TTRavg) ~ treatment + timepoint + treatment:timepoint + as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + as.factor(presenceT) + (1|crab.ID), data = (modTTR %>% filter(., is.na(TTRavg) == FALSE) %>% filter(., is.na(genotype) == FALSE)), REML = FALSE) #Run model with genotype to identify significance of predictor
anova(TTRmodelReduced, TTRpresenceT) #No significant impact of T allele
```

```{r}
TTRModelComparisons<- rbind(broom::tidy(anova(TTRnullTreatment, TTRmodel)),
                            broom::tidy(anova(TTRnullTimepoint, TTRmodel)),
                            broom::tidy(anova(TTRnullInteraction, TTRmodel)),
                            broom::tidy(anova(TTRmodelReduced, TTRfullGenotype)),
                            broom::tidy(anova(TTRmodelReduced, TTRpresenceC)),
                            broom::tidy(anova(TTRmodelReduced, TTRpresenceT))) #Create table for model comparison output
TTRModelComparisons
```

```{r}
write.csv(TTRModelComparisons, "ttr-model-comparison-stat-output.csv", quote = FALSE, row.names = FALSE) #Save model output
```

There is no significant impact of genotype on average TTR. There are some additional factors I want to understand: how genotype impacts whether or not a crab flipped, and how genotype impacted a difference in TTR.

## Threshold analysis

### Define threshold

```{r}
modTTR <- modTTR %>%
  mutate(., TTRthresh = case_when(!(is.na(trial.1 & trial.2 & trial.3) == FALSE) ~ 0,
                                  is.na(trial.1 & trial.2 & trial.3) == FALSE ~ 1)) #Create a new column where 1 = righted within 90 seconds in all trials and 0 = failure to right within that threshold
head(modTTR) #Confirm modifications
```

### Binomial model

```{r}
TTRthreshModel <- glmer(I(TTRthresh == 1) ~ treatment + timepoint + treatment:timepoint +
                          as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                          (1|crab.ID), 
                        family = binomial(), 
                        data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create a binomial model, where 1 = success of righting wihtin the defined threshold
summary(TTRthreshModel)
```

```{r}
TTRthreshNullTreatment <- glmer(I(TTRthresh == 1) ~ timepoint +
                                  as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                                  (1|crab.ID), 
                                family = binomial(), 
                                data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create null model without treatment
anova(TTRthreshNullTreatment, TTRthreshModel) #Compare models
```

```{r}
TTRthreshNullTimepoint <- glmer(I(TTRthresh == 1) ~ treatment +
                                  as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                                  (1|crab.ID), 
                                family = binomial(), 
                                data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create null model without timepoint
anova(TTRthreshNullTimepoint, TTRthreshModel) #Compare models
```

```{r}
TTRthreshNullInteraction <- glmer(I(TTRthresh == 1) ~ treatment + timepoint +
                                    as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                                    (1|crab.ID), 
                                  family = binomial(), 
                                  data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create null model without interaction
anova(TTRthreshNullInteraction, TTRthreshModel) #Compare models
```

```{r}
TTRthreshModelComparisons <- rbind(broom::tidy(anova(TTRthreshNullTreatment, TTRthreshModel)),
                                   broom::tidy(anova(TTRthreshNullTimepoint, TTRthreshModel)),
                                   broom::tidy(anova(TTRthreshNullInteraction, TTRthreshModel))) #Create a table of model comparison output
TTRthreshModelComparisons
```

```{r}
write.csv(TTRthreshModelComparisons, "ttr-binomial-model-comparison-output.csv", quote = FALSE, row.names = FALSE) #Save model comparison table
```

Alright, there are significant impacts of treatment and time on righting, but no impact of their interaction. So far everything is similar to the MA population. Now I want to test the impact of genotype on binary righting response. 

```{r}
TTRthreshModelRevised <- glmer(I(TTRthresh == 1) ~ treatment + timepoint +
                                 as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                                 (1|crab.ID), 
                               family = binomial(), 
                               data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create a binomial model, where 1 = success of righting wihtin the defined threshold
summary(TTRthreshModelRevised)
```

```{r}
TTRthreshGenotype <- glmer(I(TTRthresh == 1) ~ treatment + timepoint +
                             as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                             as.factor(genotype) +
                             (1|crab.ID), 
                           family = binomial(), 
                           data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create null model without interaction
anova(TTRthreshModelRevised, TTRthreshGenotype) #Compare models
```

```{r}
TTRthreshpresenceC <- glmer(I(TTRthresh == 1) ~ treatment + timepoint +
                              as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                              as.factor(presenceC) +
                              (1|crab.ID), 
                            family = binomial(), 
                            data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create null model without interaction
anova(TTRthreshModelRevised, TTRthreshpresenceC) #Compare models
```

```{r}
TTRthreshpresenceT <- glmer(I(TTRthresh == 1) ~ treatment + timepoint +
                              as.factor(sex) + integument.cont + carapace.width + weight + as.factor(missing.swimmer) + 
                              as.factor(presenceT) +
                              (1|crab.ID), 
                            family = binomial(), 
                            data = (modTTR %>% filter(., is.na(genotype) == FALSE))) #Create null model without interaction
anova(TTRthreshModelRevised, TTRthreshpresenceT) #Compare models
```

```{r}
TTRthreshModelComparisons <- rbind(broom::tidy(anova(TTRthreshNullTreatment, TTRthreshModel)),
                                   broom::tidy(anova(TTRthreshNullTimepoint, TTRthreshModel)),
                                   broom::tidy(anova(TTRthreshNullInteraction, TTRthreshModel)),
                                   broom::tidy(anova(TTRthreshModelRevised, TTRthreshGenotype)),
                                   broom::tidy(anova(TTRthreshModelRevised, TTRthreshpresenceC)),
                                   broom::tidy(anova(TTRthreshModelRevised, TTRthreshpresenceT))) #Create a table of model comparison output
TTRthreshModelComparisons
```

```{r}
write.csv(TTRthreshModelComparisons, "ttr-binomial-model-comparison-output.csv", quote = FALSE, row.names = FALSE) #Save model comparison table
```


No significant impact of genotype on whether or not a crab righted, but I suspect that's because I don't have all of the genotype data yet!

# Plots

## Average TTR plot

```{r}
modTTR %>%
  dplyr::select(., c(timepoint, treatment, TTRavg)) %>%
  filter(., is.na(TTRavg) == FALSE) %>%
  distinct(.) %>%
  ggplot(mapping = aes(x = as.character(timepoint), y = TTRavg, color = treatment, shape = treatment)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(position = position_dodge(width = 0.75)) +
  scale_y_continuous(name = "Average Time-to-Right (s)",
                     breaks = c(seq(0, 20, 5), seq(20, 100, 20))) +
  scale_x_discrete(name = "Exposure Time (Hours)",
                   breaks = unique(modTTR$timepoint),
                   labels = c(0, 4, 22, 28, 46, 94)) +
  scale_color_manual(values = c(plotColors[1], plotColors[2]),
                     name = "Temperature (ºC)",
                     breaks = c("1.5C", "5C"),
                     labels = c("1.5", "5")) +
  scale_shape_manual(values = c(15, 19),
                     name = "Temperature (ºC)",
                     breaks = c("1.5C", "5C"),
                     labels = c("1.5", "5")) +
  theme_classic(base_size = 15) #Select unique timepoint, treatment, and TTRavg data, and remove rows where TTRavg = NA. Plot average TTR for each crab in a boxplot. Do not show outliers with the OG boxplot, but add them in with geom_jitter. Assign colors and shapes to each treatment. Increase base font size.
ggsave("figures/time-to-right-avg-boxplot.pdf", height = 8.5, width = 11)
```

## Threshold analysis plot

```{r}
treatment_labels <- c("1.5C" = "1.5ºC",
                      "5C" = "5ºC") #Modify facet labels for treatment
timepoint_labels <- c("0" = "0",
                      "1" = "4",
                      "2" = "22",
                      "3" = "28",
                      "4" = "46",
                      "5" = "94") #Modify facet labels for timepoint
```

```{r}
modTTR %>%
  group_by(., treatment, timepoint) %>%
  summarize(., total = n(), 
            trial1No = sum(is.na(trial.1) == TRUE),
            trial2No = sum(is.na(trial.2) == TRUE),
            trial3No = sum(is.na(trial.3) == TRUE)) %>%
  mutate(., trial1Yes = total - trial1No,
         trial2Yes = total - trial2No,
         trial3Yes = total - trial3No) %>%
  pivot_longer(., cols = c(trial1No, trial1Yes,
                           trial2No, trial2Yes,
                           trial3No, trial3Yes), 
               values_to = "count", names_to = "condition") %>%
  mutate(., trial = case_when(condition == "trial1No" ~ 1,
                              condition == "trial1Yes" ~ 1,
                              condition == "trial2No" ~ 2,
                              condition == "trial2Yes" ~ 2,
                              condition == "trial3No" ~ 3,
                              condition == "trial3Yes" ~ 3)) %>%
  mutate(., condition = gsub(pattern = "trial1", replacement = "", x = condition),
         condition = gsub(pattern = "trial2", replacement = "", x = condition),
         condition = gsub(pattern = "trial3", replacement = "", x = condition)) %>%
  ggplot(., aes(x = trial, y = count, fill = condition)) +
  geom_bar(position = "fill", stat = "identity") + 
  facet_grid(treatment~timepoint, labeller = labeller(treatment = treatment_labels,
                                                      timepoint = timepoint_labels)) +
  labs(x = "Trial", y = "Proportion") +
  scale_fill_manual(name = "",
                    values = c("grey70", "grey10")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white"))
ggsave("figures/time-to-right-noFlips.pdf", height = 8.5, width = 11)

#Take TTR data and group by timepoint. Count the number of crabs that did not right within 90 seconds during each trial. Subtract that value from the total to get the number of crabs that did right within 90 seconds. Pivot the data longer. Create a trial column based on the condition column. Remove trial information from the text in the condition column. Pipe modified data into ggplot to create a stacked barplot. Facet by treatment and timepoint using facet labels. Relabel axes. Modify fill scale. Use specific base size with classic theme. For facet labels, use a white border (no border).
```

```{r}
modTTR %>%
  group_by(., timepoint, treatment) %>%
  summarize(., total = n(), 
            No = sum(is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE),
            Yes = sum(!(is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE))) %>%
  pivot_longer(., cols = c(Yes, No), 
               values_to = "count", names_to = "condition") %>%
  mutate(., condition = case_when(condition == "No" ~ "2N",
                                  condition == "Yes" ~ "1Y")) %>%
  ggplot(., aes(x = timepoint, y = count, fill = condition)) +
  geom_bar(position = "fill", stat = "identity") + 
  facet_wrap(~treatment, labeller = labeller(treatment = treatment_labels)) +
  scale_x_continuous(name = "Exposure Time (Hours)",
                     breaks = c(0:5),
                     labels = c(0, 4, 22, 28, 46, 94)) +
  scale_y_continuous(name = "Proportion") +
  scale_fill_manual(name = "",
                    values = c("grey70", plotColors[1]),
                    labels = c("Yes", "No")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white"))
ggsave("figures/time-to-right-noFlips-timepoint.pdf", height = 8.5, width = 11)
```


```{r}
presenceT_labels <- c("N" = "CC",
                      "Y" = "CT or TT") #Modify facet labels for precense of the T allele
```

```{r}
modTTR %>%
  filter(., is.na(genotype) == FALSE) %>%
  group_by(., treatment, genotype) %>%
  summarize(., total = n(), 
            trial1No = sum(is.na(trial.1) == TRUE),
            trial2No = sum(is.na(trial.2) == TRUE),
            trial3No = sum(is.na(trial.3) == TRUE)) %>%
  mutate(., trial1Yes = total - trial1No,
         trial2Yes = total - trial2No,
         trial3Yes = total - trial3No) %>%
  pivot_longer(., cols = c(trial1No, trial1Yes,
                           trial2No, trial2Yes,
                           trial3No, trial3Yes), 
               values_to = "count", names_to = "condition") %>%
  mutate(., trial = case_when(condition == "trial1No" ~ 1,
                              condition == "trial1Yes" ~ 1,
                              condition == "trial2No" ~ 2,
                              condition == "trial2Yes" ~ 2,
                              condition == "trial3No" ~ 3,
                              condition == "trial3Yes" ~ 3)) %>%
  mutate(., condition = gsub(pattern = "trial1", replacement = "", x = condition),
         condition = gsub(pattern = "trial2", replacement = "", x = condition),
         condition = gsub(pattern = "trial3", replacement = "", x = condition)) %>%
  ggplot(., aes(x = trial, y = count, fill = condition)) +
  geom_bar(position = "fill", stat = "identity") + 
  facet_grid(treatment~genotype, labeller = labeller(treatment = treatment_labels)) +
  labs(x = "Trial", y = "Proportion") +
  scale_fill_manual(name = "",
                    values = c("grey70", "grey10")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white"))
ggsave("figures/time-to-right-noFlips-tempGenotype.pdf", height = 8.5, width = 11)
```

```{r}
modTTR %>%
  filter(., is.na(genotype) == FALSE) %>%
  group_by(., timepoint, presenceT) %>%
  summarize(., total = n(), 
            No = sum(is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE),
            Yes = sum(!(is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE))) %>%
  pivot_longer(., cols = c(Yes, No), 
               values_to = "count", names_to = "condition") %>%
  mutate(., condition = case_when(condition == "No" ~ "2N",
                                  condition == "Yes" ~ "1Y")) %>%
  ggplot(., aes(x = timepoint, y = count, fill = condition)) +
  geom_bar(position = "fill", stat = "identity") + 
  facet_grid(~presenceT, labeller = labeller(presenceT = presenceT_labels)) +
  scale_x_continuous(name = "Hours",
                     breaks = c(0:5),
                     labels = c(0, 4, 22, 28, 46, 94)) +
  scale_y_continuous(name = "Proportion") +
  scale_fill_manual(name = "",
                    values = c("grey70", plotColors[1]),
                    labels = c("Yes", "No")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white"))
ggsave("figures/time-to-right-noFlips-timepointT.pdf", height = 8.5, width = 11)
```

```{r}
modTTR %>%
  filter(., is.na(genotype) == FALSE) %>%
  group_by(., presenceT) %>%
  summarize(., total = n(), 
            trial1No = sum(is.na(trial.1) == TRUE),
            trial2No = sum(is.na(trial.2) == TRUE),
            trial3No = sum(is.na(trial.3) == TRUE)) %>%
  mutate(., trial1Yes = total - trial1No,
         trial2Yes = total - trial2No,
         trial3Yes = total - trial3No) %>%
  pivot_longer(., cols = c(trial1No, trial1Yes,
                           trial2No, trial2Yes,
                           trial3No, trial3Yes), 
               values_to = "count", names_to = "condition") %>%
  mutate(., trial = case_when(condition == "trial1No" ~ 1,
                              condition == "trial1Yes" ~ 1,
                              condition == "trial2No" ~ 2,
                              condition == "trial2Yes" ~ 2,
                              condition == "trial3No" ~ 3,
                              condition == "trial3Yes" ~ 3)) %>%
  mutate(., condition = gsub(pattern = "trial1", replacement = "", x = condition),
         condition = gsub(pattern = "trial2", replacement = "", x = condition),
         condition = gsub(pattern = "trial3", replacement = "", x = condition)) %>%
  ggplot(., aes(x = trial, y = count, fill = condition)) +
  geom_bar(position = "fill", stat = "identity") + 
  facet_grid(~presenceT, labeller = labeller(presenceT = presenceT_labels)) +
  labs(x = "Trial", y = "Proportion") +
  scale_fill_manual(name = "",
                    values = c("grey70", "grey10")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white"))
ggsave("figures/time-to-right-noFlips-trialT.pdf", height = 8.5, width = 11)
```

```{r}
modTTR %>%
  filter(., is.na(genotype) == FALSE) %>%
  filter(., is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE) %>%
  group_by(., timepoint, genotype) %>%
  summarize(., total = n(), 
            No = sum(is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE),
            Yes = sum(!(is.na(trial.1) == TRUE | is.na(trial.2) == TRUE | is.na(trial.3) == TRUE))) %>%
  pivot_longer(., cols = c(Yes, No), 
               values_to = "count", names_to = "condition") %>%
  ggplot(., aes(x = timepoint, y = count, fill = genotype)) +
  geom_bar(position = "fill", stat = "identity") + 
  labs(y = "Proportion") +
  scale_x_continuous(name = "Hours",
                     breaks = c(0:5),
                     labels = c(0, 4, 22, 28, 46, 94)) +
  scale_fill_manual(name = "Genotype",
                    values = c("grey90", "grey45", "grey0"),
                    labels = c("CC", "CT","TT")) +
  theme_classic(base_size = 15)
ggsave("figures/time-to-right-noFlips-timeT.pdf", height = 8.5, width = 11)
#Take TTR data and group by timepoint. Count the number of crabs that did not right within 90 seconds during each trial. Subtract that value from the total to get the number of crabs that did right within 90 seconds. Pivot the data longer. Create a trial column based on the condition column. Remove trial information from the text in the condition column. Pipe modified data into ggplot to create a stacked barplot. Facet by treatment and timepoint using facet labels. Relabel axes. Modify fill scale. Use specific base size with classic theme. For facet labels, use a white border (no border).
```


# Individual-level analysis

## Q10 analysis

Q10 is a metric used to quantify temperature sensitivity, generally defined as the change in a physiological response over a 10 degree temperature interval. For most organisms, Q10 is around 2 (a two-fold change in a physiological or chemical response due to a 10 degree temperature change).

```{r}
modTTRQ10 <- modTTR %>%
  select(., timepoint, crab.ID, TTRavg, treatment, genotype) %>%
  group_by(., crab.ID) %>%
  pivot_wider(., names_from = timepoint, values_from = TTRavg, names_prefix = "timepoint") %>%
  mutate_all(~ifelse(is.nan(.), 91, .)) %>%
  filter(., is.na(timepoint0) == FALSE) %>%
  filter(., is.na(timepoint1) == FALSE) %>%
  filter(., is.na(timepoint2) == FALSE) %>%
  filter(., is.na(timepoint3) == FALSE) %>%
  filter(., is.na(timepoint4) == FALSE) %>%
  filter(., is.na(timepoint5) == FALSE) %>%
  mutate(temperature = case_when(treatment == "1.5C" ~ 1.5,
                                 treatment == "5C" ~ 5)) %>%
  arrange(., temperature) %>%
  mutate(., timepoint1_Q10 = Q10(R2 = timepoint0, 
                                 R1 = timepoint1, 
                                 T2 = 15, 
                                 T1 = temperature)) %>%
  mutate(., timepoint2_Q10 = Q10(R2 = timepoint0, 
                                 R1 = timepoint2, 
                                 T2 = 15, 
                                 T1 = temperature)) %>%
  mutate(., timepoint3_Q10 = Q10(R2 = timepoint0, 
                                 R1 = timepoint3, 
                                 T2 = 15, 
                                 T1 = temperature)) %>%
  mutate(., timepoint4_Q10 = Q10(R2 = timepoint0, 
                                 R1 = timepoint4, 
                                 T2 = 15, 
                                 T1 = temperature)) %>% 
  mutate(., timepoint5_Q10 = Q10(R2 = timepoint0, 
                                 R1 = timepoint5, 
                                 T2 = 15, 
                                 T1 = temperature)) %>%
  select(., crab.ID, treatment, genotype,
         timepoint1_Q10, timepoint2_Q10, timepoint3_Q10, timepoint4_Q10, timepoint5_Q10) %>%
  pivot_longer(., cols = !c(crab.ID, treatment, genotype), names_to = "timepoint", values_to = "Q10") %>%
  mutate(., timepoint = gsub(x = timepoint, pattern = "_Q10", replacement = "")) %>%
  mutate(., timepoint = gsub(x = timepoint, pattern = "timepoint", replacement = "")) %>%
  mutate(., timepoint = as.numeric(timepoint))
#Select columns of interest. Add columns for presence of C or T alleles. Pivot wider. Change NaN to 91 and remove any NA measurements. Add a column for temperature based on the treatment column, and arrange data by temperature. Use the Q10 function within the respirometry package to calculate Q10 values for timepoint 0 vs. all other timepoints. Select columns of interest and pivot longer. Use a series of gsub to modify timepoint column and convert to a character format that would be organized numerically. Use a similar approach for treatment.
head(modTTRQ10)
length(modTTRQ10$crab.ID %>% unique(.)) #28 crabs
```

```{r}
modTTRQ10 %>%
  ggplot(., mapping = aes(x = timepoint, y = Q10, color = treatment, shape = treatment, group = crab.ID)) +
  geom_point(size = 2) + geom_line(lty = 3) +
  facet_wrap(. ~ treatment, labeller = labeller(treatment = treatment_labels)) +
  scale_x_discrete(name = "Exposure Time (Hours)",
                   breaks = unique(modTTR$timepoint),
                   labels = c(0, 4, 22, 28, 46, 94)) +
  scale_color_manual(values = c(plotColors[1], plotColors[2]),
                     guide = "none") +
  scale_shape_manual(values = c(15, 17),
                     guide = "none") +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white")) #Plot Q10 data with timepoint on the x-axis, Q10 on the y-axis, and color and shape defined by treatment. Group by crab.ID. Add points and lines connecting values for each crab. Facet by treatment, allow y-axis to vary by each facet, and include facet labels. Add x-axis label and specify x-axis tick labels. Add shape and color values for each treatment but do not include a legend.
ggsave("figures/time-to-right-average-individuals-Q10.pdf", height = 8.5, width = 11)
```

```{r}
modTTRQ10 %>%
  ggplot(., mapping = aes(x = timepoint, y = Q10, color = treatment, shape = treatment, group = crab.ID)) +
  geom_point(size = 2) + geom_line(lty = 3) +
  facet_grid(genotype ~ treatment, labeller = labeller(treatment = treatment_labels)) +
  scale_x_discrete(name = "Exposure Time (Hours)",
                   breaks = unique(modTTR$timepoint),
                   labels = c(0, 4, 22, 28, 46, 94)) +
  scale_color_manual(values = c(plotColors[1], plotColors[2]),
                     guide = "none") +
  scale_shape_manual(values = c(15, 17),
                     guide = "none") +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white")) #Plot Q10 data with timepoint on the x-axis, Q10 on the y-axis, and color and shape defined by treatment. Group by crab.ID. Add points and lines connecting values for each crab. Facet by treatment, allow y-axis to vary by each facet, and include facet labels. Add x-axis label and specify x-axis tick labels. Add shape and color values for each treatment but do not include a legend.
ggsave("figures/time-to-right-average-individuals-Q10-genotype.pdf", height = 8.5, width = 11)
```

## Crabs tracked throught the experiment

I tracked certain individuals throughout the course of the experiment, so I want to see how their responses changed over time.

```{r}
left_join(modTTRQ10, modTTR, by = c("crab.ID", "treatment", "genotype", "timepoint")) %>%
  select(., -c(date:trial.3, sex:presenceT)) %>%
  filter(., treatment != "5C") %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRthresh, color = genotype, group = crab.ID)) +
  geom_point(size = 2) + geom_line(lty = 3) +
  facet_wrap(~crab.ID) +
  scale_x_continuous(name = "Exposure Time (Hours)",
                   breaks = 1:5,
                   labels = c(4, 22, 28, 46, 94)) +
  scale_y_continuous(name = "Right?",
                     breaks = c(0, 1),
                     labels = c("No", "Yes")) +
  scale_colour_manual(name = "Genotype",
                    values = c("grey70", "grey10"),
                    labels = c("CT", "TT")) +
  scale_shape_manual(name = "Genotype",
                    values = c(15, 19),
                    labels = c("CT", "TT")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white")) #Join TTR data with Q10 data to get individuals that were tracked throughout time. Remove superfluous columns and entries at 5ºC. Create point and line graph and wrap by crab ID.
ggsave("figures/time-to-right-noFlips-individuals-Q10.pdf", height = 8.5, width = 11)
```

```{r}
left_join(modTTRQ10, modTTR, by = c("crab.ID", "treatment", "genotype", "timepoint")) %>%
  select(., -c(Q10:trial.3, sex:presenceT)) %>%
  filter(., treatment != "5C") %>%
  pivot_wider(., names_from = "timepoint", names_prefix = "timepoint", values_from = "TTRthresh") %>%
  mutate(., pattern = case_when(timepoint2 == 0 & timepoint3 == 0 & timepoint4 == 0 & timepoint5 == 0 ~ "4",
                                timepoint2 == 1 & timepoint3 == 1 & timepoint4 == 1 & timepoint5 == 1 ~ "1",
                                (timepoint2 == 0 | timepoint3 == 0 | timepoint4 == 0) & timepoint5 == 1 ~ "2",
                                (timepoint3 == 1 | timepoint4 == 1) & timepoint5 == 0 ~ "3")) %>%
  group_by(., genotype, pattern) %>%
  summarize(., count = n()) %>%
  ggplot(., mapping = aes(x = genotype, y = count, fill = pattern)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(y = "Proportion") +
  scale_x_discrete(name = "Genotype",
                 labels = c("CT", "TT")) +
  scale_fill_manual(name = "Pattern",
                    labels = c("Always", "Recovered", "Failed", "Never"),
                    values = c("grey90", "grey60", "grey30", "grey0")) +
  theme_classic(base_size = 15) #Take data and pivot timepoint and threshold information. Assign patterns where 1 = always righting, 2 = recovered, 3 = failure to recover, and 4 = never righting. Group data by genotype and pattern and generate counts. Create stacked barplot.
ggsave("figures/time-to-right-average-individuals-Q10-genotype-patterns.pdf", height = 8.5, width = 11)
```


```{r}
modTTR %>%
  filter(., is.na(genotype) == FALSE) %>%
  dplyr::select(., crab.ID, treatment, genotype, timepoint, TTRthresh) %>%
  filter(., timepoint != 0) %>%
  filter(., treatment == "1.5C") %>%
  group_by(., crab.ID) %>%
  arrange(., crab.ID) %>%
  filter(n() > 3) %>%
  ggplot(., mapping = aes(x = timepoint, y = TTRthresh, color = genotype, shape = genotype, group = crab.ID)) +
  geom_point(size = 2) + geom_line(lty = 3) +
  facet_wrap(~crab.ID) +
  scale_x_continuous(name = "Exposure Time (Hours)",
                   breaks = 1:5,
                   labels = c(4, 22, 28, 46, 94)) +
  scale_y_continuous(name = "Right?",
                     breaks = c(0, 1),
                     labels = c("No", "Yes")) +
  scale_colour_manual(name = "Genotype",
                    values = c("grey70", "grey10"),
                    labels = c("CT", "TT")) +
  scale_shape_manual(name = "Genotype",
                    values = c(15, 19),
                    labels = c("CT", "TT")) +
  theme_classic(base_size = 15) + theme(strip.background = element_rect(colour = "white"))
ggsave("figures/time-to-right-noFlips-individuals.pdf", height = 8.5, width = 11)
```

```{r}
modTTR %>%
  dplyr::select(., crab.ID, treatment, genotype, timepoint, TTRthresh) %>%
  filter(., timepoint != 0) %>%
  filter(., treatment == "1.5C") %>%
  group_by(., crab.ID) %>%
  arrange(., crab.ID) %>%
  filter(n() > 3) %>%
  pivot_wider(., names_from = "timepoint", names_prefix = "timepoint", values_from = "TTRthresh") %>%
  mutate(., pattern = case_when(timepoint2 == 0 & timepoint3 == 0 & timepoint4 == 0 & timepoint5 == 0 ~ "4",
                                timepoint2 == 1 & timepoint3 == 1 & timepoint4 == 1 & timepoint5 == 1 ~ "1",
                                (timepoint2 == 0 | timepoint3 == 0 | timepoint4 == 0) & timepoint5 == 1 ~ "2",
                                (timepoint3 == 1 | timepoint4 == 1) & timepoint5 == 0 ~ "3")) %>%
  filter(., is.na(pattern) == FALSE) %>%
  group_by(., genotype, pattern) %>%
  summarize(., count = n()) %>%
  ggplot(., mapping = aes(x = genotype, y = count, fill = pattern)) +
  geom_bar(position = "fill", stat = "identity") +
  labs(y = "Proportion") +
  scale_x_discrete(name = "Genotype",
                 labels = c("CC", "CT")) +
  scale_fill_manual(name = "Pattern",
                    labels = c("Always", "Recovered", "Failed", "Never"),
                    values = c("grey90", "grey60", "grey30", "grey0")) +
  theme_classic(base_size = 15) #Take data and pivot timepoint and threshold information. Assign patterns where 1 = always righting, 2 = recovered, 3 = failure to recover, and 4 = never righting. Group data by genotype and pattern and generate counts. Create stacked barplot.
ggsave("figures/time-to-right-average-individuals-genotype-patterns.pdf", height = 8.5, width = 11)
```

