---
title: "SST-plot"
author: "Yaamini Venkataraman"
date: "4/2/2025"
output: html_document
---

In this script, I will create a SST plot using data and code from C. Tepolt. I will then overlay this information onto a map of sampling sites.

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/temperature-conditions/")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r, warnings = FALSE}
#install.packages("tidyverse")
#install.packages("data.table)
#install.packages("RColorBrewer")
#install.packages("sf")
# install.packages("rnaturalearth")
# install.packages("rnaturalearthdata")
#install.packages("patchwork")
require(tidyverse)
require(data.table)
require(RColorBrewer)
require(sf)
require(rnaturalearth)
require(rnaturalearthdata)
require(patchwork)
```

```{r}
sessionInfo()
```

# SST plot

This code was written by C. Tepolt. Modifications and annotations were made by me.

## Import and format data

```{r}
data = fread("../../data/Cm-SG-ICB_2021-2024.csv") #Import SST data
head(data) #Confirm import
```

```{r}
dat = data[,.(mean=mean(Mean), min = min(Mean), max = max(Mean)), by = .(Site, Month)] #Calculate mean, min, and max by site and month over the multiple years of data used
head(dat) #Confirm formatting
```

```{r}
#Reformat variables as needed for plotting
dat$Month = as.character(dat$Month)
dat$Month = ordered(dat$Month, levels = c("1","2","3","4","5","6","7","8","9","10","11","12"), labels = c("Jan","Feb","Mar","Apr","May","Jun","Jul","Aug","Sep","Oct","Nov","Dec"))
dat$Site = ordered(dat$Site, levels = c("WA", "MA", "ME"), labels = c("Willapa Bay, WA", "Buzzards Bay, MA", "Harpswell Sound, ME"))
head(dat) #Confirm changes
```

## Create plot

```{r}
site_cols <- c(brewer.pal(9, "Purples")[7],
               brewer.pal(9, "Blues")[7],
               brewer.pal(9, "Greens")[7]) #Create color scheme: WA, MA, ME
```

```{r}
ggplot(dat, aes(x = Month, y = mean, group = Site)) +
  geom_path(aes(color = Site), linewidth = 2) +
  geom_ribbon(aes(x = Month, ymin = min, ymax = max, fill = Site), alpha = 0.5) +
  ylab("Temperature (°C)") +
  ylim(c(0, 25)) +
  scale_color_manual(values = site_cols) +
  scale_fill_manual(values = site_cols) +
  theme_classic(base_size = 15) #Plot SST temperature
ggsave("SST-figure.pdf", height = 8.5, width = 11)
```

# Map

Code pulled from this tutorial: https://r-spatial.org/r/2018/10/25/ggplot2-sf-2.html

Willapa Bay, WA: 46.657330704 -124.006333308
Buzzards Bay, MA: 41.5666644 -70.7666636
Harpswell, ME: 43.80175, -69.98421

```{r}
collectionSites <- data.frame(lat = c(46.657330704, 41.5666644, 43.80175),
                              long = c(-124.006333308, -70.7666636, -69.98421)) #Create dataframe with lat and long information for each site
```

```{r}
us_map <- ne_countries(country = "United States of America", scale = "medium", returnclass = "sf") #Create map data for the US only
```

```{r}
ggplot(data = us_map) +
    geom_sf() +
    geom_point(data = collectionSites, 
               aes(x = long, y = lat), 
               size = 5, shape = 23, fill = site_cols) +
  coord_sf(xlim = c(-126, -66),
           ylim = c(24, 50), expand = FALSE) +
  scale_x_continuous(name = "Longitude") +
  scale_y_continuous(name = "Latitude") +
  theme_bw(base_size = 15) #Create a map of the US with collection sites indicated.
ggsave("map-figure.pdf", height = 8.5, width = 11)
```

# Multipanel plot

```{r}
mapPlot <- ggplot(data = us_map) +
    geom_sf() +
    geom_point(data = collectionSites, 
               aes(x = long, y = lat), 
               size = 5, shape = c(23, 24, 22), fill = site_cols) +
  coord_sf(xlim = c(-126, -66),
           ylim = c(24, 50), expand = FALSE) +
  scale_x_continuous(name = "Longitude") +
  scale_y_continuous(name = "Latitude") +
  ggtitle("A. Collection Sites") +
  theme_bw(base_size = 15) #Create a map of the US with collection sites indicated.
mapPlot
```

```{r}
sstPlot <- ggplot(dat, aes(x = Month, y = mean, group = Site)) +
  geom_path(aes(color = Site, linetype = Site), linewidth = 2) +
  geom_ribbon(aes(x = Month, ymin = min, ymax = max, fill = Site), alpha = 0.5) +
  ylab("Temperature (°C)") +
  ylim(c(0, 25)) +
  scale_color_manual(values = site_cols,
                     name = "") +
  scale_fill_manual(values = site_cols,
                    name = "") +
  scale_linetype_manual(values = 1:3,
                        name = "") +
  ggtitle("B. Average SST (2021-2024)") +
  theme_classic(base_size = 15) + theme(legend.position = c(0.25, 0.9),
                                        legend.background = element_rect(color = NA),
                                        legend.key.width = unit(2, "line")) #Plot SST temperature
sstPlot
```

```{r}
mapPlot + sstPlot +
  plot_layout(ncol = 1)
ggsave("multipanel-SST-map-plot.pdf", width = 8.5, height = 11)
```


