---
title: "01-temp-conditions"
author: "Yaamini Venkataraman"
date: "4/2/2025"
output: html_document
---

# Set up R Markdown document

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("../../output/mult-pop-heat/01-temp-conditions/")) #Set root directory
```

```{r}
getwd()
```

#Install packages

```{r}
#install.packages("tidyverse")
#install.packages("data.table)
#install.packages("RColorBrewer")
#install.packages("patchwork")
require(tidyverse)
require(data.table)
require(RColorBrewer)
require(patchwork)
```

```{r}
sessionInfo()
```
# WA

## Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1aWA <- read.csv("../../../data/mult-pop-heat/HOBO/WA_T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..20548771..SEN.S.N..20548771.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1bWA <- read.csv("../../../data/mult-pop-heat/HOBO/WA_T1B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..21435406..SEN.S.N..21435406.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2aWA <- read.csv("../../../data/mult-pop-heat/HOBO/WA_T2A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..20548772..SEN.S.N..20548772.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2bWA <- read.csv("../../../data/mult-pop-heat/HOBO/WA_T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20239628..SEN.S.N..20239628.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment

tank1aWA <- tank1aWA %>% 
  mutate(., tank = 1) %>%
  slice(., -1177)
tank1bWA <- tank1bWA %>% 
  mutate(., tank = 1) %>%
  slice(., -1177)

tank2aWA <- tank2aWA %>% 
  mutate(., tank = 2)
tank2bWA <- tank2bWA %>% 
  mutate(., tank = 2)
```

```{r}
tempCleanWA <- cbind(tank1aWA, tank1bWA,
                     tank2aWA, tank2bWA) %>%
  dplyr::select(c(1, 
                  2, 5,
                  8, 11)) #cbind data and select necessary columns
head(tempCleanWA) #Confirm changes
```

## Calculate treatment conditions

### For each tank

```{r}
treatmentTempStatsWA <- data.frame("tank" = c(1:2),
                                   "meanTemp" = rep(0, times = 2),
                                   "sdTemp" = rep(0, times = 2)) #Create an empty dataframe
treatmentTempStatsWA #Confirm dataframe creation
```

```{r}
for (i in seq(2, 4, 2)) {
  treatmentTempStatsWA[(i/2),2] <-
    mean(rbind(tempCleanWA[[i]][3:93], 
               tempCleanWA[[i+1]][3:93]), na.rm = TRUE)
  treatmentTempStatsWA[(i/2),3] <-
    sd(rbind(tempCleanWA[[i]][3:93], 
             tempCleanWA[[i+1]][3:93]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats
head(treatmentTempStatsWA) #Confirm calculations
```

```{r}
write.csv(treatmentTempStatsWA, "temp-mean-sd-WA.csv", quote = FALSE)
```

### Overall conditions

```{r}
#Exposure conditions 

rbind(tempCleanWA$temp1a[3:93],
      tempCleanWA$temp1b[3:93],
      tempCleanWA$temp2a[3:93],
      tempCleanWA$temp2b[3:93]) %>%
  mean(.) #Mean = 30.03588
rbind(tempCleanWA$temp1a[3:93],
      tempCleanWA$temp1b[3:93],
      tempCleanWA$temp2a[3:93],
      tempCleanWA$temp2b[3:93]) %>%
  sd(.) #Mean = 0.300692
```

```{r}
tempCleanWA$dateTime[93] - tempCleanWA$dateTime[3] #Time difference of 22.5 hours
```

It doesn't seem like the loggers were active during the acclimation period. The loggers were kept in tanks after the experiment so I can use that information to roughly calculate what the acclimation conditions were.

```{r}
#Acclimation conditions 

rbind(tempCleanWA$temp1a[123:1170],
      tempCleanWA$temp1b[123:1170],
      tempCleanWA$temp2a[123:1170],
      tempCleanWA$temp2b[123:1170]) %>%
  mean(.) #Mean = 22.78631
rbind(tempCleanWA$temp1a[123:1170],
      tempCleanWA$temp1b[123:1170],
      tempCleanWA$temp2a[123:1170],
      tempCleanWA$temp2b[123:1170]) %>%
  sd(.) #Mean = 0.4322318
```

## Statistical test

I will use a Kruskal-Wallis to confirm there are significant differences between tanks.

### Format data

```{r}
tempLongWA <- tempCleanWA[3:93,] %>%
  pivot_longer(., cols = temp1a:temp2b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp1a" | logger == "temp1b" ~ "tank1",
                             logger == "temp2a" | logger == "temp2b" ~ "tank2")) %>%
  dplyr::select(-logger) %>% 
  drop_na()
head(tempLongWA)
```

### Perform Kruskal-Wallis tests

```{r}
tempLongStatsWA <- kruskal.test(temperature ~ tank, data = tempLongWA)
tempLongStatsWA #Significant differences between treatments.

broom::tidy(tempLongStatsWA) #Chi-sq: 181.8361, p = 1.925508e-41	
```

# ME

## Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1aME <- read.csv("../../../data/mult-pop-heat/HOBO/ME_T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..20548771..SEN.S.N..20548771.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1bME <- read.csv("../../../data/mult-pop-heat/HOBO/ME_T1B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..21435406..SEN.S.N..21435406.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2aME <- read.csv("../../../data/mult-pop-heat/HOBO/ME_T2A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..20548772..SEN.S.N..20548772.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2bME <- read.csv("../../../data/mult-pop-heat/HOBO/ME_T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20239628..SEN.S.N..20239628.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment

tank1aME <- tank1aME %>% 
  mutate(., tank = 1) %>%
  slice(., -276)
tank1bME <- tank1bME %>% 
  mutate(., tank = 1) %>%
  slice(., -276)

tank2aME <- tank2aME %>% 
  mutate(., tank = 2)
tank2bME <- tank2bME %>% 
  mutate(., tank = 2)
```

```{r}
tempCleanME <- cbind(tank1aME, tank1bME,
                     tank2aME, tank2bME) %>%
  dplyr::select(c(1, 
                  2, 5,
                  8, 11)) #cbind data and select necessary columns
head(tempCleanME) #Confirm changes
```

## Calculate treatment conditions

### For each tank

```{r}
treatmentTempStatsME <- data.frame("tank" = c(1:2),
                                   "meanTemp" = rep(0, times = 2),
                                   "sdTemp" = rep(0, times = 2)) #Create an empty dataframe
treatmentTempStatsME #Confirm dataframe creation
```

```{r}
for (i in seq(2, 4, 2)) {
  treatmentTempStatsME[(i/2),2] <-
    mean(rbind(tempCleanME[[i]][-c(1:184)], 
               tempCleanME[[i+1]][-c(1:184)]), na.rm = TRUE)
  treatmentTempStatsME[(i/2),3] <-
    sd(rbind(tempCleanME[[i]][-c(1:184)], 
             tempCleanME[[i+1]][-c(1:184)]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats.
head(treatmentTempStatsME) #Confirm calculations
```

```{r}
write.csv(treatmentTempStatsME, "temp-mean-sd-ME.csv", quote = FALSE)
```

### Overall conditions

```{r}
#Exposure temperature

rbind(tempCleanME$temp1a[-c(1:184)],
      tempCleanME$temp1b[-c(1:184)],
      tempCleanME$temp2a[-c(1:184)],
      tempCleanME$temp2b[-c(1:184)]) %>%
  mean(.) #Mean = 29.91406
rbind(tempCleanME$temp1a[-c(1:184)],
      tempCleanME$temp1b[-c(1:184)],
      tempCleanME$temp2a[-c(1:184)],
      tempCleanME$temp2b[-c(1:184)]) %>%
  sd(.) #Mean = 0.3027249
```

```{r}
tempCleanME$dateTime[275] - tempCleanME$dateTime[185]
```

```{r}
tempCleanME$dateTime[185] - tempCleanME$dateTime[160] #6.25 hour ramp up
tempCleanME$temp1a[185] - tempCleanME$temp1a[160] #5.8ºC difference
```


```{r}
# Acclimation temperature

rbind(tempCleanME$temp1a[c(1:180)],
      tempCleanME$temp1b[c(1:180)],
      tempCleanME$temp2a[c(1:180)],
      tempCleanME$temp2b[c(1:180)]) %>%
  mean(.) #Mean = 24.0956
rbind(tempCleanME$temp1a[c(1:180)],
      tempCleanME$temp1b[c(1:180)],
      tempCleanME$temp2a[c(1:180)],
      tempCleanME$temp2b[c(1:180)]) %>%
  sd(.) #Mean = 0.1767467
```

## Statistical test

I will use a Kruskal-Wallis to confirm there are significant differences between tanks.

### Format data

```{r}
tempLongME <- tempCleanME %>%
  slice(-c(1:184)) %>%
  pivot_longer(., cols = temp1a:temp2b, values_to = "temperature", names_to = "logger") %>%
  mutate(., tank = case_when(logger == "temp1a" | logger == "temp1b" ~ "tank1",
                             logger == "temp2a" | logger == "temp2b" ~ "tank2")) %>%
  dplyr::select(-logger) %>% 
  drop_na()
head(tempLongME)
```

### Perform Kruskal-Wallis tests

```{r}
tempLongStatsME <- kruskal.test(temperature ~ tank, data = tempLongME)
tempLongStatsME #Significant differences between treatments.

broom::tidy(tempLongStatsME) #Chi-sq: 264.8512, p = 1.503481e-59
```

# NH

## Import and format data

Outliers from when HOBO loggers recorded data while outside of the tank were removed manually.

```{r}
#Import CSV output from HOBO loggers. Skip first line of CSV and specify header. Retain second and third columns, rename columns, and format dateTime column correctly

tank1aNH <- read.csv("../../../data/mult-pop-heat/HOBO/NH_T1A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1a = Temp...C..LGR.S.N..20548771..SEN.S.N..20548771.)  %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank1bNH <- read.csv("../../../data/mult-pop-heat/HOBO/NH_T1B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp1b = Temp...C..LGR.S.N..21435406..SEN.S.N..21435406.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))

tank2aNH <- read.csv("../../../data/mult-pop-heat/HOBO/NH_T2A.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2a = Temp...C..LGR.S.N..20548772..SEN.S.N..20548772.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
tank2bNH <- read.csv("../../../data/mult-pop-heat/HOBO/NH_T2B.csv", skip = 1, header = TRUE) %>% dplyr::select(., 2:3) %>% rename(dateTime = Date.Time..GMT.04.00, temp2b = Temp...C..LGR.S.N..20239628..SEN.S.N..20239628.) %>% mutate(dateTime = as.POSIXct(dateTime, format = "%m/%d/%y%t%H:%M"))
```

```{r}
#Add columns for tank ID and treatment

tank1aNH <- tank1aNH %>% 
  mutate(., tank = 1)
tank1bNH <- tank1bNH %>% 
  mutate(., tank = 1)

tank2aNH <- tank2aNH %>% 
  mutate(., tank = 2)
tank2bNH <- tank2bNH %>% 
  mutate(., tank = 2)
```

```{r}
tempCleanNH <- cbind(tank1aNH, tank1bNH,
                     tank2aNH, tank2bNH) %>%
  dplyr::select(c(1, 
                  2, 5,
                  8, 11)) #cbind data and select necessary columns
head(tempCleanNH) #Confirm changes
```

## Calculate treatment conditions

### For each tank

```{r}
treatmentTempStatsNH <- data.frame("tank" = c(1:2),
                                   "meanTemp" = rep(0, times = 2),
                                   "sdTemp" = rep(0, times = 2)) #Create an empty dataframe
treatmentTempStatsNH #Confirm dataframe creation
```

```{r}
for (i in seq(2, 4, 2)) {
  treatmentTempStatsNH[(i/2),2] <-
    mean(rbind(tempCleanNH[[i]], 
               tempCleanNH[[i+1]]), na.rm = TRUE)
  treatmentTempStatsNH[(i/2),3] <-
    sd(rbind(tempCleanNH[[i]], 
             tempCleanNH[[i+1]]), na.rm = TRUE)
} #For each tank, combine the treatment temperature data from both HOBO loggers, either find the average or SD, and save in treatmentTempStats.
head(treatmentTempStatsNH) #Confirm calculations
```

```{r}
write.csv(treatmentTempStatsNH, "temp-mean-sd-NH.csv", quote = FALSE)
```

### Overall conditions

```{r}
rbind(tempCleanNH$temp1a,
      tempCleanNH$temp1b,
      tempCleanNH$temp2a,
      tempCleanNH$temp2b) %>%
  mean(.) #Mean = 22.5643
rbind(tempCleanNH$temp1a,
      tempCleanNH$temp1b,
      tempCleanNH$temp2a,
      tempCleanNH$temp2b) %>%
  sd(.) #Mean = 0.3813463
```

# Plot full temperature conditions

```{bash}
mkdir figures
```

```{r}
plotColors <- c(brewer.pal(9, "Greens")[7],
                brewer.pal(9, "Oranges")[7],
                brewer.pal(9, "Purples")[7]) #Create color scheme
```

```{r}
plotWA <- tempCleanWA[1:93,] %>%
  dplyr::select(dateTime,
                temp1a, temp1b,
                temp2a, temp2b) %>%
  rowwise() %>%
  mutate(., avgTemp = mean(c(temp1a, temp1b,
                             temp2a, temp2b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = avgTemp)) +
  geom_line(color = plotColors[3]) +
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)") +
  ggtitle("A. WA Experimental Conditions") +
  theme_classic(base_size = 15)

plotME <- tempCleanME[-c(1:174),] %>%
  dplyr::select(dateTime,
                temp1a, temp1b,
                temp2a, temp2b) %>%
  rowwise() %>%
  mutate(., avgTemp = mean(c(temp1a, temp1b,
                             temp2a, temp2b), na.rm = TRUE)) %>%
  ggplot(., aes(x = dateTime, y = avgTemp)) +
  geom_line(color = plotColors[1]) +
  scale_x_datetime(name = "") +
  scale_y_continuous(name = "Temperature (ºC)") +
  ggtitle("B. ME Experimental Conditions") +
  theme_classic(base_size = 15)

plotWA
plotME
```

```{r}
plotWA / plotME
ggsave("figures/exp-temp.pdf", height = 8.5, width = 11)
```

```{r}
save.image("01-temp-conditions-mult-pop-heat.RData") #Save RData to load in another script.
```

